### 2013-04-29 ms.tex
File upload

%--------------------------------
section{Introduction}

- Contribution of neutral theory to ecology: the recognition of the importance of dispersal in structuring communities. The provision of a

- Gave roots and motivation for the metacommunity concepts. Present the different perspectives.

- Objective of the paper 
	- Already several reviews of neutral theory in the literature
	- We will use this opportunity to explore in more details the contribution of neutral theory to our understanding of metacommunities
	- We will pay a particular attention to recent developments at the crossroad of network theory
	- Move from a perspective where dispersal is either global or spatially explicity (e.g. over a lattice), and spatial constant, to a perspective focusing on the variance of dispersal. 
	- More specifically, the objective of the paper is to use neutral theory to better reveal the importance of landscape network structure on the distribution of diversity.
	- Referred as a "spatial contingency" 
	- Relevance to limnology

- Structure:
	- Start with a review of the main approaches to describe spatial networks and studies investigating them
	- Then we describe three simple toy models of metacommunity dynamics, taking this opportunity to review their assumptions and functioning.
	- R scripts for the toy models and all simulations conducted for this paper are provided as supplementary materiel.
	- We then conduct simple simulations of these models to reveal the impact of spatial network structure on diversity distribution. 
	- We conclude with a discussion on the operationally of the framework.

%--------------------------------
section{Network representation of landscapes}

% Table 1: main descriptors of spatial networks. Column 1: Name of the metric; Column 2: Definition
% We might want to distinguish network level versus node-level metrics.

%--------------------------------
section{Model description}

For each model:
	- description of the dynamics
	- description of the algorithm
	- main predictions

%--------------------------------
section{Results}

- Alpha diversity
- Beta diversity

%--------------------------------
section{Conclusion}

### 2013-04-29 codeR/simulations.R

@@ -1,0 +1,225 @@
+########################################################################
+# Source for the figures 1-3 in the paper
+# Gravel, Poisot and Desjardins. 2013. XXX. J. Limnology. 
+#
+# By: Dominique Gravel (dominique_gravel@uqar.ca) 
+# April 2013
+#
+########################################################################
+source("spatial_graphs.R")
+source("patch_model.R")
+source("lottery_model.R")
+library(vegan)
+
+###############################
+# Useful functions
+alpha_fn = function(pres) {
+	s = apply(pres,1,sum)
+	vec.col = numeric(length(s))
+	RK = rank(s)
+	for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
+	return(vec.col)	
+}
+
+plot_alpha = function(spatial_graph, vec.col) {
+  	# Plots a spatial graph 
+  	#
+  	# Args:
+  	#   spatia_graph: the output of one of the spatial graphs
+  	#
+	par(mar=c(1,1,1,1))
+	XY = spatial_graph[[1]]
+	adjMat = spatial_graph[[2]]
+
+	plot(XY[,1],XY[,2],xlab = "", ylab = "",cex = 1.5,labels = F)
+	adjVec = stack(as.data.frame(adjMat))[,1]
+	XX = expand.grid(XY[,1],XY[,1])
+	YY = expand.grid(XY[,2],XY[,2])
+	XX = subset(XX,adjVec==1)
+	YY = subset(YY,adjVec==1)
+	arrows(x0 = XX[,1],x1=XX[,2],y0 = YY[,1], y1 = YY[,2], length = 0,lwd = 0.2, col = "darkgrey")
+	points(XY[,1],XY[,2],pch=21,bg=vec.col,cex = 1.5)
+}
+
+###############################
+# General parameters
+S = 100
+n = 25
+r = 0.3
+
+###############################
+# Draw the four landscapes
+connected = connected_fn(n)
+lattice = lattice_fn(n)
+geograph = geograph_fn(n,r)
+geotree = geotree_fn(n,r)
+
+# Geographic distances among plots
+distConnected = as.matrix(dist(connected[[1]], upper = T, diag = T))
[…]

### 2013-04-29 codeR/spatial_graphs.R

@@ -1,0 +1,268 @@
+########################################################################
+# Functions to generate networks ("graphs") for spatial ecology.
+#
+# Stores the graph in a adjacency matrix
+#
+# Mains functions included:
+# - geograph: returns a random geometric graph
+# - geotree: returns a random geometric tree
+# - lattice: returns a lattice type of spatial graph
+# - connected: returns a fully connected graph
+# - plot_spatial: custom function to plot the graphs
+#
+# Associated functions:
+# - isConnected: tests if the graph is fully connected
+# - testCon: function used in isConnected
+# - spt: generates a graph of shortest path given an input graph 
+#
+# By: Dominique Gravel (dominique_gravel@uqar.ca) & Philippe Desjardins-Proulx (philippe.d.proulx@gmail.com)
+# April 2013
+########################################################################
+
+geograph_fn = function(n = 64, r = 0.32) {
+  # Generates a random geometric graph.
+  #
+  # Args:
+  #   n: Number of vertices.
+  #   r: Threshold distance to connect vertices.
+  #
+  # Returns:
+  #   A list with the xy coordinates and the adjacency matrix.
+  # Note: the algorithm tests if the graph is connected
+  attempts = 0
+  repeat {
+    attempts <- attempts + 1
+	xy = cbind(runif(n), runif(n))
+    distMat = as.matrix(dist(xy, method = 'euclidean', upper = T, diag = T))
+    adjMat = matrix(0, nr = n, nc = n)
+    adjMat[distMat < r] = 1
+    diag(adjMat) = 0
+    if (isConnected(adjMat)) {
+      return(list(xy, adjMat))
+    }
+  }
+}
+
+##############################
+geotree_fn = function(n = 64, r = 0.32) {
+  # Generates a random geometric tree.
+  #
+  # Args:
+  #   n: Number of vertices.
+  #   r: Threshold distance to connect vertices.
+  #
+  # Returns:
+  #   A list with the xy coordinatesa nd the adjacency matrix
+  source <- sample(1:n, 1)
+  g <- geograph_fn(n, r)
+  return(list(g[[1]], spt(g[[2]], source)))
+}
+
[…]

### 2013-05-01 ms.tex
File upload

@@
-%--------------------------------
-section{Introduction}
+\documentclass[letterpaper,twocolumn,showkeys]{revtex4-1}
+\usepackage[utf8]{inputenc}
+\usepackage{color,dcolumn,graphicx,hyperref}
 
-- Contribution of neutral theory to ecology: the recognition of the importance of dispersal in structuring communities. The provision of a
+\begin{document}
 
-- Gave roots and motivation for the metacommunity concepts. Present the different perspectives.
+\title{TO FIND}
+
+\author{Dominique Gravel}
+\author{Timothée Poisot}
+\author{Philippe Desjardins-Proulx}
+
+\begin{abstract}
+Background: \\
+Question: \\
+Approach: \\
+Key results: \\
+Conclusion\\
+
+\end{abstract}
+
+\keywords{metacommunity ecology, neutral theory, species
+sorting, patch dynamics, spatial networks, dispersal}
+
+\maketitle
+
+%========================================================%
+\section{Introduction}
+%DG
+
+- Contribution of neutral theory to ecology: the recognition of
+the importance of dispersal in structuring communities. Link with limnology.
+
+- Gave roots and motivation for the metacommunity concepts.
+Present the different perspectives.
 
@@
 	- Already several reviews of neutral theory in the literature
-	- We will use this opportunity to explore in more details the contribution of neutral theory to our understanding of metacommunities
-	- We will pay a particular attention to recent developments at the crossroad of network theory
-	- Move from a perspective where dispersal is either global or spatially explicity (e.g. over a lattice), and spatial constant, to a perspective focusing on the variance of dispersal. 
-	- More specifically, the objective of the paper is to use neutral theory to better reveal the importance of landscape network structure on the distribution of diversity.
+- We will use this opportunity to explore in more details the
+contribution of neutral theory to our understanding of
+metacommunities
+
+- We will pay a particular attention to recent developments at
+the crossroad of network theory
+
+- Move from a perspective where dispersal is either global or
+spatially explicity (e.g. over a lattice), and spatial constant,
+to a perspective focusing on the variance of dispersal.
+
+- More specifically, the objective of the paper is to use
+neutral theory to better reveal the importance of landscape
+network structure on the distribution of diversity.
 	- Referred as a "spatial contingency" 
@@
 - Structure:
-	- Start with a review of the main approaches to describe spatial networks and studies investigating them
-	- Then we describe three simple toy models of metacommunity dynamics, taking this opportunity to review their assumptions and functioning.
-	- R scripts for the toy models and all simulations conducted for this paper are provided as supplementary materiel.
-	- We then conduct simple simulations of these models to reveal the impact of spatial network structure on diversity distribution. 
-	- We conclude with a discussion on the operationally of the framework.
+- Start with a review of the main approaches to describe spatial
+networks and studies investigating them
+- Then we describe three simple toy models of metacommunity
+dynamics, taking this opportunity to review their assumptions
+and functioning.
+- R scripts for the toy models and all simulations conducted for
+this paper are provided as supplementary materiel.
+- We then conduct simple simulations of these models to reveal
+the impact of spatial network structure on diversity
+distribution.
+- We conclude with a discussion on the operationally of the
+framework.
 
-%--------------------------------
-section{Network representation of landscapes}
+%========================================================%
+\section{Network representation of landscapes}
+% PDP
+
+Entités discrètes --> réseaux
 
@@
 
-%--------------------------------
-section{Model description}
+%========================================================%
+\section{Model description}
+% DG
 
-For each model:
-	- description of the dynamics
-	- description of the algorithm
-	- main predictions
+General introduction to the section
 
-%--------------------------------
-section{Results}
+\subsection{Patch dynamics}
+blabla
 
-- Alpha diversity
-- Beta diversity
+\subsection{Neutral dynamics}
 
-%--------------------------------
-section{Conclusion}
+\subsection{Species sorting and mass effect}
+blabla
+
+%For each model:
+%	- description of the dynamics
+%	- description of the algorithm
+%	- main predictions
+blabla
+
+%========================================================%
+\section{Results}
+% TP
+
+\subsection{Alpha diversity}
+
+\subsection{Beta diversity}
+
+%========================================================%
+\section{Discussion}
+% TP
+
+Contrasting the three perspectives: what do we learn?
+
+New questions
+
+Making the theory operational
+
+Distribution attendue des espèces dans un paysage: utiliser la théorie neutre pour avoir un attendu (différents de la méthode classique de permutation)
+
+- Déviation locale des prédictions
+
+- Importance de la structure du paysage sur les propriétés émergentes
+
+-
+
+\end{document}
+%========================================================%

### 2013-05-04 codeR/spatial_graphs.R
Update spatial_graphs.R

@@ -29,10 +29,8 @@ geograph_fn = function(n = 64, r = 0.32) {
   # Returns:
   #   A list with the xy coordinates and the adjacency matrix.
   # Note: the algorithm tests if the graph is connected
-  attempts = 0
   repeat {
-    attempts <- attempts + 1
-	xy = cbind(runif(n), runif(n))
+    xy = cbind(runif(n), runif(n))
     distMat = as.matrix(dist(xy, method = 'euclidean', upper = T, diag = T))
     adjMat = matrix(0, nr = n, nc = n)
     adjMat[distMat < r] = 1
@@ -60,38 +58,38 @@ geotree_fn = function(n = 64, r = 0.32) {
 
 ##############################
 lattice_fn = function(n) {
-	# Generates a lattice type of graph
- 	#
- 	# Args:
- 	#   n: number of cells in the lattice
- 	#
- 	# Returns:
- 	#   A list with the xy coordinates and the adjacency matrix  
-	X = seq(0,1,by = 1/(n^0.5-1))
-	Y = seq(0,1,by = 1/(n^0.5-1))
-	XY = expand.grid(X,Y)
-	distMat = as.matrix(dist(XY,method = "euclidean", upper = T, diag = T))
-	adjMat = matrix(0, nr=n, nc=n)
-	adjMat[distMat <= 1/(n^0.5-1)*(1+1e-10)] = 1
-	diag(adjMat) = 0
-	return(list(XY,adjMat)) 	
+  # Generates a lattice type of graph
+  #
+  # Args:
+  #   n: number of cells in the lattice
+  #
+  # Returns:
+  #   A list with the xy coordinates and the adjacency matrix  
+  X = seq(0,1,by = 1/(n^0.5-1))
+  Y = seq(0,1,by = 1/(n^0.5-1))
+  XY = expand.grid(X,Y)
+  distMat = as.matrix(dist(XY,method = "euclidean", upper = T, diag = T))
+  adjMat = matrix(0, nr=n, nc=n)
+  adjMat[distMat <= 1/(n^0.5-1)*(1+1e-10)] = 1
+  diag(adjMat) = 0
+  return(list(XY,adjMat)) 	
 }
 
 ##############################
 connected_fn = function(n) {
-	# Generates a connected graph
- 	#
- 	# Args:
[…]

### 2013-05-16 ms.tex
Forka mea, regulae mea

@@
-\documentclass[letterpaper,twocolumn,showkeys]{revtex4-1}
+\documentclass[12pt]{article}
 \usepackage[utf8]{inputenc}
-\usepackage{color,dcolumn,graphicx,hyperref}
+\usepackage{color,lineno,setspace,graphicx,multirow,kpfonts}
+\usepackage[top=2.4cm,left=2.4cm,top=2.4cm,bottom=2.4cm,includefoot]{geometry}
+\usepackage[style=bes]{biblatex}
+\bibliography{./library}
 
@@
 
-\title{TO FIND}
+\linenumbers 
+\modulolinenumbers[1]
 
-\author{Dominique Gravel}
-\author{Timothée Poisot}
-\author{Philippe Desjardins-Proulx}
+\textbf{Title:}   Using neutral theory to reveal the contribution of dispersal to community assembly\\
 
-\begin{abstract}
+\textbf{Authors:}  Dominique Gravel$^{1,2,*}$, Timoth\'ee Poisot$^{1,2}$, Philippe Desjardins-Proulx$^{1,2}$\\
+
+1: Canada Research Chair on Terrestrial Ecosystems. D\'epartement de biologie, chimie et g\'eographique, Universit\'e du Qu\'ebec \`a Rimouski, 300 All\'ee des Ursulines, Qu\'ebec, Canada. G5L 3A1.\\
+
+2: Qu\'ebec Centre for Biodiversity Sciences, Stewart Biological Sciences Building, 1205 Dr.~Penfield Avenue, Montr\'eal (QC), H3A 1B1, Canada\\
+
+\textbf{Words in the abstract:}      \\
+\textbf{Words in the main text:}    \\
+\textbf{Words in the legends:}    \\
+\textbf{References:}             \\
+\textbf{Table:}                    \\
+
+\newpage
+\doublespacing
+
+%========================================================%
+
+\section*{Abstract}
 Background: \\
@@
 
-\end{abstract}
+\textbf{Keywords:} metacommunity ecology, neutral theory, species
+sorting, patch dynamics, spatial networks, dispersal\\
 
-\keywords{metacommunity ecology, neutral theory, species
-sorting, patch dynamics, spatial networks, dispersal}
-
-\maketitle
+\newpage
 
@@
 \section{Introduction}
-%DG
 
-- Contribution of neutral theory to ecology: the recognition of
-the importance of dispersal in structuring communities. Link with limnology.
+Community assembly is concerned by patterns and processes occurring at various
+spatial scales \parencite{Levin1992}. Until the development of metacommunity
+ecology, studies on community assembly were often restricted to local
+populations, with a strong focus on pairwise interactions (e.g.
+\textcite{MacArthur1972, May1973,Pimm1982,DeAngelis1992}). The emphasis on local
+communities has been vigorously criticized by \textcite{Ricklefs2008}, who has
+long recognized that local dynamics and community structure are strongly
+contingent on processes occurring at much larger spatial scales
+\parencite{Ricklefs1987}. This perspective is particularly relevant to
+limnology, where exchanges of organisms and nutrients affect community and
+ecosystem properties from the local (e.g. vertical mixing
+\parencite{Ryabov2011}) to the regional (e.g. connection of lakes
+\parencite{Leibold2004b} scales. It emphasizes the importance of dispersal
+relative to pairwise interactions in the organization of ecological communities.
+The metacommunity concept has been proposed by \textcite{Leibold2004a} as a
+novel approach to link different spatial scales in ecology. It builds on
+feedbacks between local scale processes, such as competitive interactions and
+local adaptation, and regional scale processes such as dispersal, gene flow and
+speciation. Ecologists are now required to move toward a predictive ecology,
+integrating elements of theoretical ecology \parencite{Thuiller2013}, and the
+metacommunity perspective appears naturally as the appropriate conceptual
+framework to develop the new modeling techniques required to fill this
+challenge. The development of neutral theory has been quite provocative in that
+respect, as one could see it as a step back in time. Neutral theory makes the
+provocative assumption that species are ecologically equivalent and thereby any
+variation in the environment has no impact on demography
+\parencite{Bell2000,Hubbell2001}. Only demographic stochasticity and dispersal
+drive the structure of neutral ecological communities. It therefore appears
+that, on first sight, neutral theory is useless.  We will develop in this paper
+the argument that neutral theory could be a useful tool to understand the impact
+of dispersal on community organization in landscapes of various complexities.
 
-- Gave roots and motivation for the metacommunity concepts.
-Present the different perspectives.
+Neutral theory sparked an historical debate in community ecology that is still
+lasting after more than a decade \parencite{Chave2004, Etiennee2011,
+Rosindell2012,Clark2012}. It was stimulated by the impressive ability of neutral
+models to fit several well studied empirical observations such as species
+abundance distributions and distance-decay relationships. A remarkable strenght
+of neutral theory is to provide a \emph{"formal general theory of abundance
+and diversity that will account, in a simple and economical fashion, for the
+many patterns that ecologists have documented"} \parencite{Bell2001}. Even if
+new studies rejecting neutral theory are consistently published (e.g.
+\textcite{Ricklefs2012}), there is now almost a consensus that neutral theory is
+a well-developed null hypothesis for niche theory and could even be used as an
+adequate approximation of ecological dynamics in some situations.
+\textcite{Bell2001} nicely envisionned two perspectives to neutral theory that
+are still standing today. Under the weak perspective, neutral theory provides a
+set of realistic predictions of community organization despite false
+assumptions. Even if being fundamentally wrong, neutral theory is still useful
+when used as a null hypothesis \parencite{Gotelli2006}. It is considered as an
+improvement over traditional null hypotheses based on randomization
+\parencite{Gotelli2000} because it readily integrates dispersal. The strong
+version on the other hand posits that neutral theory is a satisfying
+approximation to community dynamics and an appropriate theory to explain the
+distribution of biodiversity. It implies that the right mechanisms have been
+identified and that the consistently observed differences among species do not
+impact community organization.
 
-- Objective of the paper 
-	- Already several reviews of neutral theory in the literature
-- We will use this opportunity to explore in more details the
-contribution of neutral theory to our understanding of
-metacommunities
+Neutral theory has also been proposed as a useful tool to understand and predict
+some aspects of community dynamics. It links to an old philosophical debate
+between realism and instrumentalism \parencite{Wennekes2011}. Because every
+ecological model is a simplification of reality, any scientist has to
+subjectively decide the level of details he puts in, leaving out some elements
+judged unimportant. The realism perspective requires that all assumptions of
+theory to be true, while the utility of the theory is more important to
+instrumentalism. The utilitarian value of a theory could either be for
+understanding or prediction (another old philosophical debate, see
+\textcite{Schmueli2010}). Obviously neutral theory could only be instrumental.
+The question then is if such a 'general, large-scale, but vague' theory
+\parencite{Wennekes2011} is a satisfying approximation.
 
-- We will pay a particular attention to recent developments at
-the crossroad of network theory
+The instrumentalism view of neutral theory raises the question of why should it
+be a satisfying approximation despite knowing the pieces are wrong? We see two
+potential answers to this question. A first answer might be that stochasticity
+of various origins can blur the deterministic differences among species and
+promote the ecological drift \parencite{Gravel2011}. Much has been said the
+existence of demographic stochasticity, some ecologists even arguing that
+neutral models impede progress in community ecology by hidding niche differences
+\parencite{Clark2012}, and we therefore will keep this discussion for other
+papers. The second answer is that dispersal and historical contingencies might
+have a much more profound impact on species distribution \parencite{Bahn2007}
+and ecological dynamics. The debate over the equivalence assumption and
+demographic stochasticity has perhaps overlook the recognition of how much
+dispersal influence community assembly.
 
-- Move from a perspective where dispersal is either global or
-spatially explicity (e.g. over a lattice), and spatial constant,
-to a perspective focusing on the variance of dispersal.
+We will adopt the perspective that neutral theory is a useful tool to both
+understand and predict the impact of dispersal on community organization. Even
+for a theoretical analysis, we need a benchmark without niche differences to
+reveal the role of dispersal in structuring communities and understand the
+interaction with niche differentiation. We will explore recent applications of
+neutral theory, at the crossroad of network theory, to better represent the
+impact of landscape structure on biodiversitity distribution. This analysis will
+prove particularly relevant to limnology, where most riverine and lacustre
+habitats are characterized by a their discrete nature and spatially complex
+arrangements. We will also reveal the relative contribution of ecological
+interaction and niche differentiation by contrasting predictions of a neutral
+model to other metacommunity perspectives.
 
-- More specifically, the objective of the paper is to use
-neutral theory to better reveal the importance of landscape
-network structure on the distribution of diversity.
-	- Referred as a "spatial contingency" 
-	- Relevance to limnology
-
-- Structure:
-- Start with a review of the main approaches to describe spatial
-networks and studies investigating them
-- Then we describe three simple toy models of metacommunity
-dynamics, taking this opportunity to review their assumptions
-and functioning.
-- R scripts for the toy models and all simulations conducted for
-this paper are provided as supplementary materiel.
-- We then conduct simple simulations of these models to reveal
-the impact of spatial network structure on diversity
-distribution.
-- We conclude with a discussion on the operationally of the
-framework.
+Our main objective in this paper is to use neutral theory to stress the
+importance of landscape network structure on the distribution of diversity. We
+refer to the landscape organization as a "spatial contingency"
+\parencite{Peres-Neto2013} that could potentially affect the coexistence
+mechanisms at play. We will therefore move from a perspective where dispersal is
+either global or spatially explicity (e.g. over a lattice), and spatial
+constant, to a perspective focusing on the variance of dispersal. A second
+generation of neutral models (e.g. \textcite{Economo2008,
+Economo2011,Desjardins2012a,Desjardins2012b}, and even experiments
+\parencite{Altermat2012}, recently introduced more realistic lanscapes and found
+surprising contributions of spatial contingencies. We will start with a short
+review of the main approaches to describe spatial networks and the studies
+investigating them. Then we will describe three simple toy models of
+metacommunity dynamics, taking this opportunity to review their assumptions and
+main predictions. We provide as Supplementary Material the R scripts for the toy
+models and all simulations conducted for this paper. We then conduct simple
+simulations of these models to reveal the impact of spatial network structure on
+diversity distribution. We conclude with a discussion on the operationally of
+the framework.
 
@@
 Making the theory operational
+	- how to parameterize neutral models
+	- making predictions for specific landscapes
+	-
 
@@
 
-- Déviation locale des prédictions
+- Utiliser les déviations locales pour comprendre le rôle de la sélection
+
+- Approche par réseau: vers une approche qui intègre la variance des mécanismes de coexistence
 
 - Importance de la structure du paysage sur les propriétés émergentes
+	- importance relative de species sorting et drift va dépendre de l'organisation du paysage
+	- prédiction
+	- rôle de la distribution de degrés
+-
 
--
+%========================================================%
+\section{Acknowledgements}
+DG received financial support from NSERC and Canada Research Chair program. TP is funded by a MELS-FQRNT post-doctoral fellowship and PDP by a NSERC fellowship.
+\newpage
+
+%========================================================%
+\printbibliography
+\newpage
 

### 2013-05-16 ms.tex
Minor edits in the last paragraphs

@@
 
-We will adopt the perspective that neutral theory is a useful tool to both
-understand and predict the impact of dispersal on community organization. Even
-for a theoretical analysis, we need a benchmark without niche differences to
-reveal the role of dispersal in structuring communities and understand the
-interaction with niche differentiation. We will explore recent applications of
+In this paper, we argue that neutral theory is useful both to understand and
+predict the impact of dispersal on community organization. Even
+for purely theoretical analyses, we need a benchmark without niche differences to
+reveal the role of dispersal in structuring communities, and understand how it
+interacts with niche differentiation. We will explore recent applications of
 neutral theory, at the crossroad of network theory, to better represent the

### 2013-05-16 ms.tex
First sentence of the paragraph re. effects of landscape structure on relative importance of processes

@@
 
-The instrumentalism view of neutral theory raises the question of why should it
+The instrumentalist view of neutral theory raises the question of why it should
 be a satisfying approximation despite knowing the pieces are wrong? We see two
@@
 
+Our results point to an important, although often overlooked, consequence of
+neutral dynamics. Emerging properties of a community (i.e. the number of
+species, but also their spatial distribution and co-occurence at different local
+patches), are affected by the dispersal structure.
+
 - Importance de la structure du paysage sur les propriétés émergentes

### 2013-05-17 codeR/simulations.R
Update of the R codes

@@ -1,5 +1,5 @@
 ########################################################################
-# Source for the figures 1-3 in the paper
+# Source for the figures 1-4 in the paper
 # Gravel, Poisot and Desjardins. 2013. XXX. J. Limnology. 
 #
 # By: Dominique Gravel (dominique_gravel@uqar.ca) 
@@ -65,6 +65,18 @@ topoLattice = spm(lattice[[2]])
 topoGeograph = spm(geograph[[2]])
 topoGeotree = spm(geotree[[2]])
 
+# Degree centrality
+degGeograph = deg_cen(geograph)
+degGeotree = deg_cen(geotree)
+
+# Eigen centrality
+eigGeograph = eig_cen(geograph)
+eigGeotree = eig_cen(geotree)
+
+# Closeness centrality
+clsGeograph = cls_cen(geograph)
+clsGeotree = cls_cen(geotree)
+
 ###############################
 # Run the three models with the four landscapes
 
@@ -102,7 +114,7 @@ d = 0.1
 J = 100
 sdN = Inf
 sdE = 5
-nsteps = 10000
+nsteps = 1000
 
 # Simulations
 neutralConnected = lottery_model(m,M,d,S,J,sdE,sdN,connected,nsteps)
@@ -153,31 +165,60 @@ betaSSGeotree = as.matrix(vegdist(ssGeotree[[1]],method = "bray",diag=T,upper=T)
 
 ###############################
 # Figure 1
-quartz(height = 6, width = 6)
-t = matrix(c(1:12),nr = 4, nc = 3, byrow = T)
+quartz(height = 6, width = 3)
+t = matrix(c(1:4),nr = 4, nc = 1, byrow = T)
 layout(t)
-layout.show(12)
+layout.show(4)
 
-plot_alpha(connected,alphaPatchConnected)
+#plot_alpha(connected,alphaPatchConnected)
 plot_alpha(connected,alphaNeutralConnected)
-plot_alpha(connected,alphaSSConnected)
+#plot_alpha(connected,alphaSSConnected)
 
-plot_alpha(lattice,alphaPatchLattice)
+#plot_alpha(lattice,alphaPatchLattice)
 plot_alpha(lattice,alphaNeutralLattice)
-plot_alpha(lattice,alphaSSLattice)
[…]

### 2013-05-17 codeR/spatial_graphs.R

@@ -34,244 +34,244 @@
 # May 2013
 ########################################################################
 
-library('igraph')
+library(igraph)
 
 geograph_fn = function(n = 64, r = 0.32) {
-  # Generates a random geometric graph.
-  #
-  # Args:
-  #   n: Number of vertices.
-  #   r: Threshold distance to connect vertices.
-  #
-  # Returns:
-  #   A list with the xy coordinates and the adjacency matrix.
-  # Note: the algorithm tests if the graph is connected
-  repeat {
-    xy = cbind(runif(n), runif(n))
-    distMat = as.matrix(dist(xy, method = 'euclidean', upper = T, diag = T))
-    adjMat = matrix(0, nr = n, nc = n)
-    adjMat[distMat < r] = 1
-    diag(adjMat) = 0
-    if (connected(adjMat)) {
-      return(list(xy, adjMat))
-    }
-  }
+  # Generates a random geometric graph.
+  #
+  # Args:
+  #   n: Number of vertices.
+  #   r: Threshold distance to connect vertices.
+  #
+  # Returns:
+  #   A list with the xy coordinates and the adjacency matrix.
+  # Note: the algorithm tests if the graph is connected
+  repeat {
+    xy = cbind(runif(n), runif(n))
+    distMat = as.matrix(dist(xy, method = 'euclidean', upper = T, diag = T))
+    adjMat = matrix(0, nr = n, nc = n)
+    adjMat[distMat < r] = 1
+    diag(adjMat) = 0
+    if (isconnected(adjMat)) {
+      return(list(xy, adjMat))
+    }
+  }
 }
 
 ##############################
 geotree_fn = function(n = 64, r = 0.32) {
-  # Generates a random geometric tree.
-  #
-  # Args:
-  #   n: Number of vertices.
-  #   r: Threshold distance to connect vertices.
-  #
-  # Returns:
-  #   A list with the xy coordinatesa nd the adjacency matrix
-  source = sample(1:n, 1)
-  g = geograph_fn(n, r)
-  return(list(g[[1]], spt(g[[2]], source)))
+  # Generates a random geometric tree.
+  #
+  # Args:
+  #   n: Number of vertices.
+  #   r: Threshold distance to connect vertices.
+  #
+  # Returns:
[…]

### 2013-05-17 ms copie.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{color,lineno,setspace,graphicx,multirow,kpfonts}
\usepackage[top=2.4cm,left=2.4cm,top=2.4cm,bottom=2.4cm,includefoot]{geometry}
\usepackage[style=bes]{biblatex}
\bibliography{./library}

\begin{document}

\linenumbers 
\modulolinenumbers[1]

\textbf{Title:}   Using neutral theory to reveal the contribution of dispersal to community assembly\\

\textbf{Authors:}  Dominique Gravel$^{1,2,*}$, Timoth\'ee Poisot$^{1,2}$, Philippe Desjardins-Proulx$^{1,2}$\\

1: Canada Research Chair on Continental Ecosystem Ecology. D\'epartement de biologie, chimie et g\'eographique, Universit\'e du Qu\'ebec \`a Rimouski, 300 All\'ee des Ursulines, Qu\'ebec, Canada. G5L 3A1.\\

2: Qu\'ebec Centre for Biodiversity Sciences\\

\textbf{Words in the abstract:}      \\
\textbf{Words in the main text:}    \\
\textbf{Words in the legends:}    \\
\textbf{References:}             \\
\textbf{Table:}                    \\

\newpage
\doublespacing

%========================================================%

\section*{Abstract}
Background: \\
Question: \\
Approach: \\
Key results: \\
Conclusion\\

\textbf{Keywords:} metacommunity ecology, neutral theory, species
sorting, patch dynamics, spatial networks, dispersal\\

\newpage

%========================================================%
\section*{Introduction}

Community assembly is concerned by patterns and processes occurring at various spatial scales \parencite{Levin1992}. Until the development of metacommunity ecology, studies on community assembly were often restricted to local populations, with a strong focus on pairwise interactions (e.g. \textcite{MacArthur1972, May1973,Pimm1982,DeAngelis1992}). The emphasis on local communities has been vigorously criticized by \textcite{Ricklefs2008}, who has long recognized that local dynamics and community structure are strongly contingent on processes occurring at much larger spatial scales \parencite{Ricklefs1987}. This perspective is particularly relevant to limnology, where exchanges of organisms and nutrients impact community and ecosystem properties from the local (e.g. vertical mixing \parencite{Ryabov2011}) to the regional (e.g. connection of lakes \parencite{Leibold2004b} scales. It emphasizes the importance of dispersal relative to pairwise interactions in the organization of ecological communities. The metacommunity concept has been proposed by \textcite{Leibold2004a} as a novel approach to link different spatial scales in ecology. It builds on feedbacks between local scale processes, such as competitive interactions and local adaptation, and regional scale processes such as dispersal, gene flow and speciation. Ecologists are now required to move toward a predictive ecology, integrating elements of theoretical ecology \parencite{Thuiller2013}, and the metacommunity perspective appears naturally as the appropriate conceptual framework to develop the new modeling techniques required to fill this challenge. The development of neutral theory has been quite provocative in that respect, as one could see it as a step back in time. Neutral theory makes the provocative assumption that species are ecologically equivalent and thereby any variation in the environment has no impact on demography \parencite{Bell2000,Hubbell2001}. Only demographic stochasticity and dispersal drive the structure of neutral ecological communities. It therefore appears that, on first sight, neutral theory is useless.  We will develop in this paper the argument that neutral theory could be a useful tool to understand the impact of dispersal on community organization in landscapes of various complexities.

Neutral theory sparked an historical debate in community ecology that is still lasting after more than a decade \parencite{Chave2004, Etiennee2011, Rosindell2012,Clark2012}. It was stimulated by the impressive ability of neutral models to fit several well studied empirical observations such as species abundance distributions and distance-decay relationships. A remarkable strenght of neutral theory is to provide a \textit("formal general theory of abundance and diversity that will account, in a simple and economical fashion, for the many patterns that ecologists have documented") \parencite{Bell2001}. Even if new studies rejecting neutral theory are consistently published (e.g. \textcite{Ricklefs2012}), there is now almost a consensus that neutral theory is a well-developed null hypothesis for niche theory and could even be used as an adequate approximation of ecological dynamics in some situations. \textcite{Bell2001} nicely envisionned two perspectives to neutral theory that are still standing today. Under the weak perspective, neutral theory provides a set of realistic predictions of community organization despite false assumptions. Even if being fundamentally wrong, neutral theory is still useful when used as a null hypothesis \parencite{Gotelli2006}. It is considered as an improvement over traditional null hypotheses based on randomization \parencite{Gotelli2000} because it readily integrates dispersal. The strong version on the other hand posits that neutral theory is a satisfying approximation to community dynamics and an appropriate theory to explain the distribution of biodiversity. It implies that the right mechanisms have been identified and that the consistently observed differences among species do not impact community organization.

Neutral theory has also been proposed as a useful tool to understand and predict some aspects of community dynamics. It links to an old philosophical debate between realism and instrumentalism \parencite{Wennekes2011}. Because every ecological model is a simplification of reality, any scientist has to subjectively decide the level of details he puts in, leaving out some elements judged unimportant. The realism perspective requires that all assumptions of theory to be true, while the utility of the theory is more important to instrumentalism. The utilitarian value of a theory could either be for understanding or prediction (another old philosophical debate, see \textcite{Schmueli2010}). Obviously neutral theory could only be instrumental. The question then is if such a 'general, large-scale, but vague' theory \parencite{Wennekes2011} is a satisfying approximation.

The instrumentalism view of neutral theory raises the question of why should it be a satisfying approximation despite knowing the pieces are wrong? We see two potential answers to this question. A first answer might be that stochasticity of various origins can blur the deterministic differences among species and promote the ecological drift \parencite{Gravel2011}. Much has been said the existence of demographic stochasticity, some ecologists even arguing that neutral models impede progress in community ecology by hidding niche differences \parencite{Clark2012}, and we therefore will keep this discussion for other papers. The second answer is that dispersal and historical contingencies might have a much more profound impact on species distribution \parencite{Bahn2007} and ecological dynamics. The debate over the equivalence assumption and demographic stochasticity has perhaps overlook the recognition of how much dispersal influence community assembly.

We will adopt the perspective that neutral theory is a useful tool to both understand and predict the impact of dispersal on community organization. Even for a theoretical analysis, we need a benchmark without niche differences to reveal the role of dispersal in structuring communities and understand the interaction with niche differentiation. We will explore recent applications of neutral theory, at the crossroad of network theory, to better represent the impact of landscape structure on biodiversitity distribution. This analysis will prove particularly relevant to limnology, where most riverine and lacustre habitats are characterized by a their discrete nature and spatially complex arrangements. We will also reveal the relative contribution of ecological interaction and niche differentiation by contrasting predictions of a neutral model to other metacommunity perspectives.

Our main objective in this paper is to use neutral theory to stress the importance of landscape network structure on the distribution of diversity. We refer to the landscape organization as a "spatial contingency" \parencite{Peres-Neto2013} that could potentially impact the coexistence mechanisms at play. We will therefore move from a perspective where dispersal is either global or spatially explicity (e.g. over a lattice), and spatial constant, to a perspective focusing on the variance of dispersal. A second generation of neutral models (e.g. \textcite{Economo2008, Economo2011,Desjardins2012a,Desjardins2012b}, and even experiments \parencite{Altermat2012}, recently introduced more realistic lanscapes and found surprising contributions of spatial contingencies. We will start with a short review of the main approaches to describe spatial networks and the studies investigating them. Then we will describe three simple toy models of metacommunity dynamics, taking this opportunity to review their assumptions and main predictions. We provide as Supplementary Material the R scripts for the toy models and all simulations conducted for this paper. We then conduct simple simulations of these models to reveal the impact of spatial network structure on diversity distribution. We conclude with a discussion on the operationally of the framework.

%========================================================%
\section*{Network representation of landscapes}
% PDP

PDP

% Table 1: main descriptors of spatial networks. Column 1: Name of the metric; Column 2: Definition
% We might want to distinguish network level versus node-level metrics.

%========================================================%
\section*{Model description}

In this section we will describe three toy models representing different perspectives to metacommunity ecology: patch dynamics, neutral dynamics and species sorting. While the neutral model is interesting in itself, it is by comparison with a model without any interactions (patch dynamics) and with niche differentiation (species sorting) that we will be able to fully understand the interaction between these processes and landscape structure. Despite neutral, competitive interactions in neutral models are very strong because of the zero-sum assumption (the community is always at carrying capacity). We will first review the fundamental assumptions of each model with their description (Table 1 summarizes the parameters and variables that are used), and then briefly discuss their main predictions. Simulation results are presented in the next section, with the corresponding R code provided in the Supplementary Material.

\begin{table*}
	\centering
	\begin{tabular}{llccc}
		\hline
		 & Definition & Patch dynamics & Neutral & Species-sorting \\
		\hline
		\textbf{Variables} &     &    &    &   \\
		$p$ & Occupancy  & X &    &   \\
		$N$ & Local population size &    & X & X \\
		$Z$ & Local rel. abund. &  & X & X \\
		$P$ & Rel. abund. in the neighborhood &  & X & X \\
		$s$ & Local species richness & X & X & X \\
		$d$ & Node degree & X & X & X \\
		$C$ & Prob. of a colonization event & X &  &  \\
		$I$ & Prob. of a colonization event & X &  &  \\
		$Pr$ & Recruitment prob. &  & X & X \\
		$\lambda$ & Survival prob. &  &  & X \\

		\textbf{Indices} & & & &\\
		$x,y$ & Node location & X & X & X \\
		$i,j$ & Species & X & X & X\\
		$n$ & Microsite & & & X \\

		\textbf{Parameters} & & & &\\
		$S$ & Size of regional species pool & X & X & X \\
		$c$ & Colonization prob. & X & & \\
		$e$ & Extinction prob. & X & & \\
		$J$ & Local carrying capacity & & X & X \\
		$m$ & immigration prob. from neigh. & & X & X \\
		$M$ & immigration prob. from metaco. & & X & X \\
		$k$ & Death prob. & & X & X \\
		$u$ & Niche optimum & & & X \\
		$b$ & Niche breadth & & & X \\
		$E$ & Microsite env. conditions & & & X \\
		$\overline{E}$ & Local env. average & & & X \\
		$\sigma$ & Local env. variance & & & X \\
		$\overline{E_R}$ & Regioal env. average & & & X \\
		$\sigma_R$ & Regional env. variance & & & X \\
		\hline
	\end{tabular}
	\caption{List of variables, indices and parameters from the three models}
\end{table*}

%------------------------
\subsection*{Patch dynamics}

The simplest metacommunity model is a $S$ species extension of traditional metapopulation models \parencite(Hanski1999). The standard Levins metapopulation model \parencite(Levins1969) describes the stochastic colonizations and extinctions of population over a homogenous landscape. The basic unit is the population. The Levins model tracts the dynamics of occupancy (the fraction of the landscape that is occupied) with an ordinary differential equation and therefore assumes an infinite landscape. The simulation model we run is more realistic as it simulates a finite number $N$ of discrete patches. The rules described in the previous section were used to generate connectivity matrices along four scenarios (Fig. 1): global dispersal (connected graph), a lattice, a random geometric graph and a random tree graph. A patch $x$ shares $d_x$ links with neigbhouring patches (its degree). At each time step (the simulation model is discrete in time), the probability that a colonist coming from an occupied patch $y$ arrives at patch $x$ is $cd_y^{-1}$, where $c$ is the probability a colonization event takes place if all connected patches are occupied. The expected probability that a colonist arrives to patch $x$ from patch $y$ is then $C_{ixy}=cp_{iy}d_y^{-1}$, where $p_{i}y$ is the probability that patch $y$ is occupied by species $i$. The probability that an extinction occurs in a given patch is $e$. The Levins model is for a single species, but a basic metacommunity patch dynamics model could be run by aggregating $S$ independent metapopulation models \parencite(Hanski1997). There are no interactions in this simple model, which means there is no limit to local species richness and no carrying capacity. Competitive, mutualistic and predator-prey interactions have been added to this framework (e.g. \textcite{Tilman1994,Klausmeier1998,Holt1996}) but we will keep this model minimal for the sake of comparison with the neutral model.

Predictions of the patch dynamics metacommunity model are quite straightforward. First, a fundamental result of metapopulation ecology is that persistence will occur if colonization probability is larger than extinction probability ($c>e$). Given that all species are the same, then we should expect the regional diversity to be $S$ if this condition is satisfied and $0$ if not. The situation is however more complex in spatially explicit landscapes with complex connectivity matrices \parencite{Hanski1998}. Spatially explicit dispersal usually reduces the occupancy and thereby the likelihood of persistence. The second prediction is that, given spatial variation in connectivity, there will be spatial variation in occurrence probability. Given the above formulation of a colonization event to occur, the probability that an empty location is colonized is $I_ix=1-\prod d_x(1-C_{ixy})$. This equation basically tells us that the colonization probability will increase asymptotically with the degree of a patch (because of the product). It is easy to show from metapopulation theory that the occurrence probability in a patch is then $p_ix=I_x(I_ix+e)^{-1}$. The feedback between local and regional dynamics arises because all $p_ix$ from the landscape are dependent from each other. Simulation are usually conducted to solve the model for a large landscape, but numerical solutions are theoretically possible. The aggregation across the $S$ species of the regional species pool is obtained by taking the summation of occurrence probabilities over all species, $s_x = \sum{p_i}$. Because in this model all species are equal, we expect the local species richness to be a linear function of the patch degree. Multi-species analysis of metapopulation models also reveals interesting predictions on other aspects of community organization at various spatial scales such as species-area relationship \parencite{Hanski1997}, and proved to be useful in conservation ecology with predictions of extinctions following habitat destruction \parencite{Nee1994,Rybicki2013}.

%------------------------
\subsection*{Neutral dynamics}

Neutral theory introduces strong competitive interactions by assuming there is a finite number of individuals that could occupy a  patch. There are different ways to simulate this \emph{zero-sum rule} \parencite{Bell2000,Hubbell2001}, but they all result in the same constraint that the increase in abundance of a species could only occur after an equivalent decrease by another species. One important change in the formulation of most neutral models relative the patch dynamics model presented above is therefore that it is individual-based, not population based. We therefore considered in our toy model of neutral dynamics that each local patch is made of $J_x$ individuals. The model tracts the local abundance of all species $N_{ix}$ in each local patch. At each time step an individidual dies with probability $k$. Recruitment only occurs in vacant sites, similarly to a tree by tree replacement process in a closed canopy forest.

The formulation of the recruitment probability is the central piece of all neutral models, making possible the coupling with the metacommunity and neighbouring patches. We adopt a simple formulation in our model based on \parencite{Gravel2006}. The approach is conceptually similar to placing a seed trap in a gap in a forest and picking a seed at random among the one that fall in to determine the identity of the recruited species. The composition of the seed pool in that trap will be a mixture of local dispersal and immigrants from the metacommunity. For simplicity, we consider three spatial scales of dispersal but it would be easy to generalize the approach to a continuous seed dispersal kernel \parencite{Gravel2006}. The parameter $m$ is the probability that the recruit is a migrant from neighbouring patches, $M$ is the probability it comes from a larger (and fixed) metacommunity, and consequently by substraction $1-m-M$ is the probability it comes from local dispersal. The fraction $N_{ix}J_x^{-1}$ is the local relative abundance and $P_ix$ is the relative abundance of species $i$ in the seed pool coming from neigbouring patches $x$. The relative abundance in the neighborhood is weighted by the degree of the connected nodes because some nodes will spread their seeds across a higher number of nodes and thus contribute less to the seed pool. We consider simply $P_{ix} =\frac{\sum d_y^{-1}P_{iy}}{\sum dy^{-1}}$.  We assume for simplicity (and without loss of generality, \parencite{Bell2000}) that the relative abundance in the metacommunity is uniform, ie equal to $S^{-1}$. This immigration prevents the collapse of the metacommunity because otherwise all species except one will face extinction by ecological drift (speciation prevents this phenomenon to occur in \textcite{Hubbell2001}). The local recruitment probability is consequently $Pr_{ix} = MS^{-1} + mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$.

The model is neutral because it assumes that the probabilities of a local recruitment, an immigration and a mortality event are all equal. Demographic stochasticity is the source of variations in abundance but larger disturbances could be simulated as well, as long as they hit all species with the same probability, independently of their density. The fundamental feature of neutral dynamics is therefore the ecological drift, defined as population changes emerging from neutraly stable population dynamics. It can be measured as the variance between replicated time series of community dynamics \parencite{Gravel2011}. \textcite{Hubbell2001} provides a very comprehensive analysis of the model, with specific attention to the effect of the different parameters on the drift (and consequently variance in abundance) and time to extinction. Despite its simplicity, the neutral model is surprisingly rich in the predictions it makes. \textcite{Bell2001} and \textcite{Hubbell2001} analyzed the ability of neutral models to predict species abundance distributions, range-abundance relationship, spatial variation in abundance, species-area relationship, community turnover (beta-diversity) and co-occurrence. Other than the ecological equivalence assumption, one of the most criticized aspect of neutral models is the realism of the speciation process and the required speciation rates to sustain species richness \parencite{Ricklefs2003,Etienne2007}. Recent neutral models with more credible speciation models \parencite{Rosindell2009,Desjardins2012a} revealed the difficulty to maintain diversity in neutral models over macro-evolutionary time scales. These models however also generated interesting novel predictions on endemic species richness \parencite{Rosindell2011,Desjardins2012b}.

%------------------------
\subsection*{Species-sorting and mass effect}

The species-sorting and the mass effect perspectives build on the notion of species-specific responses to a spatially varying environment \parencite{Leibold2004a}. There are various ways to simulate such dynamics and we picked the lottery models, in line with tradition \parencite{Mouquet2002} and for its proximity to the neutral model described above. Competition for space occurs during recruitment after the death of an adult. The recruitment is a lottery against potential candidates. The recruitment probability is however biased by species specific response to local environmental conditions.

The lottery dynamics described above for the neutral model assume there is a very large number of offsprings that are candidate for recruitment but only one will survive and develop to the adult stage. The effect of a differentiation to local environmental conditions could be implemented at this stage.The $J_x$ individuals all experience unique environmental conditions $E_{nx}$ called a microsite $n$. We considered a patch average $\overline{E_x}$, with a within-patch variance $\sigma_x$. The regional average is $\overline{E_R}$ and the regional variance $\sigma_R$ (for simplicity we considered normal distributions, but different distributions will lead to different regional similarity constraints \parencite{Mouquet2003,Tilman2004,Gravel2006}). We consider that a fraction $\lambda_{inx}$ of offsprings reaching the microsite where recruitment occurs will survive. The recruitment probability is therefore biased because only some species will be able to cope with the local environmental conditions. For tractability we define the relative abundance in the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$. The calculation of the relative abundance in the seed rain is the same as the neutral model but the recruitment probability differs because only a fraction of offspring survive. It is formulated as $Pr_{ix} = \frac{\lambda_{inx}Z_{ix}}{\sum \lambda_{jnx}Z_{jx}}$. The function describing the relationship between microsite conditions and survival could take various forms and we used the traditional gaussian curve describing the niche, $\lambda_{inx} =  \exp{-\frac{(E_{nx}-u_i)^2}{2\Pi b_i^2}}$, where $u_i$ is the niche optimum and $b_i$ is niche breadth. Note that the model will converge to a neutral model when the niche breadth tends to infinity (which is in fact how we simulated neutral dynamics in the Supplementary Material to minimize the complexity of the code).

Analyses of similar models with a combination of dispersal and species-sorting shown that predictions are extremely variables and depend on the distribution of environmental conditions, niche optimums and breadth. For instance, a well-studied prediction of neutral models is the species abundance distribution. It was shown that niche models can predict similar distributions \parencite{Tilman2004,Gravel2006}. The main prediction is nonetheless that stable and predictable (meaning which species will coexist) if species are sufficiently dissimilar. Local species richness will first depend on the joint effects of local heterogeneity and niche breadth because coexistence requires a limiting similarity \parencite{Schwilk2005}. Local species richness could be increased by a mass effect when dispersal is consistently supplying individuals coming from more favorable locations (refuges). The limiting similarity required to maintain regional coexistence depends on the amount of dispersal because exchanges among communities homogenizes environmental conditions. This is one of the main result from the species sorting theory and a clever example of local-regional feedbacks: increasing dispersal promotes local coexistence, but on the other hand it diminishes regional coexistence. Only the best average competitors will remain at very high dispersal.  We therefore expect a hump-shaped relationship between dispersal and alpha diversity, with a peak at intermediate dispersal. On the other hand, we expect a monotonic decrease of beta and gamma diversity with dispersal \parencite{Mouquet2003}. This prediction has been validated in some experiments \parencite{Venail2008, Logue2011}.

%========================================================%
\section*{Results}
TP

\subsection*{Alpha diversity}
TP

\subsection*{Beta diversity} 
TP

%========================================================%
\section*{Discussion}
% TP

Contrasting the three perspectives: what do we learn?

New questions

Making the theory operational
	- how to parameterize neutral models
	- making predictions for specific landscapes
	-

Distribution attendue des espèces dans un paysage: utiliser la théorie neutre pour avoir un attendu (différents de la méthode classique de permutation)

- Utiliser les déviations locales pour comprendre le rôle de la sélection

- Approche par réseau: vers une approche qui intègre la variance des mécanismes de coexistence

- Importance de la structure du paysage sur les propriétés émergentes
	- importance relative de species sorting et drift va dépendre de l'organisation du paysage
	- prédiction
	- rôle de la distribution de degrés
-

%========================================================%
\section{Acknowledgements}
DG received financial support from NSERC and Canada Research Chair program. TP is funded by a MELS-FQRNT post-doctoral fellowship and PDP by a NSERC fellowship.
\newpage

%========================================================%
\printbibliography

\newpage
%========================================================%
\section*{Figure legends}

%------------------------
\subsection*{Figure 1}
\textbf{Illustration of the four simulated landscapes}. The color codes represents the local species richness simulated with a neutral model, ranked from the poorest (red) to the richest (blue). Parameters are: $P = 100$, $N = 25$, $m = 0.2$, $M = 0.01$, $k = 0.1$, $J_x = 100$.

%------------------------
\subsection*{Figure 2}
\textbf{Relationship between local species richness and metrics of node centrality}. The upper two panels are simulation results conducted with the random geometric graph illustrated at Fig. 1 and the lower two panels are runs with the random tree graph.

%------------------------
\subsection*{Figure 3}
\textbf{Bray curtis dissimilarity as a function of geographic distance}.

%------------------------
\subsection*{Figure 4}
\textbf{Bray curtis dissimilarity as a function of topological distance}.

\newpage
%========================================================%

\section*{Figures}

%------------------------
\subsection*{Figure 1}

\begin{figure}[ht!]
	\centering\includegraphics[width=0.3\textwidth]{Networks.pdf}
\end{figure}

\newpage

%------------------------
\subsection*{Figure 2}

\begin{figure}[ht!]
	\centering\includegraphics[width=0.75\textwidth]{Centrality.pdf}
\end{figure}

\newpage

%------------------------
\subsection*{Figure 3}

\begin{figure}[ht!]
	\centering\includegraphics[width=0.65\textwidth]{BetaGeoDist.pdf}
\end{figure}

\newpage

%------------------------
\subsection*{Figure 4}

\begin{figure}[ht!]
	\centering\includegraphics[width=0.65\textwidth]{BetaTopoDist.pdf}
\end{figure}

\newpage

\end{document}
%========================================================%

### 2013-05-17 codeR/simulations.R
Moving files around

@@ -186,7 +186,7 @@ plot_alpha(geograph,alphaNeutralGeograph)
 plot_alpha(geotree,alphaNeutralGeotree)
 #plot_alpha(geotree,alphaSSGeotree)
 
-dev.copy2pdf(file = "Figure1.pdf")
+dev.copy2pdf(file = "Networks.pdf")
 
 ###############################
 # Figure 2
@@ -215,7 +215,7 @@ plot(eigGeotree,apply(patchGeotree[[1]],1,sum),pch = 19, xlab = "Eigen centralit
 points(eigGeotree,apply(neutralGeotree[[1]],1,sum),pch = 21, bg = "red")
 points(eigGeotree,apply(ssGeotree[[1]],1,sum),pch = 21, bg = "blue")
 
-dev.copy2pdf(file = "Figure2.pdf")
+dev.copy2pdf(file = "Centrality.pdf")
 
 ###############################
 # Figure 3
@@ -238,7 +238,7 @@ plot(distLattice,betaSSLattice,pch = 19,labels = F,xlab = "",ylab = "",cex = 0.8
 plot(distGeograph,betaSSGeograph,pch = 19,labels = F,xlab = "",ylab = "",cex = 0.8)
 plot(distGeotree,betaSSGeotree,pch = 19,labels = F,xlab = "",ylab = "",cex = 0.8)
 
-dev.copy2pdf(file = "Figure3.pdf")
+dev.copy2pdf(file = "BetaGeoDist.pdf")
 
 ###############################
 # Figure 4
@@ -261,6 +261,6 @@ plot(topoLattice,betaSSLattice,pch = 19,labels = F,xlab = "",ylab = "",cex = 0.8
 plot(topoGeograph,betaSSGeograph,pch = 19,labels = F,xlab = "",ylab = "",cex = 0.8)
 plot(topoGeotree,betaSSGeotree,pch = 19,labels = F,xlab = "",ylab = "",cex = 0.8)
 
-dev.copy2pdf(file = "Figure4.pdf")
+dev.copy2pdf(file = "BetaTopoDist.pdf")
 
 

### 2013-05-17 ms.tex
Line wrapping

@@
 \doublespacing
-
-%========================================================%
-
-\section*{Abstract}
-Background: \\
-Question: \\
-Approach: \\
-Key results: \\
-Conclusion\\
-
-\textbf{Keywords:} metacommunity ecology, neutral theory, species
-sorting, patch dynamics, spatial networks, dispersal\\
-
-\newpage
 
@@
 %========================================================%
-\section{Model description}
-% DG
+\section*{Model description}
 
-General introduction to the section
+In this section we describe three toy models representing different
+perspectives of metacommunity ecology: patch dynamics, neutral dynamics and
+species sorting. While the neutral model is interesting in itself, it is by
+comparison with a model without any interactions (patch dynamics) and with niche
+differentiation (species sorting) that we will be able to fully understand the
+interaction between these processes and landscape structure. Despite neutral,
+competitive interactions in neutral models are very strong because of the
+zero-sum assumption (the community is always at carrying capacity). We will
+first review the fundamental assumptions of each model with their description
+(Table 1 summarizes the parameters and variables that are used), and then
+briefly discuss their main predictions. Simulation results are presented in the
+next section, with the corresponding R code provided in the Supplementary
+Material.
 
-\subsection{Patch dynamics}
-blabla
+\begin{table*}
+	\centering
+	\begin{tabular}{llccc}
+		\hline
+		 & Definition & Patch dynamics & Neutral & Species-sorting \\
+		\hline
+		\textbf{Variables} &     &    &    &   \\
+		$p$ & Occupancy  & X &    &   \\
+		$N$ & Local population size &    & X & X \\
+		$Z$ & Local rel. abund. &  & X & X \\
+		$P$ & Rel. abund. in the neighborhood &  & X & X \\
+		$s$ & Local species richness & X & X & X \\
+		$d$ & Node degree & X & X & X \\
+		$C$ & Prob. of a colonization event & X &  &  \\
+		$I$ & Prob. of a colonization event & X &  &  \\
+		$Pr$ & Recruitment prob. &  & X & X \\
+		$\lambda$ & Survival prob. &  &  & X \\
 
-\subsection{Neutral dynamics}
+		\textbf{Indices} & & & &\\
+		$x,y$ & Node location & X & X & X \\
+		$i,j$ & Species & X & X & X\\
+		$n$ & Microsite & & & X \\
 
-\subsection{Species sorting and mass effect}
-blabla
+		\textbf{Parameters} & & & &\\
+		$S$ & Size of regional species pool & X & X & X \\
+		$c$ & Colonization prob. & X & & \\
+		$e$ & Extinction prob. & X & & \\
+		$J$ & Local carrying capacity & & X & X \\
+		$m$ & immigration prob. from neigh. & & X & X \\
+		$M$ & immigration prob. from metaco. & & X & X \\
+		$k$ & Death prob. & & X & X \\
+		$u$ & Niche optimum & & & X \\
+		$b$ & Niche breadth & & & X \\
+		$E$ & Microsite env. conditions & & & X \\
+		$\overline{E}$ & Local env. average & & & X \\
+		$\sigma$ & Local env. variance & & & X \\
+		$\overline{E_R}$ & Regioal env. average & & & X \\
+		$\sigma_R$ & Regional env. variance & & & X \\
+		\hline
+	\end{tabular}
+	\caption{List of variables, indices and parameters from the three models}
+\end{table*}
 
-%For each model:
-%	- description of the dynamics
-%	- description of the algorithm
-%	- main predictions
-blabla
+%------------------------
+\subsection*{Patch dynamics}
 
-%========================================================%
+The simplest metacommunity model is a $S$ species extension of traditional
+metapopulation models \parencite(Hanski1999). The standard Levins metapopulation
+model \parencite(Levins1969) describes the stochastic colonizations and
+extinctions of population over a homogenous landscape. The basic unit is the
+population. The Levins model tracts the dynamics of occupancy (the fraction of
+the landscape that is occupied) with an ordinary differential equation and
+therefore assumes an infinite landscape. The simulation model we run is more
+realistic as it simulates a finite number $N$ of discrete patches. The rules
+described in the previous section were used to generate connectivity matrices
+along four scenarios (Fig. 1): global dispersal (connected graph), a lattice, a
+random geometric graph and a random tree graph. A patch $x$ shares $d_x$ links
+with neigbhouring patches (its degree). At each time step (the simulation model
+is discrete in time), the probability that a colonist coming from an occupied
+patch $y$ arrives at patch $x$ is $cd_y^{-1}$, where $c$ is the probability a
+colonization event takes place if all connected patches are occupied. The
+expected probability that a colonist arrives to patch $x$ from patch $y$ is then
+$C_{ixy}=cp_{iy}d_y^{-1}$, where $p_{i}y$ is the probability that patch $y$ is
+occupied by species $i$. The probability that an extinction occurs in a given
+patch is $e$. The Levins model is for a single species, but a basic
+metacommunity patch dynamics model could be run by aggregating $S$ independent
+metapopulation models \parencite(Hanski1997). There are no interactions in this
+simple model, which means there is no limit to local species richness and no
+carrying capacity. Competitive, mutualistic and predator-prey interactions have
+been added to this framework (e.g.
+\textcite{Tilman1994,Klausmeier1998,Holt1996}) but we will keep this model
+minimal for the sake of comparison with the neutral model.
+
+Predictions of the patch dynamics metacommunity model are quite straightforward.
+First, a fundamental result of metapopulation ecology is that persistence will
+occur if colonization probability is larger than extinction probability ($c>e$).
+Given that all species are the same, then we should expect the regional
+diversity to be $S$ if this condition is satisfied and $0$ if not. The situation
+is however more complex in spatially explicit landscapes with complex
+connectivity matrices \parencite{Hanski1998}. Spatially explicit dispersal
+usually reduces the occupancy and thereby the likelihood of persistence. The
+second prediction is that, given spatial variation in connectivity, there will
+be spatial variation in occurrence probability. Given the above formulation of a
+colonization event to occur, the probability that an empty location is colonized
+is $I_ix=1-\prod d_x(1-C_{ixy})$. This equation basically tells us that the
+colonization probability will increase asymptotically with the degree of a patch
+(because of the product). It is easy to show from metapopulation theory that the
+occurrence probability in a patch is then $p_ix=I_x(I_ix+e)^{-1}$. The feedback
+between local and regional dynamics arises because all $p_ix$ from the landscape
+are dependent from each other. Simulation are usually conducted to solve the
+model for a large landscape, but numerical solutions are theoretically possible.
+The aggregation across the $S$ species of the regional species pool is obtained
+by taking the summation of occurrence probabilities over all species, $s_x =
+\sum{p_i}$. Because in this model all species are equal, we expect the local
+species richness to be a linear function of the patch degree. Multi-species
+analysis of metapopulation models also reveals interesting predictions on other
+aspects of community organization at various spatial scales such as species-area
+relationship \parencite{Hanski1997}, and proved to be useful in conservation
+ecology with predictions of extinctions following habitat destruction
+\parencite{Nee1994,Rybicki2013}.
+
+%------------------------
+\subsection*{Neutral dynamics}
+
+Neutral theory introduces strong competitive interactions by assuming there is a
+finite number of individuals that could occupy a  patch. There are different
+ways to simulate this \emph{zero-sum rule} \parencite{Bell2000,Hubbell2001}, but
+they all result in the same constraint that the increase in abundance of a
+species could only occur after an equivalent decrease by another species. One
+important change in the formulation of most neutral models relative the patch
+dynamics model presented above is therefore that it is individual-based, not
+population based. We therefore considered in our toy model of neutral dynamics
+that each local patch is made of $J_x$ individuals. The model tracts the local
+abundance of all species $N_{ix}$ in each local patch. At each time step an
+individidual dies with probability $k$. Recruitment only occurs in vacant sites,
+similarly to a tree by tree replacement process in a closed canopy forest.
+
+The formulation of the recruitment probability is the central piece of all
+neutral models, making possible the coupling with the metacommunity and
+neighbouring patches. We adopt a simple formulation in our model based on
+\parencite{Gravel2006}. The approach is conceptually similar to placing a seed
+trap in a gap in a forest and picking a seed at random among the one that fall
+in to determine the identity of the recruited species. The composition of the
+seed pool in that trap will be a mixture of local dispersal and immigrants from
+the metacommunity. For simplicity, we consider three spatial scales of dispersal
+but it would be easy to generalize the approach to a continuous seed dispersal
+kernel \parencite{Gravel2006}. The parameter $m$ is the probability that the
+recruit is a migrant from neighbouring patches, $M$ is the probability it comes
+from a larger (and fixed) metacommunity, and consequently by substraction
+$1-m-M$ is the probability it comes from local dispersal. The fraction
+$N_{ix}J_x^{-1}$ is the local relative abundance and $P_ix$ is the relative
+abundance of species $i$ in the seed pool coming from neigbouring patches $x$.
+The relative abundance in the neighborhood is weighted by the degree of the
+connected nodes because some nodes will spread their seeds across a higher
+number of nodes and thus contribute less to the seed pool. We consider simply
+$P_{ix} =\frac{\sum d_y^{-1}P_{iy}}{\sum dy^{-1}}$.  We assume for simplicity
+(and without loss of generality, \parencite{Bell2000}) that the relative
+abundance in the metacommunity is uniform, ie equal to $S^{-1}$. This
+immigration prevents the collapse of the metacommunity because otherwise all
+species except one will face extinction by ecological drift (speciation prevents
+this phenomenon to occur in \textcite{Hubbell2001}). The local recruitment
+probability is consequently $Pr_{ix} = MS^{-1} + mP_{ix} +
+(1-m-M)N_{ix}J_x^{-1}$.
+
+The model is neutral because it assumes that the probabilities of a local
+recruitment, an immigration and a mortality event are all equal. Demographic
+stochasticity is the source of variations in abundance but larger disturbances
+could be simulated as well, as long as they hit all species with the same
+probability, independently of their density. The fundamental feature of neutral
+dynamics is therefore the ecological drift, defined as population changes
+emerging from neutraly stable population dynamics. It can be measured as the
+variance between replicated time series of community dynamics
+\parencite{Gravel2011}. \textcite{Hubbell2001} provides a very comprehensive
+analysis of the model, with specific attention to the effect of the different
+parameters on the drift (and consequently variance in abundance) and time to
+extinction. Despite its simplicity, the neutral model is surprisingly rich in
+the predictions it makes. \textcite{Bell2001} and \textcite{Hubbell2001}
+analyzed the ability of neutral models to predict species abundance
+distributions, range-abundance relationship, spatial variation in abundance,
+species-area relationship, community turnover (beta-diversity) and
+co-occurrence. Other than the ecological equivalence assumption, one of the most
+criticized aspect of neutral models is the realism of the speciation process and
+the required speciation rates to sustain species richness
+\parencite{Ricklefs2003,Etienne2007}. Recent neutral models with more credible
+speciation models \parencite{Rosindell2009,Desjardins2012a} revealed the
+difficulty to maintain diversity in neutral models over macro-evolutionary time
+scales. These models however also generated interesting novel predictions on
+endemic species richness \parencite{Rosindell2011,Desjardins2012b}.
+
+%------------------------
+\subsection*{Species-sorting and mass effect}
+
+The species-sorting and the mass effect perspectives build on the notion of
+species-specific responses to a spatially varying environment
+\parencite{Leibold2004a}. There are various ways to simulate such dynamics and
+we picked the lottery models, in line with tradition \parencite{Mouquet2002} and
+for its proximity to the neutral model described above. Competition for space
+occurs during recruitment after the death of an adult. The recruitment is a
+lottery against potential candidates. The recruitment probability is however
+biased by species specific response to local environmental conditions.
+
+The lottery dynamics described above for the neutral model assume there is a
+very large number of offsprings that are candidate for recruitment but only one
+will survive and develop to the adult stage. The effect of a differentiation to
+local environmental conditions could be implemented at this stage.The $J_x$
+individuals all experience unique environmental conditions $E_{nx}$ called a
+microsite $n$. We considered a patch average $\overline{E_x}$, with a
+within-patch variance $\sigma_x$. The regional average is $\overline{E_R}$ and
+the regional variance $\sigma_R$ (for simplicity we considered normal
+distributions, but different distributions will lead to different regional
+similarity constraints \parencite{Mouquet2003,Tilman2004,Gravel2006}). We
+consider that a fraction $\lambda_{inx}$ of offsprings reaching the microsite
+where recruitment occurs will survive. The recruitment probability is therefore
+biased because only some species will be able to cope with the local
+environmental conditions. For tractability we define the relative abundance in
+the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$. The
+calculation of the relative abundance in the seed rain is the same as the
+neutral model but the recruitment probability differs because only a fraction of
+offspring survive. It is formulated as $Pr_{ix} =
+\frac{\lambda_{inx}Z_{ix}}{\sum \lambda_{jnx}Z_{jx}}$. The function describing
+the relationship between microsite conditions and survival could take various
+forms and we used the traditional gaussian curve describing the niche,
+$\lambda_{inx} =  \exp{-\frac{(E_{nx}-u_i)^2}{2\Pi b_i^2}}$, where $u_i$ is the
+niche optimum and $b_i$ is niche breadth. Note that the model will converge to a
+neutral model when the niche breadth tends to infinity (which is in fact how we
+simulated neutral dynamics in the Supplementary Material to minimize the
+complexity of the code).
+
+Analyses of similar models with a combination of dispersal and species-sorting
+shown that predictions are extremely variables and depend on the distribution of
+environmental conditions, niche optimums and breadth. For instance, a
+well-studied prediction of neutral models is the species abundance distribution.
+It was shown that niche models can predict similar distributions
+\parencite{Tilman2004,Gravel2006}. The main prediction is nonetheless that
+stable and predictable (meaning which species will coexist) if species are
+sufficiently dissimilar. Local species richness will first depend on the joint
+effects of local heterogeneity and niche breadth because coexistence requires a
+limiting similarity \parencite{Schwilk2005}. Local species richness could be
+increased by a mass effect when dispersal is consistently supplying individuals
+coming from more favorable locations (refuges). The limiting similarity required
+to maintain regional coexistence depends on the amount of dispersal because
+exchanges among communities homogenizes environmental conditions. This is one of
+the main result from the species sorting theory and a clever example of
+local-regional feedbacks: increasing dispersal promotes local coexistence, but
+on the other hand it diminishes regional coexistence. Only the best average
+competitors will remain at very high dispersal.  We therefore expect a
+hump-shaped relationship between dispersal and alpha diversity, with a peak at
+intermediate dispersal. On the other hand, we expect a monotonic decrease of
+beta and gamma diversity with dispersal \parencite{Mouquet2003}. This prediction
+has been validated in some experiments \parencite{Venail2008, Logue2011}.
+
 \section{Results}
-% TP
 
-\subsection{Alpha diversity}
+In fig. 1, we illustrate the outcome of simulations using the neutral model on
+the four different network shapes considered in this article. All of these
+networks have the same number of nodes (e.g. spatial sampling sites), but both
+different number of edges (e.g. dispersal routes between sampling sites) and
+patterns of connectivity between nodes. We ask how these differences in
+connectivity will shape the emerging properties of the community under the
+scenarios represented by each meta-community model.
 
-\subsection{Beta diversity}
+In fig. 2,
 
@@
 neutral dynamics. Emerging properties of a community (i.e. the number of
-species, but also their spatial distribution and co-occurence at different local
+species, but also their spatial distribution and co-occurence in different local
 patches), are affected by the dispersal structure.
@@
 \section{Acknowledgements}
-DG received financial support from NSERC and Canada Research Chair program. TP is funded by a MELS-FQRNT post-doctoral fellowship and PDP by a NSERC fellowship.
+DG received financial support from NSERC and Canada Research Chair program. TP
+is funded by a MELS-FQRNT post-doctoral fellowship and PDP by a NSERC
+fellowship.
 \newpage
@@
 \printbibliography
+
+\newpage
+%========================================================%
+\section*{Figure legends}
+
+%------------------------
+\subsection*{Figure 1}
+\textbf{Illustration of the four simulated landscapes}. The color code
+represents the local species richness simulated with a neutral model, ranked
+from the poorest (red) to the richest (blue). Parameters are: $P = 100$, $N =
+25$, $m = 0.2$, $M = 0.01$, $k = 0.1$, $J_x = 100$.
+
+%------------------------
+\subsection*{Figure 2}
+\textbf{Relationship between local species richness and metrics of node
+centrality}. The upper two panels are simulation results conducted with the
+random geometric graph illustrated at Fig. 1 and the lower two panels are runs
+with the random tree graph.
+
+%------------------------
+\subsection*{Figure 3}
+\textbf{Bray curtis dissimilarity as a function of geographic distance}.
+
+%------------------------
+\subsection*{Figure 4}
+\textbf{Bray curtis dissimilarity as a function of topological distance}.
+
+\newpage
+%========================================================%
+
+\section*{Figures}
+
+%------------------------
+\subsection*{Figure 1}
+
+\begin{figure}[ht!]
+	\centering\includegraphics[width=0.3\textwidth]{Networks.pdf}
+\end{figure}
+
+\newpage
+
+%------------------------
+\subsection*{Figure 2}
+
+\begin{figure}[ht!]
+	\centering\includegraphics[width=0.75\textwidth]{Centrality.pdf}
+\end{figure}
+
+\newpage
+
+%------------------------
+\subsection*{Figure 3}
+
+\begin{figure}[ht!]
+	\centering\includegraphics[width=0.65\textwidth]{BetaGeoDist.pdf}
+\end{figure}
+
+\newpage
+
+%------------------------
+\subsection*{Figure 4}
+
+\begin{figure}[ht!]
+	\centering\includegraphics[width=0.65\textwidth]{BetaTopoDist.pdf}
+\end{figure}
+
 \newpage

### 2013-05-17 ms.tex
Results fig 2

@@
 
-In fig. 2,
+In fig. 2, we present the richness of each node of the network, as a function of
+the centrality of the node, under different assumptions of meta-community
+dynamics and network structure. It appears that both in the goeographical and
+tree graph, the path dynamics model allows more diversity to persist overall.
+However, the local diversity increases with the node centrality, whether it is
+measured by the degree or the eigencentrality. All situations examined give a
+remarkably congruent signal. The patch dynamics model, \emph{i.e.} when there
+are no interactions between species, allows more $\alpha$ diversity. The neutral
+model allowed the maintenance of slightly more diversity than the
+species-sorting model. In all cases however, it is clear that the richness
+maintained tend to increase with the node centrality.
 

### 2013-05-18 ms.tex
Write text on spatial networks, add references

--- 2013-05-16 ms.tex
@@
-\documentclass[12pt]{article}
+\documentclass[letterpaper,twocolumn,showkeys]{revtex4-1}
 \usepackage[utf8]{inputenc}
-\usepackage{color,lineno,setspace,graphicx,multirow,kpfonts}
-\usepackage[top=2.4cm,left=2.4cm,top=2.4cm,bottom=2.4cm,includefoot]{geometry}
-\usepackage[style=bes]{biblatex}
-\bibliography{./library}
+\usepackage{color,dcolumn,graphicx,hyperref}
 
@@
 
-\linenumbers 
-\modulolinenumbers[1]
+\title{TO FIND}
 
-\textbf{Title:}   Using neutral theory to reveal the contribution of dispersal to community assembly\\
+\author{Dominique Gravel}
+\author{Timothée Poisot}
+\author{Philippe Desjardins-Proulx}
 
-\textbf{Authors:}  Dominique Gravel$^{1,2,*}$, Timoth\'ee Poisot$^{1,2}$, Philippe Desjardins-Proulx$^{1,2}$\\
-
-1: Canada Research Chair on Terrestrial Ecosystems. D\'epartement de biologie, chimie et g\'eographique, Universit\'e du Qu\'ebec \`a Rimouski, 300 All\'ee des Ursulines, Qu\'ebec, Canada. G5L 3A1.\\
-
-2: Qu\'ebec Centre for Biodiversity Sciences, Stewart Biological Sciences Building, 1205 Dr.~Penfield Avenue, Montr\'eal (QC), H3A 1B1, Canada\\
-
-\textbf{Words in the abstract:}      \\
-\textbf{Words in the main text:}    \\
-\textbf{Words in the legends:}    \\
-\textbf{References:}             \\
-\textbf{Table:}                    \\
-
-\newpage
-\doublespacing
-
-%========================================================%
-
-\section*{Abstract}
+\begin{abstract}
 Background: \\
@@
 
-\textbf{Keywords:} metacommunity ecology, neutral theory, species
-sorting, patch dynamics, spatial networks, dispersal\\
+\end{abstract}
 
-\newpage
+\keywords{metacommunity ecology, neutral theory, species
+sorting, patch dynamics, spatial networks, dispersal}
+
+\maketitle
 
@@
 \section{Introduction}
+%DG
 
-Community assembly is concerned by patterns and processes occurring at various
-spatial scales \parencite{Levin1992}. Until the development of metacommunity
-ecology, studies on community assembly were often restricted to local
-populations, with a strong focus on pairwise interactions (e.g.
-\textcite{MacArthur1972, May1973,Pimm1982,DeAngelis1992}). The emphasis on local
-communities has been vigorously criticized by \textcite{Ricklefs2008}, who has
-long recognized that local dynamics and community structure are strongly
-contingent on processes occurring at much larger spatial scales
-\parencite{Ricklefs1987}. This perspective is particularly relevant to
-limnology, where exchanges of organisms and nutrients affect community and
-ecosystem properties from the local (e.g. vertical mixing
-\parencite{Ryabov2011}) to the regional (e.g. connection of lakes
-\parencite{Leibold2004b} scales. It emphasizes the importance of dispersal
-relative to pairwise interactions in the organization of ecological communities.
-The metacommunity concept has been proposed by \textcite{Leibold2004a} as a
-novel approach to link different spatial scales in ecology. It builds on
-feedbacks between local scale processes, such as competitive interactions and
-local adaptation, and regional scale processes such as dispersal, gene flow and
-speciation. Ecologists are now required to move toward a predictive ecology,
-integrating elements of theoretical ecology \parencite{Thuiller2013}, and the
-metacommunity perspective appears naturally as the appropriate conceptual
-framework to develop the new modeling techniques required to fill this
-challenge. The development of neutral theory has been quite provocative in that
-respect, as one could see it as a step back in time. Neutral theory makes the
-provocative assumption that species are ecologically equivalent and thereby any
-variation in the environment has no impact on demography
-\parencite{Bell2000,Hubbell2001}. Only demographic stochasticity and dispersal
-drive the structure of neutral ecological communities. It therefore appears
-that, on first sight, neutral theory is useless.  We will develop in this paper
-the argument that neutral theory could be a useful tool to understand the impact
-of dispersal on community organization in landscapes of various complexities.
+- Contribution of neutral theory to ecology: the recognition of
+the importance of dispersal in structuring communities. Link with limnology.
 
-Neutral theory sparked an historical debate in community ecology that is still
-lasting after more than a decade \parencite{Chave2004, Etiennee2011,
-Rosindell2012,Clark2012}. It was stimulated by the impressive ability of neutral
-models to fit several well studied empirical observations such as species
-abundance distributions and distance-decay relationships. A remarkable strenght
-of neutral theory is to provide a \emph{"formal general theory of abundance
-and diversity that will account, in a simple and economical fashion, for the
-many patterns that ecologists have documented"} \parencite{Bell2001}. Even if
-new studies rejecting neutral theory are consistently published (e.g.
-\textcite{Ricklefs2012}), there is now almost a consensus that neutral theory is
-a well-developed null hypothesis for niche theory and could even be used as an
-adequate approximation of ecological dynamics in some situations.
-\textcite{Bell2001} nicely envisionned two perspectives to neutral theory that
-are still standing today. Under the weak perspective, neutral theory provides a
-set of realistic predictions of community organization despite false
-assumptions. Even if being fundamentally wrong, neutral theory is still useful
-when used as a null hypothesis \parencite{Gotelli2006}. It is considered as an
-improvement over traditional null hypotheses based on randomization
-\parencite{Gotelli2000} because it readily integrates dispersal. The strong
-version on the other hand posits that neutral theory is a satisfying
-approximation to community dynamics and an appropriate theory to explain the
-distribution of biodiversity. It implies that the right mechanisms have been
-identified and that the consistently observed differences among species do not
-impact community organization.
+- Gave roots and motivation for the metacommunity concepts.
+Present the different perspectives.
 
-Neutral theory has also been proposed as a useful tool to understand and predict
-some aspects of community dynamics. It links to an old philosophical debate
-between realism and instrumentalism \parencite{Wennekes2011}. Because every
-ecological model is a simplification of reality, any scientist has to
-subjectively decide the level of details he puts in, leaving out some elements
-judged unimportant. The realism perspective requires that all assumptions of
-theory to be true, while the utility of the theory is more important to
-instrumentalism. The utilitarian value of a theory could either be for
-understanding or prediction (another old philosophical debate, see
-\textcite{Schmueli2010}). Obviously neutral theory could only be instrumental.
-The question then is if such a 'general, large-scale, but vague' theory
-\parencite{Wennekes2011} is a satisfying approximation.
+- Objective of the paper 
+	- Already several reviews of neutral theory in the literature
+- We will use this opportunity to explore in more details the
+contribution of neutral theory to our understanding of
+metacommunities
 
-The instrumentalist view of neutral theory raises the question of why it should
-be a satisfying approximation despite knowing the pieces are wrong? We see two
-potential answers to this question. A first answer might be that stochasticity
-of various origins can blur the deterministic differences among species and
-promote the ecological drift \parencite{Gravel2011}. Much has been said the
-existence of demographic stochasticity, some ecologists even arguing that
-neutral models impede progress in community ecology by hidding niche differences
-\parencite{Clark2012}, and we therefore will keep this discussion for other
-papers. The second answer is that dispersal and historical contingencies might
-have a much more profound impact on species distribution \parencite{Bahn2007}
-and ecological dynamics. The debate over the equivalence assumption and
-demographic stochasticity has perhaps overlook the recognition of how much
-dispersal influence community assembly.
+- We will pay a particular attention to recent developments at
+the crossroad of network theory
 
-In this paper, we argue that neutral theory is useful both to understand and
-predict the impact of dispersal on community organization. Even
-for purely theoretical analyses, we need a benchmark without niche differences to
-reveal the role of dispersal in structuring communities, and understand how it
-interacts with niche differentiation. We will explore recent applications of
-neutral theory, at the crossroad of network theory, to better represent the
-impact of landscape structure on biodiversitity distribution. This analysis will
-prove particularly relevant to limnology, where most riverine and lacustre
-habitats are characterized by a their discrete nature and spatially complex
-arrangements. We will also reveal the relative contribution of ecological
-interaction and niche differentiation by contrasting predictions of a neutral
-model to other metacommunity perspectives.
+- Move from a perspective where dispersal is either global or
+spatially explicity (e.g. over a lattice), and spatial constant,
+to a perspective focusing on the variance of dispersal.
 
-Our main objective in this paper is to use neutral theory to stress the
-importance of landscape network structure on the distribution of diversity. We
-refer to the landscape organization as a "spatial contingency"
-\parencite{Peres-Neto2013} that could potentially affect the coexistence
-mechanisms at play. We will therefore move from a perspective where dispersal is
-either global or spatially explicity (e.g. over a lattice), and spatial
-constant, to a perspective focusing on the variance of dispersal. A second
-generation of neutral models (e.g. \textcite{Economo2008,
-Economo2011,Desjardins2012a,Desjardins2012b}, and even experiments
-\parencite{Altermat2012}, recently introduced more realistic lanscapes and found
-surprising contributions of spatial contingencies. We will start with a short
-review of the main approaches to describe spatial networks and the studies
-investigating them. Then we will describe three simple toy models of
-metacommunity dynamics, taking this opportunity to review their assumptions and
-main predictions. We provide as Supplementary Material the R scripts for the toy
-models and all simulations conducted for this paper. We then conduct simple
-simulations of these models to reveal the impact of spatial network structure on
-diversity distribution. We conclude with a discussion on the operationally of
-the framework.
+- More specifically, the objective of the paper is to use
+neutral theory to better reveal the importance of landscape
+network structure on the distribution of diversity.
+	- Referred as a "spatial contingency" 
+	- Relevance to limnology
+
+- Structure:
+- Start with a review of the main approaches to describe spatial
+networks and studies investigating them
+- Then we describe three simple toy models of metacommunity
+dynamics, taking this opportunity to review their assumptions
+and functioning.
+- R scripts for the toy models and all simulations conducted for
+this paper are provided as supplementary materiel.
+- We then conduct simple simulations of these models to reveal
+the impact of spatial network structure on diversity
+distribution.
+- We conclude with a discussion on the operationally of the
+framework.
 
@@
 
-Entités discrètes --> réseaux
+A network is a discrete mathematical object made of two sets: a set of 
+vertices (or nodes) $V$ and a set of edges $E$ connecting the vertices
+\cite{new10}. The term ``graph'' is often preferred in computer science and 
+mathematics \cite{gro06}, with graph algorithms being an important and active 
+area of research \cite{sed01}. A network is a combinatorial object: it is 
+used to study how discrete entities are connected and how they combine 
+together to create complex structures. They are used to study molecules, food 
+webs, social networks, or even the relationship between variables in 
+statistics \cite{wri21,new10}. We are especially interested in spatial 
+networks, a special kind of network mixing the combinatorial properties of 
+networks with a topological space \cite{kob94}. Thus, the vertices in a 
+spatial graph are embedded in some other space, most often the two or three-
+dimensional Euclidean space. This object brings a rich representation to 
+spatial ecology and is particularly suited for systems of lakes and rivers, which can 
+easily be represented by vertices and edges. There are two notions of 
+distance in spatial networks. Euclidean distance represents the geographical 
+distance between the vertices $(i, j)$, i.e.: $\sqrt{(x_i - x_j)^2 + (y_i - y_
+j)^2}$. Geodesic distance is the distance in the graph space, i.e.: the 
+length of the shortest path \cite{dij59}. For example, two lakes could be 
+very close on a map (short Euclidean distance) but if they are not directly 
+linked by a river the geodesic distance could be great.
+
+For simulations, spatial networks can easily be generated with the random 
+geometric graph algorithm \cite{sed01}. In this algorithm, all vertices are 
+assigned to a position in some two-dimensional space, most often the unit 
+square. Then, all pairs of vertices within some threshold Euclidean distance $r$
+are connected with an edge. The resulting networks have the desirable 
+property of locality: if a vertex $A$ is connected to two vertices $B$ and $C$.
+then $B$ and $C$ are more likely to be connected than two random vertices. 
+Random geometric networks have been extensively studied \cite{app97a,app97
+b,app02a,app02b,pen03} and we provide a R function to generate them. We also 
+provide the code for another structure that we call the geometric tree of 
+doom. It builds a tree from the shortest path tree \cite{dij59} of a vertex in 
+a random geometric network. This random geometric tree of doom looks like a
+lake connected by rivers to a series of smaller lakes.
+
+Spatial graphs are increasingly popular in spatial ecology and conservation
+biology, where patterns of connections can be used to study and influence the
+flow of organisms \cite{min07,fal07,min08,gar08,urb09,dal10}. In the neutral 
+theory, networks were pioneered by Economo and Keitt \cite{eco08, eco10}. 
+They used networks to study how different spatial structures influenced 
+diversity. They were also used to study how the spatial structure influenced 
+nonsympatric speciation \cite{des12,des12b}.
 
@@
 % We might want to distinguish network level versus node-level metrics.
+Concepts
+  Path                    A sequence of edges forming a sequence of vertices.
+  Connection              Two vertices are connected iff there is a path between them.
+  Geodesic distance       Length of the shortest path between two vertices.
+Network-level metrics
+  Order                   Total number of vertices.
+  Size                    Total number of edges.
+  Connectivity            A measure of robustness: the minimum number of elements to remove
+                          to isolate the vertices.
+  Components              The number of connected subsets.
+Vertex-level metrics
+  Degree                  The number of edges of a vertex.
+  Closeness centrality    Average geodesic distance between a vertex and all other vertex.
+  Eigenvector centrality  A measure of centrality based on the concept that connection to
+                          highly connected vertices are more important.
+  Betweenness centrality  The number of shortest paths from all vertices to all others that
+                          pass through that vertex.
 
@@
 Making the theory operational
-	- how to parameterize neutral models
-	- making predictions for specific landscapes
-	-
 
@@
 
-- Utiliser les déviations locales pour comprendre le rôle de la sélection
-
-- Approche par réseau: vers une approche qui intègre la variance des mécanismes de coexistence
-
-Our results point to an important, although often overlooked, consequence of
-neutral dynamics. Emerging properties of a community (i.e. the number of
-species, but also their spatial distribution and co-occurence at different local
-patches), are affected by the dispersal structure.
+- Déviation locale des prédictions
 
 - Importance de la structure du paysage sur les propriétés émergentes
-	- importance relative de species sorting et drift va dépendre de l'organisation du paysage
-	- prédiction
-	- rôle de la distribution de degrés
+
 -
 
-%========================================================%
-\section{Acknowledgements}
-DG received financial support from NSERC and Canada Research Chair program. TP is funded by a MELS-FQRNT post-doctoral fellowship and PDP by a NSERC fellowship.
-\newpage
-
-%========================================================%
-\printbibliography
-\newpage
+\bibliography{../refs}
+\bibliographystyle{plain}
 
 \end{document}
-%========================================================%

### 2013-05-22 codeR/simulations.R
Clean

@@ -1,225 +1,0 @@
-########################################################################
-# Source for the figures 1-3 in the paper
-# Gravel, Poisot and Desjardins. 2013. XXX. J. Limnology. 
-#
-# By: Dominique Gravel (dominique_gravel@uqar.ca) 
-# April 2013
-#
-########################################################################
-source("spatial_graphs.R")
-source("patch_model.R")
-source("lottery_model.R")
-library(vegan)
-
-###############################
-# Useful functions
-alpha_fn = function(pres) {
-	s = apply(pres,1,sum)
-	vec.col = numeric(length(s))
-	RK = rank(s)
-	for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
-	return(vec.col)	
-}
-
-plot_alpha = function(spatial_graph, vec.col) {
-  	# Plots a spatial graph 
-  	#
-  	# Args:
-  	#   spatia_graph: the output of one of the spatial graphs
-  	#
-	par(mar=c(1,1,1,1))
-	XY = spatial_graph[[1]]
-	adjMat = spatial_graph[[2]]
-
-	plot(XY[,1],XY[,2],xlab = "", ylab = "",cex = 1.5,labels = F)
-	adjVec = stack(as.data.frame(adjMat))[,1]
-	XX = expand.grid(XY[,1],XY[,1])
-	YY = expand.grid(XY[,2],XY[,2])
-	XX = subset(XX,adjVec==1)
-	YY = subset(YY,adjVec==1)
-	arrows(x0 = XX[,1],x1=XX[,2],y0 = YY[,1], y1 = YY[,2], length = 0,lwd = 0.2, col = "darkgrey")
-	points(XY[,1],XY[,2],pch=21,bg=vec.col,cex = 1.5)
-}
-
-###############################
-# General parameters
-S = 100
-n = 25
-r = 0.3
-
-###############################
-# Draw the four landscapes
-connected = connected_fn(n)
-lattice = lattice_fn(n)
-geograph = geograph_fn(n,r)
-geotree = geotree_fn(n,r)
-
-# Geographic distances among plots
-distConnected = as.matrix(dist(connected[[1]], upper = T, diag = T))
[…]

### 2013-05-22 codeR/spatial_graphs.R

@@ -1,277 +1,0 @@
-########################################################################
-# Functions to generate networks ("graphs") for spatial ecology.
-#
-# Stores the graph in a adjacency matrix
-#
-# Functions to generate graphs:
-# - geograph_fn: returns a random geometric graph.
-# - geotree_fn: returns a random geometric tree (of doom!).
-# - lattice_fn: returns a lattice type of spatial graph.
-# - connected_fn: returns a fully connected graph.
-#
-# Useful functions:
-# - plot_spatial: custom function to plot spatial graphs.
-# - deg_cen: computes degree centrality.
-# - cls_cen: computes closeness centrality.
-# - eig_cen: computers eigen-centrality.
-#
-# Associated functions:
-# - connected: tests if the graph is fully connected.
-# - test_con: recursive function used by 'connected'.
-# - spm: Generates the shortest path distance matrix.
-# - spd: Generates the shortest path distances starting with a given source.
-# - spt: Generates the shortest path tree as an adjacency matrix.
-#
-# Examples:
-# > plot_spatial(geograph_fn(n = 25, r = 0.3))
-# > plot_spatial(geotree_fn(n = 25, r = 0.3))
-# > plot_spatial(connected_fn(n = 25))
-# > plot_spatial(lattice_fn(n = 25))
-#
-# By: Dominique Gravel (dominique_gravel@uqar.ca)
-#     Philippe Desjardins-Proulx (philippe.d.proulx@gmail.com)
-#
-# May 2013
-########################################################################
-
-library('igraph')
-
-geograph_fn = function(n = 64, r = 0.32) {
-  # Generates a random geometric graph.
-  #
-  # Args:
-  #   n: Number of vertices.
-  #   r: Threshold distance to connect vertices.
-  #
-  # Returns:
-  #   A list with the xy coordinates and the adjacency matrix.
-  # Note: the algorithm tests if the graph is connected
-  repeat {
-    xy = cbind(runif(n), runif(n))
-    distMat = as.matrix(dist(xy, method = 'euclidean', upper = T, diag = T))
-    adjMat = matrix(0, nr = n, nc = n)
-    adjMat[distMat < r] = 1
-    diag(adjMat) = 0
-    if (connected(adjMat)) {
[…]

### 2013-05-22 examples.R

@@ -1,88 +1,0 @@
-########################################################################
-# Examples of dynamics on spatial graphs with different metacommunity perspectives
-#
-# By: Dominique Gravel (dominique_gravel@uqar.ca) 
-# April 2013
-#
-########################################################################
-###############################
-# General parameters
-S = 100
-n = 25
-r = 0.3
-nsteps = 100
-
-########################################
-########################################
-# Patch dynamics example
-source("spatial_graphs.R")
-source("patch_model.R")
-c = 0.2
-e = 0.1
-spatial_graph = geograph(n, r)
-results = patch_model(c, e, S, spatial_graph, nsteps)
-
-# Figure 1: time series
-x11(height = 5.5, width = 6)
-plot(c(1:nsteps),apply(results[[2]][,2:(S+1)],1,sum),xlab = "Time", ylab = "Average local species richness",cex.lab = 1.5, cex.axis = 1.25, type = "l", ylim = c(0,S))
-
-# Figure 2: local species richness
-s = apply(results[[1]],1,sum)
-vec.col = numeric(length(s))
-RK = rank(s)
-for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
-plot_spatial(spatial_graph, vec.col)
-
-########################################
-########################################
-# Neutral example
-source("spatial_graphs.R")
-source("lottery_model.R")
-# Default parameters
-m = 0.2
-M = 0.01
-d = 0.1
-J = 100
-sdN = Inf
-sdE = 5
-spatial_graph = geograph_fn(n,r)
-results = lottery_model(m,M,d,S,J,sdE,sdN,spatial_graph,nsteps)
-
-# Figure 1: time series
-x11(height = 5.5, width = 6)
-plot(c(1:nsteps),apply(results[[2]][,2:(S+1)],1,sum),xlab = "Time", ylab = "Average local species richness",cex.lab = 1.5, cex.axis = 1.25, type = "l", ylim = c(0,S))
-
-# Figure 2: local species richness
-s = apply(results[[1]],1,sum)
-vec.col = numeric(length(s))
-RK = rank(s)
-for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
-plot_spatial(spatial_graph, vec.col)
-
-
-########################################
[…]

### 2013-05-22 simulations.R

@@ -1,266 +1,0 @@
-########################################################################
-# Source for the figures 1-4 in the paper
-# Gravel, Poisot and Desjardins. 2013. XXX. J. Limnology. 
-#
-# By: Dominique Gravel (dominique_gravel@uqar.ca) 
-# April 2013
-#
-########################################################################
-source("spatial_graphs.R")
-source("patch_model.R")
-source("lottery_model.R")
-library(vegan)
-
-###############################
-# Useful functions
-alpha_fn = function(pres) {
-	s = apply(pres,1,sum)
-	vec.col = numeric(length(s))
-	RK = rank(s)
-	for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
-	return(vec.col)	
-}
-
-plot_alpha = function(spatial_graph, vec.col) {
-  	# Plots a spatial graph 
-  	#
-  	# Args:
-  	#   spatia_graph: the output of one of the spatial graphs
-  	#
-	par(mar=c(1,1,1,1))
-	XY = spatial_graph[[1]]
-	adjMat = spatial_graph[[2]]
-
-	plot(XY[,1],XY[,2],xlab = "", ylab = "",cex = 1.5,labels = F)
-	adjVec = stack(as.data.frame(adjMat))[,1]
-	XX = expand.grid(XY[,1],XY[,1])
-	YY = expand.grid(XY[,2],XY[,2])
-	XX = subset(XX,adjVec==1)
-	YY = subset(YY,adjVec==1)
-	arrows(x0 = XX[,1],x1=XX[,2],y0 = YY[,1], y1 = YY[,2], length = 0,lwd = 0.2, col = "darkgrey")
-	points(XY[,1],XY[,2],pch=21,bg=vec.col,cex = 1.5)
-}
-
-###############################
-# General parameters
-S = 100
-n = 25
-r = 0.3
-
-###############################
-# Draw the four landscapes
-connected = connected_fn(n)
-lattice = lattice_fn(n)
-geograph = geograph_fn(n,r)
-geotree = geotree_fn(n,r)
-
-# Geographic distances among plots
-distConnected = as.matrix(dist(connected[[1]], upper = T, diag = T))
[…]

### 2013-05-22 spatial_graphs.R

@@ -1,277 +1,0 @@
-########################################################################
-# Functions to generate networks ("graphs") for spatial ecology.
-#
-# Stores the graph in a adjacency matrix
-#
-# Functions to generate graphs:
-# - geograph_fn: returns a random geometric graph.
-# - geotree_fn: returns a random geometric tree (of doom!).
-# - lattice_fn: returns a lattice type of spatial graph.
-# - connected_fn: returns a fully connected graph.
-#
-# Useful functions:
-# - plot_spatial: custom function to plot spatial graphs.
-# - deg_cen: computes degree centrality.
-# - cls_cen: computes closeness centrality.
-# - eig_cen: computers eigen-centrality.
-#
-# Associated functions:
-# - connected: tests if the graph is fully connected.
-# - test_con: recursive function used by 'connected'.
-# - spm: Generates the shortest path distance matrix.
-# - spd: Generates the shortest path distances starting with a given source.
-# - spt: Generates the shortest path tree as an adjacency matrix.
-#
-# Examples:
-# > plot_spatial(geograph_fn(n = 25, r = 0.3))
-# > plot_spatial(geotree_fn(n = 25, r = 0.3))
-# > plot_spatial(connected_fn(n = 25))
-# > plot_spatial(lattice_fn(n = 25))
-#
-# By: Dominique Gravel (dominique_gravel@uqar.ca)
-#     Philippe Desjardins-Proulx (philippe.d.proulx@gmail.com)
-#
-# May 2013
-########################################################################
-
-library(igraph)
-
-geograph_fn = function(n = 64, r = 0.32) {
-  # Generates a random geometric graph.
-  #
-  # Args:
-  #   n: Number of vertices.
-  #   r: Threshold distance to connect vertices.
-  #
-  # Returns:
-  #   A list with the xy coordinates and the adjacency matrix.
-  # Note: the algorithm tests if the graph is connected
-  repeat {
-    xy = cbind(runif(n), runif(n))
-    distMat = as.matrix(dist(xy, method = 'euclidean', upper = T, diag = T))
-    adjMat = matrix(0, nr = n, nc = n)
-    adjMat[distMat < r] = 1
-    diag(adjMat) = 0
-    if (isconnected(adjMat)) {
[…]

### 2013-05-22 ms.tex

--- 2013-05-17 ms.tex
@@
 
+Finally, fig. 3 and fig. 4 present, respectively, the between patch $\beta$
+diversity as a function of the shape of the network, under three dynamic models.
+In figure 3, the distance is expressed as the euclidean (geographic) distance
+between two patches. Although this neglects how dispersal connects the different
+patches, there is already a clear signal of $\beta$ diversity. In both the
+neutral and patch dynamics model, local communities become increasingly
+dissimilar when the distance between them increases. In other words, two
+communities which are close to each other will share a large proportion of their
+species, whereas two communities which are afar will share a small proportion.
+In the species-sorting model, the relationship between distance and
+dissimilarity is similar, Nonetheless, it forms an enveloppe of points (with
+most points lying in the upper-left part of the graph). While two distant
+communities will be dissimilar, there is no telling how dissimilar two close
+communities will be.
+
+To a vast extent, these relationships are preseved when looking at the
+topological distances, i.e. along how many edges should one travel to connect
+two patchs. Interestingly enough, the distance-dissimilarity relationship for
+the neutral model is markedly hump-shaped, with sites being at a medium distance
+having the maximal dissimilarity.
+
 %========================================================%

### 2013-05-22 ms.tex
Phil's section on spatial graphs

@@
 \section{Network representation of landscapes}
-% PDP
 
-Entités discrètes --> réseaux
+A network is a discrete mathematical object made of two sets: a set of
+vertices (or nodes) $V$ and a set of edges $E$ connecting the vertices
+\cite{new10}. The term ``graph'' is often preferred in computer science and
+mathematics \cite{gro06}, with graph algorithms being an important and active
+area of research \cite{sed01}. A network is a combinatorial object: it is
+used to study how discrete entities are connected and how they combine
+together to create complex structures. They are used to study molecules, food
+webs, social networks, or even the relationship between variables in
+statistics \cite{wri21,new10}. We are especially interested in spatial
+networks, a special kind of network mixing the combinatorial properties of
+networks with a topological space \cite{kob94}. Thus, the vertices in a
+spatial graph are embedded in some other space, most often the two or three-
+dimensional Euclidean space. This object brings a rich representation to
+spatial ecology and is particularly suited for systems of lakes and rivers, which can
+easily be represented by vertices and edges. There are two notions of
+distance in spatial networks. Euclidean distance represents the geographical
+distance between the vertices $(i, j)$, i.e.: $\sqrt{(x_i - x_j)^2 + (y_i - y_
+j)^2}$. Geodesic distance is the distance in the graph space, i.e.: the
+length of the shortest path \cite{dij59}. For example, two lakes could be
+very close on a map (short Euclidean distance) but if they are not directly
+linked by a river the geodesic distance could be great.
+
+For simulations, spatial networks can easily be generated with the random
+geometric graph algorithm \cite{sed01}. In this algorithm, all vertices are
+assigned to a position in some two-dimensional space, most often the unit
+square. Then, all pairs of vertices within some threshold Euclidean distance $r$
+are connected with an edge. The resulting networks have the desirable
+property of locality: if a vertex $A$ is connected to two vertices $B$ and $C$.
+then $B$ and $C$ are more likely to be connected than two random vertices.
+Random geometric networks have been extensively studied \cite{app97a,app97
+b,app02a,app02b,pen03} and we provide a R function to generate them. We also
+provide the code for another structure that we call the geometric tree of
+doom. It builds a tree from the shortest path tree \cite{dij59} of a vertex in
+a random geometric network. This random geometric tree of doom looks like a
+lake connected by rivers to a series of smaller lakes.
+
+Spatial graphs are increasingly popular in spatial ecology and conservation
+biology, where patterns of connections can be used to study and influence the
+flow of organisms \cite{min07,fal07,min08,gar08,urb09,dal10}. In the neutral
+theory, networks were pioneered by Economo and Keitt \cite{eco08, eco10}.
+They used networks to study how different spatial structures influenced
+diversity. They were also used to study how the spatial structure influenced
+nonsympatric speciation \cite{des12,des12b}.
 
@@
 % We might want to distinguish network level versus node-level metrics.
+Concepts
+  Path A sequence of edges forming a sequence of vertices.
+  Connection Two vertices are connected iff there is a path between them.
+  Geodesic distance Length of the shortest path between two vertices.
+Network-level metrics
+  Order Total number of vertices.
+  Size Total number of edges.
+  Connectivity A measure of robustness: the minimum number of elements to remove
+                          to isolate the vertices.
+  Components The number of connected subsets.
+Vertex-level metrics
+  Degree The number of edges of a vertex.
+  Closeness centrality Average geodesic distance between a vertex and all other vertex.
+  Eigenvector centrality A measure of centrality based on the concept that connection to
+                          highly connected vertices are more important.
+  Betweenness centrality The number of shortest paths from all vertices to all others that
+                          pass through that vertex.
 
@@
 \section{Discussion}
-% TP
 
@@
 
-New questions
+Network approach brings spatial variation in coexistence mechanisms; emergent properties
+	- role of the degree distribution
 
@@
 	- making predictions for specific landscapes
-	-
 
-Distribution attendue des espèces dans un paysage: utiliser la théorie neutre pour avoir un attendu (différents de la méthode classique de permutation)
+Working with more realistic landscapes
 
-- Utiliser les déviations locales pour comprendre le rôle de la sélection
-
-- Approche par réseau: vers une approche qui intègre la variance des mécanismes de coexistence
-
-Our results point to an important, although often overlooked, consequence of
-neutral dynamics. Emerging properties of a community (i.e. the number of
-species, but also their spatial distribution and co-occurence in different local
-patches), are affected by the dispersal structure.
-
-- Importance de la structure du paysage sur les propriétés émergentes
-	- importance relative de species sorting et drift va dépendre de l'organisation du paysage
-	- prédiction
-	- rôle de la distribution de degrés
--
+How to add other types of interactions
+	- Wootton: variation arises from unaccounted interactions
 

### 2013-05-23 ms.tex

@@
 \usepackage[utf8]{inputenc}
-\usepackage{color,lineno,setspace,graphicx,multirow,kpfonts}
+\usepackage{color,lineno,setspace,multirow,kpfonts}
+\usepackage{graphicx}
 \usepackage[top=2.4cm,left=2.4cm,top=2.4cm,bottom=2.4cm,includefoot]{geometry}
@@
 \bibliography{./library}
+
+% TODO:
+% TP, PDP & DG: Complete the discussion
+% PDP: Additional paragraph on the metrics
+% DG: Figure legends
+% DG: Abstract
+% DG: Compile references
+% TP & PDP: fix bugs I generated!!!
 
@@
 
-\textbf{Title:}   Using neutral theory to reveal the contribution of dispersal to community assembly\\
+\textbf{Title:}   Using neutral theory to reveal the contribution of dispersal to community assembly in complex landscales\\
 
@@
 
-Community assembly is concerned by patterns and processes occurring at various
-spatial scales \parencite{Levin1992}. Until the development of metacommunity
-ecology, studies on community assembly were often restricted to local
-populations, with a strong focus on pairwise interactions (e.g.
-\textcite{MacArthur1972, May1973,Pimm1982,DeAngelis1992}). The emphasis on local
-communities has been vigorously criticized by \textcite{Ricklefs2008}, who has
-long recognized that local dynamics and community structure are strongly
-contingent on processes occurring at much larger spatial scales
-\parencite{Ricklefs1987}. This perspective is particularly relevant to
+Ecologists are required to move toward a predictive ecology,
+integrating elements of theoretical ecology \parencite{Thuiller2013}, and the
+metacommunity perspective appears naturally as the appropriate conceptual
+framework to develop the new modeling techniques required to fill this
+challenge. The metacommunity concept builds on 
+feedbacks between local scale processes, such as competitive interactions and
+local adaptation, and regional scale processes such as dispersal, gene flow and
+speciation \textcite{Leibold2004a}. This perspective is particularly relevant to
 limnology, where exchanges of organisms and nutrients affect community and
@@
 relative to pairwise interactions in the organization of ecological communities.
-The metacommunity concept has been proposed by \textcite{Leibold2004a} as a
-novel approach to link different spatial scales in ecology. It builds on
-feedbacks between local scale processes, such as competitive interactions and
-local adaptation, and regional scale processes such as dispersal, gene flow and
-speciation. Ecologists are now required to move toward a predictive ecology,
-integrating elements of theoretical ecology \parencite{Thuiller2013}, and the
-metacommunity perspective appears naturally as the appropriate conceptual
-framework to develop the new modeling techniques required to fill this
-challenge. The development of neutral theory has been quite provocative in that
+
+The development of neutral theory has been quite provocative in that
 respect, as one could see it as a step back in time. Neutral theory makes the
@@
 
-Neutral theory sparked an historical debate in community ecology that is still
+Neutral theory sparked an historical debate that is still
 lasting after more than a decade \parencite{Chave2004, Etiennee2011,
-Rosindell2012,Clark2012}. It was stimulated by the impressive ability of neutral
+Rosindell2012,Clark2012}. It was stimulated by the surprising ability of neutral
 models to fit several well studied empirical observations such as species
@@
 of various origins can blur the deterministic differences among species and
-promote the ecological drift \parencite{Gravel2011}. Much has been said the
+promote ecological drift \parencite{Gravel2011}. Much has been said the
 existence of demographic stochasticity, some ecologists even arguing that
 neutral models impede progress in community ecology by hidding niche differences
-\parencite{Clark2012}, and we therefore will keep this discussion for other
+\parencite{Clark2012}, and we therefore will let this discussion for other
 papers. The second answer is that dispersal and historical contingencies might
-have a much more profound impact on species distribution \parencite{Bahn2007}
+have a much more profound impact on species distribution \parencite{Bahn2007, Boulangeat2012}
 and ecological dynamics. The debate over the equivalence assumption and
@@
 habitats are characterized by a their discrete nature and spatially complex
-arrangements. We will also reveal the relative contribution of ecological
+arrangements \parencite{Peterson2013}. We will also reveal the relative contribution of ecological
 interaction and niche differentiation by contrasting predictions of a neutral
@@
 models and all simulations conducted for this paper. We then conduct simple
-simulations of these models to reveal the impact of spatial network structure on
+simulations of these models to reveal with simple examples the impact of spatial network structure on
 diversity distribution. We conclude with a discussion on the operationally of
@@
 Random geometric networks have been extensively studied \cite{app97a,app97
-b,app02a,app02b,pen03} and we provide a R function to generate them. We also
-provide the code for another structure that we call the geometric tree of
-doom. It builds a tree from the shortest path tree \cite{dij59} of a vertex in
-a random geometric network. This random geometric tree of doom looks like a
-lake connected by rivers to a series of smaller lakes.
+b,app02a,app02b,pen03} and we provide a R function to generate them. The 
+position of vertices is typically random, but we could also imagine alterations
+where they are either more aggregated or segregated than expected by chance alone.
+
+We also provide the code for a second structure that we call a random geometric tree. 
+It builds a tree from the shortest path tree \cite{dij59} of a vertex in a random geometric 
+network. This random geometric tree does not exaxctly represent dendritic landscapes 
+but is a convenient model to simulate lake connected by rivers to a series of smaller lakes.
 
@@
 nonsympatric speciation \cite{des12,des12b}.
+
+% TO PHIL: IT WOULD BE GREAT IF YOU COULD EXPAND ONE OR TWO PARAGRAPHS ON THE 
+% MAIN STATISTICS USED TO DESCRIBE SPATIAL GRAPHS. AT LEAST PRESENT THE MAIN 
+% DIFFERENCE BETWEEN NETWORK AND VERTEX LEVEL METRICS, AND THEN REFER TO THE 
+% TABLE.
 
@@
 first review the fundamental assumptions of each model with their description
-(Table 1 summarizes the parameters and variables that are used), and then
+(Table 2 summarizes the parameters and variables that are used), and then
 briefly discuss their main predictions. Simulation results are presented in the
@@
 between local and regional dynamics arises because all $p_ix$ from the landscape
-are dependent from each other. Simulation are usually conducted to solve the
+are dependent from each other. Simulations are usually conducted to solve the
 model for a large landscape, but numerical solutions are theoretically possible.
@@
 \sum{p_i}$. Because in this model all species are equal, we expect the local
-species richness to be a linear function of the patch degree. Multi-species
+species richness to be a linear function of the patch degree (number of edges). Multi-species
 analysis of metapopulation models also reveals interesting predictions on other
-aspects of community organization at various spatial scales such as species-area
+aspects of community organization at various spatial scales such as the species-area
 relationship \parencite{Hanski1997}, and proved to be useful in conservation
@@
 population based. We therefore considered in our toy model of neutral dynamics
-that each local patch is made of $J_x$ individuals. The model tracts the local
+that each local patch holds $J_x$ individuals. The model tracts the local
 abundance of all species $N_{ix}$ in each local patch. At each time step an
@@
 \parencite{Gravel2006}. The approach is conceptually similar to placing a seed
-trap in a gap in a forest and picking a seed at random among the one that fall
+trap in a canopy gap and picking a seed at random among the ones falling
 in to determine the identity of the recruited species. The composition of the
@@
 recruit is a migrant from neighbouring patches, $M$ is the probability it comes
-from a larger (and fixed) metacommunity, and consequently by substraction
+from a larger (and fixed) metacommunity, and consequently, by substraction,
 $1-m-M$ is the probability it comes from local dispersal. The fraction
@@
 (and without loss of generality, \parencite{Bell2000}) that the relative
-abundance in the metacommunity is uniform, ie equal to $S^{-1}$. This
+abundance in the metacommunity is uniform, i.-e. equal to $S^{-1}$. This
 immigration prevents the collapse of the metacommunity because otherwise all
@@
 The model is neutral because it assumes that the probabilities of a local
-recruitment, an immigration and a mortality event are all equal. Demographic
-stochasticity is the source of variations in abundance but larger disturbances
+recruitment, an immigration and a mortality event are all equal across species. Demographic
+stochasticity is the source of variations in abundance, but larger disturbances
 could be simulated as well, as long as they hit all species with the same
@@
 the predictions it makes. \textcite{Bell2001} and \textcite{Hubbell2001}
-analyzed the ability of neutral models to predict species abundance
+analyzed the performance of neutral models to predict species abundance
 distributions, range-abundance relationship, spatial variation in abundance,
 species-area relationship, community turnover (beta-diversity) and
-co-occurrence. Other than the ecological equivalence assumption, one of the most
+co-occurrence. Recent trophic neutral models also predicted realistic ecological networks \parenciteCanard2012}. Other than the ecological equivalence assumption, one of the most
 criticized aspect of neutral models is the realism of the speciation process and
@@
 \parencite{Leibold2004a}. There are various ways to simulate such dynamics and
-we picked the lottery models, in line with tradition \parencite{Mouquet2002} and
+we picked the lottery model, in line with tradition \parencite{Mouquet2002} and
 for its proximity to the neutral model described above. Competition for space
 occurs during recruitment after the death of an adult. The recruitment is a
-lottery against potential candidates. The recruitment probability is however
-biased by species specific response to local environmental conditions.
+lottery among potential candidates as in the neutral model. The recruitment probability 
+is however biased by species specific response to local environmental conditions.
 
@@
 will survive and develop to the adult stage. The effect of a differentiation to
-local environmental conditions could be implemented at this stage.The $J_x$
-individuals all experience unique environmental conditions $E_{nx}$ called a
-microsite $n$. We considered a patch average $\overline{E_x}$, with a
+local environmental conditions could be implemented at this stage with a biased 
+survival probability. The $J_x$ individuals all experience a unique environmental condition 
+$E_{nx}$ called a microsite $n$. We considered a patch average $\overline{E_x}$, with a
 within-patch variance $\sigma_x$. The regional average is $\overline{E_R}$ and
@@
 where recruitment occurs will survive. The recruitment probability is therefore
-biased because only some species will be able to cope with the local
-environmental conditions. For tractability we define the relative abundance in
+biased in favour of the species with highest survival because only some species will be able to cope with the local environmental conditions. For tractability we define the relative abundance in
 the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$. The
@@
 \frac{\lambda_{inx}Z_{ix}}{\sum \lambda_{jnx}Z_{jx}}$. The function describing
-the relationship between microsite conditions and survival could take various
-forms and we used the traditional gaussian curve describing the niche,
+the relationship between a microsite condition and survival could take various
+forms; we used the traditional gaussian curve describing the niche,
 $\lambda_{inx} =  \exp{-\frac{(E_{nx}-u_i)^2}{2\Pi b_i^2}}$, where $u_i$ is the
@@
 well-studied prediction of neutral models is the species abundance distribution.
-It was shown that niche models can predict similar distributions
+It was shown that niche models can predict similar distributions given appropriate parameters
 \parencite{Tilman2004,Gravel2006}. The main prediction is nonetheless that
@@
 effects of local heterogeneity and niche breadth because coexistence requires a
-limiting similarity \parencite{Schwilk2005}. Local species richness could be
+sufficient dissimilaritya mong species \parencite{Schwilk2005}. Local species richness could be
 increased by a mass effect when dispersal is consistently supplying individuals
@@
 
-In fig. 1, we illustrate the outcome of simulations using the neutral model on
-the four different network shapes considered in this article. All of these
+In this section we provide simple simulation results to illustrate the impact of saptial contingencies on species distribution and coexistence.
+We consider four different landscapes, illustrated at Fig.
+1.
+with the outcome of simulations using the neutral model. All of these
 networks have the same number of nodes (e.g. spatial sampling sites), but both
@@
 scenarios represented by each meta-community model.
+Our analysis is not exhaustive, it is provided simply to illustrate the interaction between metacommunity perspectives and landscape structures on alpha, beta and gamma diversity.
 
-In fig. 2, we present the richness of each node of the network, as a function of
-the centrality of the node, under different assumptions of meta-community
-dynamics and network structure. It appears that both in the goeographical and
-tree graph, the path dynamics model allows more diversity to persist overall.
-However, the local diversity increases with the node centrality, whether it is
+In Fig. 2, we present the species richness of each node of the network (alpha diversity), as a function of
+the centrality of the node, under different assumptions of metacommunity
+dynamics and network structure. We scaled the species richness by the maximal local species richness to 
+facilitate comparison between models. The model parameterization is responsible for differences in both alpha and gamma diversity.  It appears that both in the goeographical and
+tree graph, the path dynamics model has a much more considerable variation in local species richness. 
+However, in all cases the alpha diversity increases with the node centrality, whether it is
 measured by the degree or the eigencentrality. All situations examined give a
-remarkably congruent signal. The patch dynamics model, \emph{i.e.} when there
-are no interactions between species, allows more $\alpha$ diversity. The neutral
-model allowed the maintenance of slightly more diversity than the
-species-sorting model. In all cases however, it is clear that the richness
-maintained tend to increase with the node centrality.
+remarkably congruent signal.
 
-Finally, fig. 3 and fig. 4 present, respectively, the between patch $\beta$
-diversity as a function of the shape of the network, under three dynamic models.
-In figure 3, the distance is expressed as the euclidean (geographic) distance
+% TO CHANGE WITH NEW FIGURE (SCALED DIVERSITY)
+%The patch dynamics model, \emph{i.e.} when there
+%are no interactions between species, allows more $\alpha$ diversity. The neutral
+%model allowed the maintenance of slightly more diversity than the
+%species-sorting model. In all cases however, it is clear that the richness
+%maintained tend to increase with the node centrality.
+
+Finally, Figs. 3 and 4 present, respectively, the between patch $\beta$
+diversity as a function of the shape of the network, under the three dynamic models. We simply used Bray-Curtis measure of dissimilarity between patches. 
+In Fig. 3, the distance is expressed as the euclidean (geographic) distance
 between two patches. Although this neglects how dispersal connects the different
-patches, there is already a clear signal of $\beta$ diversity. In both the
+patches, there is already a clear signal of geographic distance on $\beta$ diversity. In both the
 neutral and patch dynamics model, local communities become increasingly
@@
 communities will be.
+Note this relationship for SS varies significantly with the spatial distribution of microsites (not shown).
+At one extreme, if all patches hold the same average conditions, then we should expect no relationship between dissimilarity and distance.
+On the other hand, if the average conditions are highly variable among localities (as in here), then we should expect two communities close to be potentially dissimilar (if conditions are different) or similar (if they are the same).
+The variance should thus be larger.
+A distance-dissimilarity relationship arises in the situation where dispersal promotes a mass effect (as in here).
+Such results emphasize the interaction between spatial contingencies (here connectivity and distribution of environmental conditions).
 
 To a vast extent, these relationships are preseved when looking at the
-topological distances, i.e. along how many edges should one travel to connect
+topological distances (Fiig. 4), i.e. along how many edges should one travel to connect
 two patchs. Interestingly enough, the distance-dissimilarity relationship for
@@
 
+%TP:
 Contrasting the three perspectives: what do we learn?
 
+%TP OR DG
 Network approach brings spatial variation in coexistence mechanisms; emergent properties
 	- role of the degree distribution
-
+%DG
 Making the theory operational
@@
 
+%PDP
 Working with more realistic landscapes
 
+%TP
 How to add other types of interactions

### 2013-05-23 ms.tex
New  results of figure 2

@@
 
-Ecologists are required to move toward a predictive ecology,
-integrating elements of theoretical ecology \parencite{Thuiller2013}, and the
-metacommunity perspective appears naturally as the appropriate conceptual
-framework to develop the new modeling techniques required to fill this
-challenge. The metacommunity concept builds on 
-feedbacks between local scale processes, such as competitive interactions and
-local adaptation, and regional scale processes such as dispersal, gene flow and
-speciation \textcite{Leibold2004a}. This perspective is particularly relevant to
-limnology, where exchanges of organisms and nutrients affect community and
-ecosystem properties from the local (e.g. vertical mixing
+Ecologists are required to move toward a predictive ecology, integrating
+elements of theoretical ecology \parencite{Thuiller2013}, and the metacommunity
+perspective appears naturally as the appropriate conceptual framework to develop
+the new modeling techniques required to fill this challenge. The metacommunity
+concept builds on feedbacks between local scale processes, such as competitive
+interactions and local adaptation, and regional scale processes such as
+dispersal, gene flow and speciation \textcite{Leibold2004a}. This perspective is
+particularly relevant to limnology, where exchanges of organisms and nutrients
+affect community and ecosystem properties from the local (e.g. vertical mixing
 \parencite{Ryabov2011}) to the regional (e.g. connection of lakes
@@
 
-The development of neutral theory has been quite provocative in that
-respect, as one could see it as a step back in time. Neutral theory makes the
-provocative assumption that species are ecologically equivalent and thereby any
-variation in the environment has no impact on demography
-\parencite{Bell2000,Hubbell2001}. Only demographic stochasticity and dispersal
-drive the structure of neutral ecological communities. It therefore appears
-that, on first sight, neutral theory is useless.  We will develop in this paper
-the argument that neutral theory could be a useful tool to understand the impact
-of dispersal on community organization in landscapes of various complexities.
+The development of neutral theory has been quite provocative in that respect, as
+one could see it as a step back in time. Neutral theory makes the provocative
+assumption that species are ecologically equivalent and thereby any variation in
+the environment has no impact on demography \parencite{Bell2000,Hubbell2001}.
+Only demographic stochasticity and dispersal drive the structure of neutral
+ecological communities. It therefore appears that, on first sight, neutral
+theory is useless.  We will develop in this paper the argument that neutral
+theory could be a useful tool to understand the impact of dispersal on community
+organization in landscapes of various complexities.
 
-Neutral theory sparked an historical debate that is still
-lasting after more than a decade \parencite{Chave2004, Etiennee2011,
-Rosindell2012,Clark2012}. It was stimulated by the surprising ability of neutral
-models to fit several well studied empirical observations such as species
-abundance distributions and distance-decay relationships. A remarkable strenght
-of neutral theory is to provide a \emph{"formal general theory of abundance
-and diversity that will account, in a simple and economical fashion, for the
-many patterns that ecologists have documented"} \parencite{Bell2001}. Even if
-new studies rejecting neutral theory are consistently published (e.g.
-\textcite{Ricklefs2012}), there is now almost a consensus that neutral theory is
-a well-developed null hypothesis for niche theory and could even be used as an
-adequate approximation of ecological dynamics in some situations.
-\textcite{Bell2001} nicely envisionned two perspectives to neutral theory that
-are still standing today. Under the weak perspective, neutral theory provides a
-set of realistic predictions of community organization despite false
-assumptions. Even if being fundamentally wrong, neutral theory is still useful
-when used as a null hypothesis \parencite{Gotelli2006}. It is considered as an
-improvement over traditional null hypotheses based on randomization
-\parencite{Gotelli2000} because it readily integrates dispersal. The strong
-version on the other hand posits that neutral theory is a satisfying
-approximation to community dynamics and an appropriate theory to explain the
-distribution of biodiversity. It implies that the right mechanisms have been
-identified and that the consistently observed differences among species do not
-impact community organization.
+Neutral theory sparked an historical debate that is still lasting after more
+than a decade \parencite{Chave2004, Etiennee2011, Rosindell2012,Clark2012}. It
+was stimulated by the surprising ability of neutral models to fit several well
+studied empirical observations such as species abundance distributions and
+distance-decay relationships. A remarkable strenght of neutral theory is to
+provide a \emph{"formal general theory of abundance and diversity that will
+account, in a simple and economical fashion, for the many patterns that
+ecologists have documented"} \parencite{Bell2001}. Even if new studies rejecting
+neutral theory are consistently published (e.g.  \textcite{Ricklefs2012}), there
+is now almost a consensus that neutral theory is a well-developed null
+hypothesis for niche theory and could even be used as an adequate approximation
+of ecological dynamics in some situations.  \textcite{Bell2001} nicely
+envisionned two perspectives to neutral theory that are still standing today.
+Under the weak perspective, neutral theory provides a set of realistic
+predictions of community organization despite false assumptions. Even if being
+fundamentally wrong, neutral theory is still useful when used as a null
+hypothesis \parencite{Gotelli2006}. It is considered as an improvement over
+traditional null hypotheses based on randomization \parencite{Gotelli2000}
+because it readily integrates dispersal. The strong version on the other hand
+posits that neutral theory is a satisfying approximation to community dynamics
+and an appropriate theory to explain the distribution of biodiversity. It
+implies that the right mechanisms have been identified and that the consistently
+observed differences among species do not impact community organization.
 
@@
 papers. The second answer is that dispersal and historical contingencies might
-have a much more profound impact on species distribution \parencite{Bahn2007, Boulangeat2012}
-and ecological dynamics. The debate over the equivalence assumption and
-demographic stochasticity has perhaps overlook the recognition of how much
-dispersal influence community assembly.
+have a much more profound impact on species distribution \parencite{Bahn2007,
+Boulangeat2012} and ecological dynamics. The debate over the equivalence
+assumption and demographic stochasticity has perhaps overlook the recognition of
+how much dispersal influence community assembly.
 
 In this paper, we argue that neutral theory is useful both to understand and
-predict the impact of dispersal on community organization. Even
-for purely theoretical analyses, we need a benchmark without niche differences to
-reveal the role of dispersal in structuring communities, and understand how it
+predict the impact of dispersal on community organization. Even for purely
+theoretical analyses, we need a benchmark without niche differences to reveal
+the role of dispersal in structuring communities, and understand how it
 interacts with niche differentiation. We will explore recent applications of
@@
 habitats are characterized by a their discrete nature and spatially complex
-arrangements \parencite{Peterson2013}. We will also reveal the relative contribution of ecological
-interaction and niche differentiation by contrasting predictions of a neutral
-model to other metacommunity perspectives.
+arrangements \parencite{Peterson2013}. We will also reveal the relative
+contribution of ecological interaction and niche differentiation by contrasting
+predictions of a neutral model to other metacommunity perspectives.
 
@@
 models and all simulations conducted for this paper. We then conduct simple
-simulations of these models to reveal with simple examples the impact of spatial network structure on
-diversity distribution. We conclude with a discussion on the operationally of
-the framework.
+simulations of these models to reveal with simple examples the impact of spatial
+network structure on diversity distribution. We conclude with a discussion on
+the operationally of the framework.
 
@@
 
-We also provide the code for a second structure that we call a random geometric tree. 
-It builds a tree from the shortest path tree \cite{dij59} of a vertex in a random geometric 
-network. This random geometric tree does not exaxctly represent dendritic landscapes 
-but is a convenient model to simulate lake connected by rivers to a series of smaller lakes.
+We also provide the code for a second structure that we call a random geometric
+tree.  It builds a tree from the shortest path tree \cite{dij59} of a vertex in
+a random geometric network. This random geometric tree does not exaxctly
+represent dendritic landscapes but is a convenient model to simulate lake
+connected by rivers to a series of smaller lakes.
 
@@
 flow of organisms \cite{min07,fal07,min08,gar08,urb09,dal10}. In the neutral
-theory, networks were pioneered by Economo and Keitt \cite{eco08, eco10}.
-They used networks to study how different spatial structures influenced
-diversity. They were also used to study how the spatial structure influenced
-nonsympatric speciation \cite{des12,des12b}.
+theory, networks were pioneered by Economo and Keitt \cite{eco08, eco10}.  They
+used networks to study how different spatial structures influenced diversity.
+They were also used to study how the spatial structure influenced nonsympatric
+speciation \cite{des12,des12b}.
 
@@
 
-In this section we provide simple simulation results to illustrate the impact of saptial contingencies on species distribution and coexistence.
-We consider four different landscapes, illustrated at Fig.
-1.
-with the outcome of simulations using the neutral model. All of these
-networks have the same number of nodes (e.g. spatial sampling sites), but both
-different number of edges (e.g. dispersal routes between sampling sites) and
-patterns of connectivity between nodes. We ask how these differences in
-connectivity will shape the emerging properties of the community under the
-scenarios represented by each meta-community model.
-Our analysis is not exhaustive, it is provided simply to illustrate the interaction between metacommunity perspectives and landscape structures on alpha, beta and gamma diversity.
+In this section we provide simple simulation results to illustrate the impact of
+saptial contingencies on species distribution and coexistence. We consider four
+different landscapes, illustrated at Fig. 1,  with the outcome of simulations
+using the neutral model. All of these networks have the same number of nodes
+(e.g. spatial sampling sites), but both different number of edges (e.g.
+dispersal routes between sampling sites) and patterns of connectivity between
+nodes. We ask how these differences in connectivity will shape the emerging
+properties of the community under the scenarios represented by each
+meta-community model. Our analysis is not exhaustive, but is provided simply to
+illustrate the interaction between metacommunity perspectives and landscape
+structures on alpha, beta and gamma diversity.
 
-In Fig. 2, we present the species richness of each node of the network (alpha diversity), as a function of
-the centrality of the node, under different assumptions of metacommunity
-dynamics and network structure. We scaled the species richness by the maximal local species richness to 
-facilitate comparison between models. The model parameterization is responsible for differences in both alpha and gamma diversity.  It appears that both in the goeographical and
-tree graph, the path dynamics model has a much more considerable variation in local species richness. 
-However, in all cases the alpha diversity increases with the node centrality, whether it is
-measured by the degree or the eigencentrality. All situations examined give a
-remarkably congruent signal.
+In Fig. 2, we present the species richness of each node of the network ($\alpha$
+diversity), as a function of the centrality of the node, under different
+assumptions of metacommunity dynamics and network structure. We scaled the
+species richness by the maximal local species richness to facilitate comparison
+between models. The model parameterization is responsible for differences in
+both $\alpha$ and $\gamma$ diversity, meaning that only the shape of the
+relationship between centrality and richness ought to be looked at.  It appears
+that both in the geographical and tree graph, the path dynamics model has a
+much more considerable variation in local species richness.  However, in all
+cases the alpha diversity increases with the node degree centrality, meaning
+that nodes with more connections also host a more diverse community.
+Eigen-centrality gave a far less clear-cut result, which can probably be
+attributed to the fact that our networks are relatively small in size.
+Eigen-centrality reports how well your neighbors are connected, and in graphs
+with a short diameter (i.e. the two farthest points are not extremely far
+apart), this measure might hold less information.
 
@@
 Finally, Figs. 3 and 4 present, respectively, the between patch $\beta$
-diversity as a function of the shape of the network, under the three dynamic models. We simply used Bray-Curtis measure of dissimilarity between patches. 
-In Fig. 3, the distance is expressed as the euclidean (geographic) distance
-between two patches. Although this neglects how dispersal connects the different
-patches, there is already a clear signal of geographic distance on $\beta$ diversity. In both the
-neutral and patch dynamics model, local communities become increasingly
-dissimilar when the distance between them increases. In other words, two
-communities which are close to each other will share a large proportion of their
-species, whereas two communities which are afar will share a small proportion.
-In the species-sorting model, the relationship between distance and
-dissimilarity is similar, Nonetheless, it forms an enveloppe of points (with
-most points lying in the upper-left part of the graph). While two distant
-communities will be dissimilar, there is no telling how dissimilar two close
-communities will be.
-Note this relationship for SS varies significantly with the spatial distribution of microsites (not shown).
-At one extreme, if all patches hold the same average conditions, then we should expect no relationship between dissimilarity and distance.
-On the other hand, if the average conditions are highly variable among localities (as in here), then we should expect two communities close to be potentially dissimilar (if conditions are different) or similar (if they are the same).
-The variance should thus be larger.
-A distance-dissimilarity relationship arises in the situation where dispersal promotes a mass effect (as in here).
-Such results emphasize the interaction between spatial contingencies (here connectivity and distribution of environmental conditions).
+diversity as a function of the shape of the network, under the three dynamic
+models. We simply used Bray-Curtis measure of dissimilarity between patches.  In
+Fig. 3, the distance is expressed as the euclidean (geographic) distance between
+two patches. Although this neglects how dispersal connects the different
+patches, there is already a clear signal of geographic distance on $\beta$
+diversity. In both the neutral and patch dynamics model, local communities
+become increasingly dissimilar when the distance between them increases. In
+other words, two communities which are close to each other will share a large
+proportion of their species, whereas two communities which are afar will share a
+small proportion.  In the species-sorting model, the relationship between
+distance and dissimilarity is similar, Nonetheless, it forms an enveloppe of
+points (with most points lying in the upper-left part of the graph). While two
+distant communities will be dissimilar, there is no telling how dissimilar two
+close communities will be. Note this relationship for SS varies significantly
+with the spatial distribution of microsites (not shown). At one extreme, if all
+patches hold the same average conditions, then we should expect no relationship
+between dissimilarity and distance. On the other hand, if the average conditions
+are highly variable among localities (as in here), then we should expect two
+communities close to be potentially dissimilar (if conditions are different) or
+similar (if they are the same). The variance should thus be larger. A
+distance-dissimilarity relationship arises in the situation where dispersal
+promotes a mass effect (as in here). Such results emphasize the interaction
+between spatial contingencies (here connectivity and distribution of
+environmental conditions).
 

### 2013-05-23 ms.tex
Removed commented lines
Edits on Results

--- 2013-05-23 ms.tex
@@
 Finally, Figs. 3 and 4 present, respectively, the between patch $\beta$
-diversity as a function of the shape of the network, under the three dynamic models. We simply used Bray-Curtis measure of dissimilarity between patches. 
-In Fig. 3, the distance is expressed as the euclidean (geographic) distance
-between two patches. Although this neglects how dispersal connects the different
-patches, there is already a clear signal of geographic distance on $\beta$ diversity. In both the
-neutral and patch dynamics model, local communities become increasingly
-dissimilar when the distance between them increases. In other words, two
-communities which are close to each other will share a large proportion of their
-species, whereas two communities which are afar will share a small proportion.
-In the species-sorting model, the relationship between distance and
-dissimilarity is similar, Nonetheless, it forms an enveloppe of points (with
-most points lying in the upper-left part of the graph). While two distant
-communities will be dissimilar, there is no telling how dissimilar two close
-communities will be.
-Note this relationship for SS varies significantly with the spatial distribution of microsites (not shown).
-At one extreme, if all patches hold the same average conditions, then we should expect no relationship between dissimilarity and distance.
-On the other hand, if the average conditions are highly variable among localities (as in here), then we should expect two communities close to be potentially dissimilar (if conditions are different) or similar (if they are the same).
-The variance should thus be larger.
-A distance-dissimilarity relationship arises in the situation where dispersal promotes a mass effect (as in here).
-Such results emphasize the interaction between spatial contingencies (here connectivity and distribution of environmental conditions).
+diversity as a function of the shape of the network, under the three dynamic
+models. We simply used Bray-Curtis measure of dissimilarity between patches. In
+Fig. 3, the distance is expressed as the euclidean (geographic) distance between
+two patches. Although this neglects how dispersal connects the different
+patches, there is already a clear signal of geographic distance on $\beta$
+diversity. In both the neutral and patch dynamics model, local communities
+become increasingly dissimilar when the distance between them increases. In
+other words, two communities which are close to each other will share a large
+proportion of their species, whereas two communities which are afar will share a
+small proportion. In the species-sorting model, the relationship between
+distance and dissimilarity is similar, Nonetheless, it forms an enveloppe of
+points (with most points lying in the upper-left part of the graph). While two
+distant communities will be dissimilar, there is no telling how dissimilar two
+close communities will be. Note this relationship for SS varies significantly
+with the spatial distribution of microsites (not shown). At one extreme, if all
+patches hold the same average conditions, then we should expect no relationship
+between dissimilarity and distance. On the other hand, if the average conditions
+are highly variable among localities (as in here), then we should expect two
+communities close to be potentially dissimilar (if conditions are different) or
+similar (if they are the same). The variance should thus be larger. A
+distance-dissimilarity relationship arises in the situation where dispersal
+promotes a mass effect (as in here). Such results emphasize the interaction
+between spatial contingencies (here connectivity and distribution of
+environmental conditions).
 
 To a vast extent, these relationships are preseved when looking at the
-topological distances (Fiig. 4), i.e. along how many edges should one travel to connect
-two patchs. Interestingly enough, the distance-dissimilarity relationship for
-the neutral model is markedly hump-shaped, with sites being at a medium distance
-having the maximal dissimilarity.
+topological distances (Fiig. 4), i.e. along how many edges should one travel to
+connect two patchs. Interestingly enough, the distance-dissimilarity
+relationship for the neutral model is markedly hump-shaped, with sites being at
+a medium distance having the maximal dissimilarity.
 
@@
 
-%TP OR DG
-Network approach brings spatial variation in coexistence mechanisms; emergent properties
-	- role of the degree distribution
+% DG
+Network approach to spatial landscapes is a major advancement in metacommunity ecology.
+- First step to make the concept operational (instrumental perspective) because it accounts for more realistic landscape structures and dispersal kernels.
+- The realism comes at a first degree from spatially explicit dispersal instead of the simple island-mainland or global dispersal approaches used previously (e.g. Tilman, Mouquet, Hubbell)
+- But dispersal is also spatially explicit over a lattice and it does not make the landscape more realistic
+- The fundamental feature is the heterogeneity of dispersal
+- We found that the more central nodes are usually more diverse
+- They might also contribute more to maintain regional diversity, as they are essential for species to spread throughout the landscape.
+- Keystone communities \parencite{Mouquet2013} 
+- The comparison between metaco perspectives revealed some interesting results in that respect. 
+- The patch vs neutral comparison shows that competitive interactions....
+- On the other hand the effect of network topology is much weaker with the ss model.
+We could even hypothesize it will vanish in the extreme case of niche differentiation (with low overlap for instance) and low mass effect.
+- The neutral versus niche comparison therefore illustrates that very strong unequal competitive interactions could overwhelm the impact of dispersal.
+
+The network approach brings spatial variation in coexistence mechanisms.
+- If we take the SS perspective for instance, we find that alpha div. is higher in more central nodes
+- Since the environment is constant, it means diversity in these communities must be maintained by a stronger mass effect. 
+- It results in spatial variation in the relative importance of ss, mass effect and to a certain extent neutral drift. 
+- Because the degree centrality was the best variable explaining diversity (here; Altermat 2012; Desjardins 2012b), we should expect the degree distribution to strongly impact the relative importance of these coexistence mechanisms.
+- For a given set of ecological processes and distribution of species traits, we might expect the coexistence mechanisms to differ from one landscape to another.
+
 %DG
-Making the theory operational
-	- how to parameterize neutral models
-	- making predictions for specific landscapes
+We introduced this article arguing that neutral theory could be used as an instrument to make predictions. 
+- So far we have treated only theoretical models, but we could also envision to parameterize them and simulate real landscapes. 
+- The recruitment probabilities defined above could be used as satistical models to fit to empirical data. 
+- Prior information could be used to define apriori dispersal kernels  and then fit the model
+- Long tradition of this in metapopulation ecology (Hasnki 1998) and now new species distribution models are including dispersal limitations (Boulangeat2012)
+- Given the parameterization, one could run neutral models to generate null hypotheses that could be eventually compared to observed distribution. 
+- This would make a significant improvement over traditional null models in ecology (Gotelli1996) where there are no interactions and no dispersal limitations.
+- In addition, neutral models could be used to predict the consequences of habitat destruction and fragmentation (ref Hubbell PNAS)
+- The spatially explicit description of the landscape is a major improvement toward that end, providing much flexibility in the scenarios that could be explored.
 

### 2013-05-23 ms.tex
First paragraphs of the discussion

--- 2013-05-23 ms.tex
@@
 species-area relationship, community turnover (beta-diversity) and
-co-occurrence. Recent trophic neutral models also predicted realistic ecological networks \parenciteCanard2012}. Other than the ecological equivalence assumption, one of the most
+co-occurrence. Recent trophic neutral models also predicted realistic ecological networks \parencite{Canard2012}. Other than the ecological equivalence assumption, one of the most
 criticized aspect of neutral models is the realism of the speciation process and
@@
 
-In this section we provide simple simulation results to illustrate the impact of saptial contingencies on species distribution and coexistence.
-We consider four different landscapes, illustrated at Fig.
-1.
-with the outcome of simulations using the neutral model. All of these
-networks have the same number of nodes (e.g. spatial sampling sites), but both
-different number of edges (e.g. dispersal routes between sampling sites) and
-patterns of connectivity between nodes. We ask how these differences in
-connectivity will shape the emerging properties of the community under the
-scenarios represented by each meta-community model.
-Our analysis is not exhaustive, it is provided simply to illustrate the interaction between metacommunity perspectives and landscape structures on alpha, beta and gamma diversity.
+In this section we provide simple simulation results to illustrate the impact of
+saptial contingencies on species distribution and coexistence. We consider four
+different landscapes, illustrated at Fig. 1. with the outcome of simulations
+using the neutral model. All of these networks have the same number of nodes
+(e.g. spatial sampling sites), but both different number of edges (e.g.
+dispersal routes between sampling sites) and patterns of connectivity between
+nodes. We ask how these differences in connectivity will shape the emerging
+properties of the community under the scenarios represented by each
+meta-community model. Our analysis is not exhaustive, it is provided simply to
+illustrate the interaction between metacommunity perspectives and landscape
+structures on alpha, beta and gamma diversity.
 
-In Fig. 2, we present the species richness of each node of the network (alpha diversity), as a function of
-the centrality of the node, under different assumptions of metacommunity
-dynamics and network structure. We scaled the species richness by the maximal local species richness to 
-facilitate comparison between models. The model parameterization is responsible for differences in both alpha and gamma diversity.  It appears that both in the goeographical and
-tree graph, the path dynamics model has a much more considerable variation in local species richness. 
-However, in all cases the alpha diversity increases with the node centrality, whether it is
-measured by the degree or the eigencentrality. All situations examined give a
-remarkably congruent signal.
-
-% TO CHANGE WITH NEW FIGURE (SCALED DIVERSITY)
-%The patch dynamics model, \emph{i.e.} when there
-%are no interactions between species, allows more $\alpha$ diversity. The neutral
-%model allowed the maintenance of slightly more diversity than the
-%species-sorting model. In all cases however, it is clear that the richness
-%maintained tend to increase with the node centrality.
+In Fig. 2, we present the species richness of each node of the network (alpha
+diversity), as a function of the centrality of the node, under different
+assumptions of metacommunity dynamics and network structure. We scaled the
+species richness by the maximal local species richness to facilitate comparison
+between models. The model parameterization is responsible for differences in
+both alpha and gamma diversity and thus prevents meaningful absolute
+comparisons. It appears that both in the geographical and random tree graphs,
+the path dynamics model has a much more considerable variation in local species
+richness. However, in all cases the alpha diversity increases with the node
+degree centrality and not so much with eigencentrality. There are some clear
+differences among metacommunity perspectives. First, the relationship between
+degree centrality and species richness is strongest for patch dynamics, followed
+by neutral dynamics and then species sorting. Second, there is apparently much
+more scatter in the species sorting scenario. Thirdly, the richness-centrality
+relationship appears weaker for the random tree graph.
 
 Finally, Figs. 3 and 4 present, respectively, the between patch $\beta$
-diversity as a function of the shape of the network, under the three dynamic models. We simply used Bray-Curtis measure of dissimilarity between patches. 
-In Fig. 3, the distance is expressed as the euclidean (geographic) distance
-between two patches. Although this neglects how dispersal connects the different
-patches, there is already a clear signal of geographic distance on $\beta$ diversity. In both the
-neutral and patch dynamics model, local communities become increasingly
-dissimilar when the distance between them increases. In other words, two
-communities which are close to each other will share a large proportion of their
-species, whereas two communities which are afar will share a small proportion.
-In the species-sorting model, the relationship between distance and
-dissimilarity is similar, Nonetheless, it forms an enveloppe of points (with
-most points lying in the upper-left part of the graph). While two distant
-communities will be dissimilar, there is no telling how dissimilar two close
-communities will be.
-Note this relationship for SS varies significantly with the spatial distribution of microsites (not shown).
-At one extreme, if all patches hold the same average conditions, then we should expect no relationship between dissimilarity and distance.
-On the other hand, if the average conditions are highly variable among localities (as in here), then we should expect two communities close to be potentially dissimilar (if conditions are different) or similar (if they are the same).
-The variance should thus be larger.
-A distance-dissimilarity relationship arises in the situation where dispersal promotes a mass effect (as in here).
-Such results emphasize the interaction between spatial contingencies (here connectivity and distribution of environmental conditions).
+diversity as a function of the shape of the network, under the three dynamic
+models. We simply used Bray-Curtis measure of dissimilarity between patches. In
+Fig. 3, the distance is expressed as the euclidean (geographic) distance between
+two patches. Although this neglects how dispersal connects the different
+patches, there is already a clear signal of geographic distance on $\beta$
+diversity. In both the neutral and patch dynamics model, local communities
+become increasingly dissimilar when the distance between them increases. In
+other words, two communities which are close to each other will share a large
+proportion of their species, whereas two communities which are afar will share a
+small proportion. In the species-sorting model, the relationship between
+distance and dissimilarity is similar, Nonetheless, it forms an enveloppe of
+points (with most points lying in the upper-left part of the graph). While two
+distant communities will be dissimilar, there is no telling how dissimilar two
+close communities will be. Note this relationship for SS varies significantly
+with the spatial distribution of microsites (not shown). At one extreme, if all
+patches hold the same average conditions, then we should expect no relationship
+between dissimilarity and distance. On the other hand, if the average conditions
+are highly variable among localities (as in here), then we should expect two
+communities close to be potentially dissimilar (if conditions are different) or
+similar (if they are the same). The variance should thus be larger. A
+distance-dissimilarity relationship arises in the situation where dispersal
+promotes a mass effect (as in here). Such results emphasize the interaction
+between spatial contingencies (here connectivity and distribution of
+environmental conditions).
 
 To a vast extent, these relationships are preseved when looking at the
-topological distances (Fiig. 4), i.e. along how many edges should one travel to connect
-two patchs. Interestingly enough, the distance-dissimilarity relationship for
-the neutral model is markedly hump-shaped, with sites being at a medium distance
-having the maximal dissimilarity.
+topological distances (Fig. 4), i.e. along how many edges should one travel to
+connect two patchs. Interestingly enough, the distance-dissimilarity
+relationship for the neutral model is markedly hump-shaped, with sites being at
+a medium distance having the maximal dissimilarity.
 
@@
 
-%TP:
-Contrasting the three perspectives: what do we learn?
+Our objective in this paper was to review the main assumptions of three
+metacommunity models and illustrate how the implementation of more realistic
+landsapes could reveal the importance of dispersal on community structure. We
+argued in the introduction that neutral theory is useful both to understand and
+predict the impact of dispersal on community organization. The review of the
+different models shows that the fundamental difference between a neutral model
+and the patch dynamics model is the effect of competitive interactions on
+distribution, while the difference between the neutral and the species sorting
+models is the effect of unequal competitive interactions. The neutral model is
+thus a usefull tool to understand the joint effects of dispersal and community
+interactions. Our comparison of the distribution of alpha diversity was
+particularly meaningful in that respect. The simulation results show that
+competitive interactions buffer the potential impact of landscape structure. The
+strongest centrality-species richness relationship was observerd for the patch
+dynamics, a model without any interactions. On the other hand, strong and
+unequal competitive interactions minimized the effect of centrality. Our model
+analysis greatly illustrates the growing recognition in metacommunity ecology
+that we must move toward more realistic landscapes \parencite{Gilarranz2012}.
+For field ecologists, and particularly limnologists, our review emphasize that
+we need to go beyond geographic based analysis of beta-diversity (e.g.
+\parencite{Legendre2005}) to topological based analyses \parencite{Peterson2013}.
 
-%TP OR DG
-Network approach brings spatial variation in coexistence mechanisms; emergent properties
-	- role of the degree distribution
-%DG
-Making the theory operational
-	- how to parameterize neutral models
-	- making predictions for specific landscapes
+The network approach to the study of spatially explicit landscapes is a major
+advancement in metacommunity ecology. It is a first step to make the concept
+operational because it accounts for more realistic landscape structures and
+dispersal kernels. It makes a significant departure to simple island-mainland or
+global dispersal approaches used previously (e.g. \parencite{Tilman1994,
+Mouquet2002, Hubbell2001}). But dispersal is also spatially explicit in a
+lattice model and it does not make the landscape more realistic. We believe the
+fundamental contribution of this approach is the consideration of spatial
+heterogeneity of dispersal. In agreement with previous theoretical
+\parencite{Economo2011, Desjardins2012} and experimemental studies
+\parencite{Altermat2012}, the simulations show that the degree centrality has a
+significant impact on local species richness. Central nodes might also
+contribute more to maintain regional diversity, as they are essential for
+species to spread throughout the landscape. The nodes could be potentially
+quantified as keystone for the metacommunity \parencite{Mouquet2013}.
+Interestingly, but not surprisingly, this effect is weaker with species sorting
+dynamics. We could even hypothesize it will vanish in the extreme case of niche
+differentiation (with low overlap for instance) and low mass effect. The neutral
+versus niche comparison therefore illustrates that very strong unequal
+competitive interactions could overwhelm the impact of dispersal.
+
+The network approach and the comparison between metacommunity perspectives
+reveal there could be spatial variation in coexistence mechanisms. If we take
+the species-sorting perspective for instance, we find that alpha diversity is
+higher in more central nodes. Since the environment is constant, it implies that
+diversity in these communities is maintained by a stronger mass effect. It
+results in spatial variation in the relative importance of species-sorting, the
+mass effect and to a certain extent the neutral drift. Because the degree
+centrality was the best variable explaining diversity, we should expect the
+degree distribution to strongly impact the relative importance of these
+coexistence mechanisms. For a given set of ecological processes and distribution
+of species traits, we might expect the coexistence mechanisms to differ from one
+landscape to another.
+
+We introduced this article arguing that neutral theory could be used as an
+instrument to predict species distribution in spatially heterogeneous
+landscapes. So far we have treated only theoretical models, but we could also
+envision to parameterize them and simulate real landscapes. The recruitment
+probabilities defined above could be used as satistical models (likelihood
+functions) to fit to empirical data. Prior information could be used to define
+apriori dispersal kernels and then fit the model as in \textcite{Gravel2008}.
+The fit of metapopulation models to spatially explicit landscapes was pionneered
+by \textcite{Hanski1998} and recently extended to species distribution models
+including both species sorting and dispersal limitations
+\parencite{Boulangeat2012}. Given the parameterization, one could run neutral
+models to generate null hypotheses that could be eventually compared to observed
+distribution. This would make a significant improvement over traditional null
+models in ecology (Gotelli1996) where there are no interactions and no dispersal
+limitations.
+
+The multivariate variance partitioning framework originally proposed by Borcard
+et al. (1992) and further developped by Borcard and Legendre (2002) has been
+widely used to quantify the relative importance of species sorting and dispersal
+limitations in species distribution. This framework was originally proposed to
+model species distribution as a function of environmental variables, taking into
+accound the spatial autocorrelation of species distriution (Leduc1992,
+Borcard1992, Legendre1993). This methodology has been widely used over the last
+decade as a test of the neutral theory, its underlying assumption and a
+quantification of dispersal limitations (e.g. Svenning2004; Hardy2004;
+Gilbert2004; Cottenie2005). One problem of this approach is however that it
+makes a weak test of neutrality (McGill2003), based on the description of
+spatial community structure, rather than hypothesis testing. The different
+models we reviewed in this article could be better employed if used to generate
+expectations based on different hypotheses and then compare them. But most of
+all, parameterized spatially explicit neutral models could be more useful if
+used to generate expectations. For instance, neutral models could be used to
+predict the consequences of habitat destruction, fragmentation or a change in
+the connectinity matrix \parencite{Hubbell2008}. The spatially explicit
+description of the landscape is a major improvement toward that end, providing
+much flexibility in the scenarios that could be explored.
 

### 2013-05-23 codeR/lottery_model.R

@@ -1,0 +1,122 @@
+#######################################################################
+# Individual based lottery model with spatially explicit
+# dispersal and environmental heterogeneity
+# 
+# The model collapse to a neutral model when the niche breadth tends to infinity
+#
+# Store the results in a time series of occupancy, a presence-absence matrix
+# and a local abundance matrix
+# 
+# By: Dominique Gravel (dominique_gravel@uqar.ca) 
+# April 2013
+########################################################################
+
+# Main function
+lottery_model = function(m,M,d,S,J,sdE,sdN,spatial_graph,nsteps) {
+	# Args: 
+	# 	m: immigration probability from the neighbourhood
+	#	M: immigration probability from outside the metacommunity 
+	#	d: local death rate 	
+	# 	S: number of species
+	#	J: local community size
+	# 	sdE: within patch standard deviation of the environment
+	# 	sdN: niche breadth
+	#	spatial_graph: a spatial graph object
+	#	nsteps: number of time steps to run the simulation
+	#
+	# Returns:
+	#	A list with a time series of occupancy for each species, a site presence-absence matrix	
+	#	and a site-abundance matrix from the last time step
+	#	
+	########################################
+	# Prepare the simulation
+	
+	####################
+	# Adjency matrix
+	adjMat = spatial_graph[[2]] 	# The original matrix
+	n = nrow(adjMat)				# Number of nodes
+	degrees = apply(adjMat,2,sum) 	# Number of degrees for each node
+	w = adjMat*matrix(degrees^-1,nr = n,nc=n, byrow=T) # Weighted adjancy matrix
+
+	####################
+	# Local environmental conditions
+	Env = matrix(nr=n, nc = J)
+	for(i in 1:n) Env[i,] = rnorm(J,mean = runif(1,0,100), sd = sdE) 	# For a random distribution of environments
+#	for(i in 1:n) Env[i,] = rnorm(J,mean = i/n*100, sd = 0)				# For a uniform distribution of environments		
+		
+	####################
+	# Niche optimums
+	u = runif(S, 0, 100) 	# For a random distribution of optimums
+#	u = 100*c(1:S)/S		# For a uniform distribution of optimums
+	
+	####################
[…]

### 2013-05-23 codeR/simulations.R

@@ -1,6 +1,6 @@
 ########################################################################
 # Source for the figures 1-4 in the paper
-# Gravel, Poisot and Desjardins. 2013. XXX. J. Limnology. 
+# Gravel, Poisot and Desjardins. 2013. Using neutral theory to reveal the contribution of dispersal to community assembly in complex landscapes. J. Limnology. 
 #
 # By: Dominique Gravel (dominique_gravel@uqar.ca) 
 # April 2013
@@ -195,25 +195,37 @@ t = matrix(c(1:4),nr = 2, nc = 2, byrow = T)
 layout(t)
 layout.show(4)
 
+apply(patchGeograph[[1]],2,sum)
+
+maxPatchGeograph = max(apply(patchGeograph[[1]],1,sum))
+maxNeutralGeograph = max(apply(neutralGeograph[[1]],1,sum))
+maxSSGeograph = max(apply(ssGeograph[[1]],1,sum))
+
+maxPatchGeotree = max(apply(patchGeotree[[1]],1,sum))
+maxNeutralGeotree = max(apply(neutralGeotree[[1]],1,sum))
+maxSSGeotree = max(apply(ssGeotree[[1]],1,sum))
+
 par(mar=c(3,6,4,1))
-plot(degGeograph,apply(patchGeograph[[1]],1,sum),pch = 19, xlab = "", ylab = "Species richness", ylim = c(0,S),cex.lab = 1.5, cex.axis = 1.25)
-points(degGeograph,apply(neutralGeograph[[1]],1,sum),pch = 21, bg = "red")
-points(degGeograph,apply(ssGeograph[[1]],1,sum),pch = 21, bg = "blue")
+plot(degGeograph,apply(patchGeograph[[1]],1,sum)/maxPatchGeograph,pch = 19, xlab = "", ylab = "Species richness 
+(scaled)", ylim = c(0,1),cex.lab = 1.5, cex.axis = 1.25)
+points(degGeograph,apply(neutralGeograph[[1]],1,sum)/maxNeutralGeograph,pch = 21, bg = "red")
+points(degGeograph,apply(ssGeograph[[1]],1,sum)/maxSSGeograph,pch = 21, bg = "blue")
 
 par(mar=c(3,4,4,3))
[…]

### 2013-05-23 ms.tex

@@
 species-area relationship, community turnover (beta-diversity) and
-co-occurrence. Recent trophic neutral models also predicted realistic ecological networks \parencite{Canard2012}. Other than the ecological equivalence assumption, one of the most
+co-occurrence. Recent trophic neutral models also predicted realistic ecological networks \parenciteCanard2012}. Other than the ecological equivalence assumption, one of the most
 criticized aspect of neutral models is the realism of the speciation process and
@@
 
-In Fig. 2, we present the species richness of each node of the network (alpha
+In Fig. 2, we present the species richness of each node of the network ($\alpha$
 diversity), as a function of the centrality of the node, under different
@@
 between models. The model parameterization is responsible for differences in
-both alpha and gamma diversity and thus prevents meaningful absolute
-comparisons. It appears that both in the geographical and random tree graphs,
-the path dynamics model has a much more considerable variation in local species
-richness. However, in all cases the alpha diversity increases with the node
-degree centrality and not so much with eigencentrality. There are some clear
-differences among metacommunity perspectives. First, the relationship between
-degree centrality and species richness is strongest for patch dynamics, followed
-by neutral dynamics and then species sorting. Second, there is apparently much
-more scatter in the species sorting scenario. Thirdly, the richness-centrality
-relationship appears weaker for the random tree graph.
+both $\alpha$ and $\gamma$ diversity, meaning that only the shape of the
+relationship between centrality and richness ought to be looked at.  It appears
+that both in the geographical and tree graph, the path dynamics model has a
+much more considerable variation in local species richness.  However, in all
+cases the alpha diversity increases with the node degree centrality, meaning
+that nodes with more connections also host a more diverse community.
+Eigen-centrality gave a far less clear-cut result, which can probably be
+attributed to the fact that our networks are relatively small in size.
+Eigen-centrality reports how well your neighbors are connected, and in graphs
+with a short diameter (i.e. the two farthest points are not extremely far
+apart), this measure might hold less information.
 
@@
 To a vast extent, these relationships are preseved when looking at the
-topological distances (Fig. 4), i.e. along how many edges should one travel to
-connect two patchs. Interestingly enough, the distance-dissimilarity
-relationship for the neutral model is markedly hump-shaped, with sites being at
-a medium distance having the maximal dissimilarity.
+topological distances (Fiig. 4), i.e. along how many edges should one travel to connect
+two patchs. Interestingly enough, the distance-dissimilarity relationship for
+the neutral model is markedly hump-shaped, with sites being at a medium distance
+having the maximal dissimilarity.
 

### 2013-05-23 ms.tex
Neutrality and other interactions

@@
 diversity as a function of the shape of the network, under the three dynamic
-models. We simply used Bray-Curtis measure of dissimilarity between patches.  In
+models. We used Bray-Curtis measure of dissimilarity between patches. In
 Fig. 3, the distance is expressed as the euclidean (geographic) distance between
@@
 other words, two communities which are close to each other will share a large
-proportion of their species, whereas two communities which are afar will share a
+proportion of their species pool, whereas two communities which are afar will share a
 small proportion.  In the species-sorting model, the relationship between
@@
 
-To a vast extent, these relationships are preseved when looking at the
-topological distances (Fiig. 4), i.e. along how many edges should one travel to connect
-two patchs. Interestingly enough, the distance-dissimilarity relationship for
-the neutral model is markedly hump-shaped, with sites being at a medium distance
-having the maximal dissimilarity.
+To a vast extent, these relationships are preserved when looking at the
+topological distances (Fiig. 4), i.e. along how many edges should one travel to
+connect two patchs. Interestingly enough, the distance-dissimilarity
+relationship for the neutral model is markedly hump-shaped, with sites being at
+a medium distance having the maximal dissimilarity.
 
@@
 %TP
-How to add other types of interactions
-	- Wootton: variation arises from unaccounted interactions
+Finally, our analyses emphasize the need to expand on the canonial neutral
+theory. As pointed out by Wootoon2005Nature433, most of the unexplained
+deviation of empirical community from the prediction of accurately calibrated
+neutral models can be attirbuted to non-competitive interactions. Canard2012
+proposed that neutral processes can explain the network structure of trophic
+interactions with a good accuracy. Incorporating reasonable complexity in the
+mechanisms adressed by neutral models is not a theoreticians' exercice: it will
+re-inforce the usefulness of the neutral theory as an operational concept,
+specifically one that can be used to derive baseline predictions about (i) the
+expected local species richness, and (ii) the expected species pool
+dissimilarity at the between-site and regional scales. These predictions are the
+benchmark against which empirical relev\'es of species richness and community
+structure ought to be compared, and coming up with realistic parameters to
+calibrate these models calls for a closer cooperation and dialogue between
+theoreticians and empiricists.
 

### 2013-05-27 ms.tex
Update ms.tex

@@
 
-\textbf{Title:}   Using neutral theory to reveal the contribution of dispersal to community assembly in complex landscales\\
+\textbf{Title:}   Using neutral theory to reveal the contribution of dispersal to community assembly in complex landscapes\\
 
@@
 
-Ecologists are required to move toward a predictive ecology, integrating
-elements of theoretical ecology \parencite{Thuiller2013}, and the metacommunity
-perspective appears naturally as the appropriate conceptual framework to develop
-the new modeling techniques required to fill this challenge. The metacommunity
+Ecology needs to move toward a more predictive approach. Integrating
+elements of theoretical ecology \parencite{Thuiller2013} with the metacommunity
+perspective appears naturally as a good candidate to fill this challenge. The metacommunity
 concept builds on feedbacks between local scale processes, such as competitive
@@
 
-The development of neutral theory has been quite provocative in that respect, as
-one could see it as a step back in time. Neutral theory makes the provocative
-assumption that species are ecologically equivalent and thereby any variation in
-the environment has no impact on demography \parencite{Bell2000,Hubbell2001}.
-Only demographic stochasticity and dispersal drive the structure of neutral
-ecological communities. It therefore appears that, on first sight, neutral
-theory is useless.  We will develop in this paper the argument that neutral
-theory could be a useful tool to understand the impact of dispersal on community
-organization in landscapes of various complexities.
-
-Neutral theory sparked an historical debate that is still lasting after more
+At first sight, the development of neutral theory appears as a step back. Based on Kimura's neutral theory of molecular evolution, the neutral theory of biodiversity makes the provocative
+assumption that species are ecologically equivalent. Thus, variation in
+the environment has no impact on demography \parencite{Bell2000,Hubbell2001}. Neutral
+ecological communities are driven only by demographic stochasticity and dispersal. Neutral theory sparked an historical debate still lasting after more
 than a decade \parencite{Chave2004, Etiennee2011, Rosindell2012,Clark2012}. It
-was stimulated by the surprising ability of neutral models to fit several well
+was stimulated by the surprising ability of neutral models to fit some well
 studied empirical observations such as species abundance distributions and
-distance-decay relationships. A remarkable strenght of neutral theory is to
+distance-decay relationships. A remarkable strength of the theory is to
 provide a \emph{"formal general theory of abundance and diversity that will
@@
 ecologists have documented"} \parencite{Bell2001}. Even if new studies rejecting
-neutral theory are consistently published (e.g.  \textcite{Ricklefs2012}), there
-is now almost a consensus that neutral theory is a well-developed null
+neutral theory are consistently published (e.g.  \textcite{Ricklefs2012}), a consensus is forming that neutral theory is a well-developed null
 hypothesis for niche theory and could even be used as an adequate approximation
 of ecological dynamics in some situations.  \textcite{Bell2001} nicely
-envisionned two perspectives to neutral theory that are still standing today.
+envisioned two perspectives to neutral theory that are still standing today.
 Under the weak perspective, neutral theory provides a set of realistic
@@
 between realism and instrumentalism \parencite{Wennekes2011}. Because every
-ecological model is a simplification of reality, any scientist has to
-subjectively decide the level of details he puts in, leaving out some elements
-judged unimportant. The realism perspective requires that all assumptions of
-theory to be true, while the utility of the theory is more important to
+ecological model is a simplification of reality, scientists have to
+subjectively decide the level of details they put in, leaving out some elements they consider unimportant. The realism perspective requires that all assumptions of
+theory must be true, while the utility of the theory is more important to
 instrumentalism. The utilitarian value of a theory could either be for
@@
 The instrumentalist view of neutral theory raises the question of why it should
-be a satisfying approximation despite knowing the pieces are wrong? We see two
-potential answers to this question. A first answer might be that stochasticity
-of various origins can blur the deterministic differences among species and
-promote ecological drift \parencite{Gravel2011}. Much has been said the
-existence of demographic stochasticity, some ecologists even arguing that
-neutral models impede progress in community ecology by hidding niche differences
-\parencite{Clark2012}, and we therefore will let this discussion for other
-papers. The second answer is that dispersal and historical contingencies might
+be a satisfying approximation despite knowing the pieces are wrong? Perhaps stochasticity
+of various origins blur the deterministic differences among species and
+promote ecological drift \parencite{Gravel2011}. A second answer is that dispersal and historical contingencies might
 have a much more profound impact on species distribution \parencite{Bahn2007,
@@
 
-In this paper, we argue that neutral theory is useful both to understand and
-predict the impact of dispersal on community organization. Even for purely
+In this paper, we argue that neutral
+theory can be a useful tool to understand the impact of dispersal on community
+organization in landscapes of various complexities. Even for purely
 theoretical analyses, we need a benchmark without niche differences to reveal
-the role of dispersal in structuring communities, and understand how it
+the role of dispersal in structuring communities and understand how it
 interacts with niche differentiation. We will explore recent applications of
 neutral theory, at the crossroad of network theory, to better represent the
-impact of landscape structure on biodiversitity distribution. This analysis will
+impact of landscape structure on biodiversity distribution. This analysis will
 prove particularly relevant to limnology, where most riverine and lacustre
-habitats are characterized by a their discrete nature and spatially complex
-arrangements \parencite{Peterson2013}. We will also reveal the relative
+habitats are characterized by their discrete nature and spatially complex
+arrangements \parencite{Peterson2013}. We will also explore the relative
 contribution of ecological interaction and niche differentiation by contrasting
@@
 
-Our main objective in this paper is to use neutral theory to stress the
+Our main objective is to use neutral theory to stress the
 importance of landscape network structure on the distribution of diversity. We
@@
 mechanisms at play. We will therefore move from a perspective where dispersal is
-either global or spatially explicity (e.g. over a lattice), and spatial
+either global or spatially explicit (e.g. over a lattice), and spatial
 constant, to a perspective focusing on the variance of dispersal. A second
@@
 Economo2011,Desjardins2012a,Desjardins2012b}, and even experiments
-\parencite{Altermat2012}, recently introduced more realistic lanscapes and found
-surprising contributions of spatial contingencies. We will start with a short
-review of the main approaches to describe spatial networks and the studies
-investigating them. Then we will describe three simple toy models of
-metacommunity dynamics, taking this opportunity to review their assumptions and
+\parencite{Altermat2012}, recently introduced more realistic landscapes and found
+surprising contributions of spatial contingencies. We start with a short
+review of the main approaches to describe spatial networks. Then we describe three simple toy models of
+metacommunity dynamics, using this opportunity to review their assumptions and
 main predictions. We provide as Supplementary Material the R scripts for the toy
@@
 length of the shortest path \cite{dij59}. For example, two lakes could be
-very close on a map (short Euclidean distance) but if they are not directly
-linked by a river the geodesic distance could be great.
+very close on a map (short Euclidean distance) but the geodesic distance could be great if they are not directly
+linked by a river.
+
+The popularity of network theory stem for their ability to model complex
+structures while allowing us to extract useful metrics (Table 1). At a very high level, 
+a network can be described by a number of vertices (the order) and a number
+of edges (the size). Looking more closely, the relationship between vertices
+is influenced by paths, which are ordered series of nodes.
+Centrality is a central (no pun intended) concept in network theory, where it
+can be seen to as a measure of ``importance''. The simplest measure
+of the centrality of a node is its degree, which is the number of nodes directly connected to it. Of course, this is a very rough description of centrality. For
+example, two lakes can have the same degree, with one being connected to
+a small isolated cluster, while the other one is part of one of the biggest 
+network of lake. In this case, measures of centrality like eigen-centrality will
+weight the importance of the connection so a node connected to well-connected
+nodes will have higher centrality than a node connected to isolated nodes.
 
@@
 tree.  It builds a tree from the shortest path tree \cite{dij59} of a vertex in
-a random geometric network. This random geometric tree does not exaxctly
+a random geometric network. This random geometric tree does not exactly
 represent dendritic landscapes but is a convenient model to simulate lake
@@
 speciation \cite{des12,des12b}.
-
-% TO PHIL: IT WOULD BE GREAT IF YOU COULD EXPAND ONE OR TWO PARAGRAPHS ON THE 
-% MAIN STATISTICS USED TO DESCRIBE SPATIAL GRAPHS. AT LEAST PRESENT THE MAIN 
-% DIFFERENCE BETWEEN NETWORK AND VERTEX LEVEL METRICS, AND THEN REFER TO THE 
-% TABLE.
 

### 2013-05-27 ms.tex
Discussion

@@
 metacommunity models and illustrate how the implementation of more realistic
-landsapes could reveal the importance of dispersal on community structure. We
+landscapes could reveal the importance of dispersal on community structure. We
 argued in the introduction that neutral theory is useful both to understand and
@@
 competitive interactions buffer the potential impact of landscape structure. The
-strongest centrality-species richness relationship was observerd for the patch
+strongest centrality-species richness relationship was observed for the patch
 dynamics, a model without any interactions. On the other hand, strong and
@@
 landscapes. So far we have treated only theoretical models, but we could also
-envision to parameterize them and simulate real landscapes. The recruitment
-probabilities defined above could be used as satistical models (likelihood
+envision to parametrize them and simulate real landscapes. The recruitment
+probabilities defined above could be used as statistical models (likelihood
 functions) to fit to empirical data. Prior information could be used to define
 apriori dispersal kernels and then fit the model as in \textcite{Gravel2008}.
-The fit of metapopulation models to spatially explicit landscapes was pionneered
+The fit of metapopulation models to spatially explicit landscapes was pioneered
 by \textcite{Hanski1998} and recently extended to species distribution models
 including both species sorting and dispersal limitations
-\parencite{Boulangeat2012}. Given the parameterization, one could run neutral
+\parencite{Boulangeat2012}. Given the parametrization, one could run neutral
 models to generate null hypotheses that could be eventually compared to observed
@@
 The multivariate variance partitioning framework originally proposed by Borcard
-et al. (1992) and further developped by Borcard and Legendre (2002) has been
+et al. (1992) and further developed by Borcard and Legendre (2002) has been
 widely used to quantify the relative importance of species sorting and dispersal
@@
 model species distribution as a function of environmental variables, taking into
-accound the spatial autocorrelation of species distriution (Leduc1992,
+account the spatial autocorrelation of species distribution (Leduc1992,
 Borcard1992, Legendre1993). This methodology has been widely used over the last
@@
 expectations based on different hypotheses and then compare them. But most of
-all, parameterized spatially explicit neutral models could be more useful if
+all, parametrized spatially explicit neutral models could be more useful if
 used to generate expectations. For instance, neutral models could be used to
 predict the consequences of habitat destruction, fragmentation or a change in
-the connectinity matrix \parencite{Hubbell2008}. The spatially explicit
+the connectivity matrix \parencite{Hubbell2008}. The spatially explicit
 description of the landscape is a major improvement toward that end, providing
@@
 %PDP
-Working with more realistic landscapes
+Working with more complex representations on landscapes has several
+advantages. Real landscapes are not flat geometric objects, they are highly
+structured and diverse. This diverse structure has long been recognized as
+a key component of diversity.
+In the 19th century, Moritz Wagner noted that patterns of rivers \cite{coy04} could explain how beetles diversified.
+Yet, to this day, the relationship between spatial structures and biodiversity is not resolved, but networks provide powerful tools to analyze landscapes and generate
+testable predictions on the relationship between community assembly
+and spatial structures. For example, the neutral theory predict less diversity
+in isolated communities and constant speciation (regardless of isolation) \cite{eco08}. Adding the effect of gene flow changes the predictions, but
+the neutral theory do not predict adaptive radiation and might thus face
+problems for predicting many long-term ecological processes.
 
@@
 deviation of empirical community from the prediction of accurately calibrated
-neutral models can be attirbuted to non-competitive interactions. Canard2012
+neutral models can be attributed to non-competitive interactions. Canard2012
 proposed that neutral processes can explain the network structure of trophic
 interactions with a good accuracy. Incorporating reasonable complexity in the
-mechanisms adressed by neutral models is not a theoreticians' exercice: it will
-re-inforce the usefulness of the neutral theory as an operational concept,
+mechanisms addressed by neutral models is not a theoreticians' exercise: it will
+re-enforce the usefulness of the neutral theory as an operational concept,
 specifically one that can be used to derive baseline predictions about (i) the

### 2013-05-27 ms.tex
First complete reading

@@
 % TODO:
-% TP, PDP & DG: Complete the discussion
-% PDP: Additional paragraph on the metrics
-% DG: Figure legends
-% DG: Abstract
+% DG: First complete reading
+% DG: Figures
+% DG: Legends
 % DG: Compile references
 % TP & PDP: fix bugs I generated!!!
+% DG: Abstract
 
@@
 
-Ecology needs to move toward a more predictive approach. Integrating
-elements of theoretical ecology \parencite{Thuiller2013} with the metacommunity
-perspective appears naturally as a good candidate to fill this challenge. The metacommunity
-concept builds on feedbacks between local scale processes, such as competitive
-interactions and local adaptation, and regional scale processes such as
-dispersal, gene flow and speciation \textcite{Leibold2004a}. This perspective is
-particularly relevant to limnology, where exchanges of organisms and nutrients
-affect community and ecosystem properties from the local (e.g. vertical mixing
-\parencite{Ryabov2011}) to the regional (e.g. connection of lakes
-\parencite{Leibold2004b} scales. It emphasizes the importance of dispersal
-relative to pairwise interactions in the organization of ecological communities.
+Ecology needs to move toward a more predictive approach, integrating elements of
+theoretical ecology \parencite{Thuiller2013}. The metacommunity perspective \parencite{Leibold2004a}
+appears naturally as the appropriate conceptual framework to fill this
+challenge. The metacommunity concept builds on feedbacks between local scale
+processes, such as competitive interactions and local adaptation, and regional
+scale processes such as dispersal, gene flow and speciation. It is particularly relevant to limnology,
+where exchanges of organisms and nutrients affect community and ecosystem
+properties from the local (e.g. vertical mixing \parencite{Ryabov2011}) to the
+regional (e.g. connection of lakes \parencite{Leibold2004b, Gravel2010} scales.
+It emphasizes the importance of dispersal relative to pairwise interactions in
+the organization of ecological communities.
 
-At first sight, the development of neutral theory appears as a step back. Based on Kimura's neutral theory of molecular evolution, the neutral theory of biodiversity makes the provocative
-assumption that species are ecologically equivalent. Thus, variation in
-the environment has no impact on demography \parencite{Bell2000,Hubbell2001}. Neutral
-ecological communities are driven only by demographic stochasticity and dispersal. Neutral theory sparked an historical debate still lasting after more
-than a decade \parencite{Chave2004, Etiennee2011, Rosindell2012,Clark2012}. It
-was stimulated by the surprising ability of neutral models to fit some well
-studied empirical observations such as species abundance distributions and
-distance-decay relationships. A remarkable strength of the theory is to
-provide a \emph{"formal general theory of abundance and diversity that will
-account, in a simple and economical fashion, for the many patterns that
-ecologists have documented"} \parencite{Bell2001}. Even if new studies rejecting
-neutral theory are consistently published (e.g.  \textcite{Ricklefs2012}), a consensus is forming that neutral theory is a well-developed null
-hypothesis for niche theory and could even be used as an adequate approximation
-of ecological dynamics in some situations.  \textcite{Bell2001} nicely
-envisioned two perspectives to neutral theory that are still standing today.
-Under the weak perspective, neutral theory provides a set of realistic
-predictions of community organization despite false assumptions. Even if being
-fundamentally wrong, neutral theory is still useful when used as a null
-hypothesis \parencite{Gotelli2006}. It is considered as an improvement over
-traditional null hypotheses based on randomization \parencite{Gotelli2000}
-because it readily integrates dispersal. The strong version on the other hand
-posits that neutral theory is a satisfying approximation to community dynamics
-and an appropriate theory to explain the distribution of biodiversity. It
-implies that the right mechanisms have been identified and that the consistently
-observed differences among species do not impact community organization.
+At first sight, the development of neutral theory appears as a step back.
+Neutral theory of biodiversity makes the provocative assumption that species are
+ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral ecological communities are
+driven only by demographic stochasticity and dispersa and thus, variation in the environment has no impact on
+demography. Neutral theory sparked
+an historical debate still lasting after more than a decade
+\parencite{Chave2004, Etiennee2011, Rosindell2012,Clark2012}. It was stimulated
+by the surprising ability of neutral models to fit some well studied empirical
+observations such as species abundance distributions and distance-decay
+relationships.
+
+A remarkable strength of the theory is to provide a \emph{"formal general theory
+of abundance and diversity that will account, in a simple and economical
+fashion, for the many patterns that ecologists have documented"}
+\parencite{Bell2001}. Even if new studies rejecting neutral theory are
+consistently published (e.g. \textcite{Ricklefs2012}), a consensus is forming
+that neutral theory is a well-developed null hypothesis for niche theory and
+could even be used as an adequate approximation of ecological dynamics in some
+situations. \textcite{Bell2001} nicely envisioned two perspectives to neutral
+theory that are still standing today. Under the weak perspective, neutral theory
+provides a set of realistic predictions of community organization despite false
+assumptions. Even if being fundamentally wrong, neutral theory would still
+useful when used as a null hypothesis \parencite{Gotelli2006}. It is considered
+as an improvement over traditional null hypotheses based on randomization
+\parencite{Gotelli2000} because it readily integrates dispersal. On the other
+had, the strong version posits that neutral theory is a satisfying approximation
+to community dynamics and an appropriate theory to explain the distribution of
+biodiversity. It implies that the right mechanisms have been identified and that
+the consistently observed differences among species do not impact community
+organization.
 
@@
 between realism and instrumentalism \parencite{Wennekes2011}. Because every
-ecological model is a simplification of reality, scientists have to
-subjectively decide the level of details they put in, leaving out some elements they consider unimportant. The realism perspective requires that all assumptions of
-theory must be true, while the utility of the theory is more important to
+ecological model is a simplification of reality, scientists have to subjectively
+decide the level of details they put in, leaving out some elements they consider
+unimportant. The realism perspective requires that all assumptions of the theory
+must be true, while the utility of the theory is more important to
 instrumentalism. The utilitarian value of a theory could either be for
-understanding or prediction (another old philosophical debate, see
+understanding or for prediction (another old philosophical debate, see
 \textcite{Schmueli2010}). Obviously neutral theory could only be instrumental.
@@
 The instrumentalist view of neutral theory raises the question of why it should
-be a satisfying approximation despite knowing the pieces are wrong? Perhaps stochasticity
-of various origins blur the deterministic differences among species and
-promote ecological drift \parencite{Gravel2011}. A second answer is that dispersal and historical contingencies might
-have a much more profound impact on species distribution \parencite{Bahn2007,
-Boulangeat2012} and ecological dynamics. The debate over the equivalence
-assumption and demographic stochasticity has perhaps overlook the recognition of
-how much dispersal influence community assembly.
+be a satisfying approximation despite knowing the pieces are wrong? Perhaps
+stochasticity of various origins blur the deterministic differences among
+species and promote ecological drift \parencite{Gravel2011a}. Much has been said
+the existence of demographic stochasticity \parencite{Clark2012}, and we
+therefore will let this discussion for other papers. A second answer is that
+dispersal and historical contingencies might have a much more profound impact on
+species distribution \parencite{Bahn2007, Boulangeat2012} and ecological
+dynamics. The debate over the equivalence assumption and demographic
+stochasticity has perhaps overlook the recognition of how much dispersal
+influence community assembly.
 
-In this paper, we argue that neutral
-theory can be a useful tool to understand the impact of dispersal on community
-organization in landscapes of various complexities. Even for purely
-theoretical analyses, we need a benchmark without niche differences to reveal
-the role of dispersal in structuring communities and understand how it
-interacts with niche differentiation. We will explore recent applications of
-neutral theory, at the crossroad of network theory, to better represent the
-impact of landscape structure on biodiversity distribution. This analysis will
-prove particularly relevant to limnology, where most riverine and lacustre
-habitats are characterized by their discrete nature and spatially complex
-arrangements \parencite{Peterson2013}. We will also explore the relative
+In this paper, we argue that neutral theory can be a useful tool to understand
+the impact of dispersal on community organization in landscapes of various
+complexities. Even for purely theoretical analyses, we need a benchmark without
+niche differences to reveal the role of dispersal in structuring communities and
+understand how it interacts with niche differentiation. We will explore recent
+applications of neutral theory, at the crossroad of network theory, to better
+represent the impact of landscape structure on biodiversity distribution. This
+analysis will prove particularly relevant to limnology, where most riverine and
+lacustre habitats are characterized by their discrete nature and spatially
+complex arrangements \parencite{Peterson2013}. We will also explore the relative
 contribution of ecological interaction and niche differentiation by contrasting
@@
 
-Our main objective is to use neutral theory to stress the
-importance of landscape network structure on the distribution of diversity. We
-refer to the landscape organization as a "spatial contingency"
-\parencite{Peres-Neto2013} that could potentially affect the coexistence
-mechanisms at play. We will therefore move from a perspective where dispersal is
-either global or spatially explicit (e.g. over a lattice), and spatial
-constant, to a perspective focusing on the variance of dispersal. A second
-generation of neutral models (e.g. \textcite{Economo2008,
-Economo2011,Desjardins2012a,Desjardins2012b}, and even experiments
-\parencite{Altermat2012}, recently introduced more realistic landscapes and found
-surprising contributions of spatial contingencies. We start with a short
-review of the main approaches to describe spatial networks. Then we describe three simple toy models of
-metacommunity dynamics, using this opportunity to review their assumptions and
-main predictions. We provide as Supplementary Material the R scripts for the toy
-models and all simulations conducted for this paper. We then conduct simple
-simulations of these models to reveal with simple examples the impact of spatial
-network structure on diversity distribution. We conclude with a discussion on
-the operationally of the framework.
+Our main objective is to use neutral theory to stress the importance of
+landscape network structure on the distribution of diversity. We refer to the
+landscape organization as a "spatial contingency" \parencite{Peres-Neto2013}
+that could potentially affect the coexistence mechanisms at play. We will
+therefore move from a perspective where dispersal is either global or spatially
+explicit (e.g. over a lattice), and spatial constant, to a perspective focusing
+on the variance of dispersal. A second generation of neutral models (e.g.
+\textcite{Economo2008, Economo2011,Desjardins2012a,Desjardins2012b}, and even
+experiments \parencite{Altermat2012}, recently introduced more realistic
+landscapes and found surprising contributions of spatial contingencies. We start
+with a short review of the main approaches to describe spatial networks. Then we
+describe three simple toy models of metacommunity dynamics, using this
+opportunity to review their assumptions and main predictions. We provide as
+Supplementary Material the R scripts for the toy models and all simulations
+conducted for this paper. We then conduct simple simulations of these models to
+reveal with simple examples the impact of spatial network structure on diversity
+distribution. We conclude with a discussion on the operationally of the
+framework.
 
@@
 
-A network is a discrete mathematical object made of two sets: a set of
-vertices (or nodes) $V$ and a set of edges $E$ connecting the vertices
-\cite{new10}. The term ``graph'' is often preferred in computer science and
-mathematics \cite{gro06}, with graph algorithms being an important and active
-area of research \cite{sed01}. A network is a combinatorial object: it is
-used to study how discrete entities are connected and how they combine
-together to create complex structures. They are used to study molecules, food
-webs, social networks, or even the relationship between variables in
-statistics \cite{wri21,new10}. We are especially interested in spatial
-networks, a special kind of network mixing the combinatorial properties of
-networks with a topological space \cite{kob94}. Thus, the vertices in a
-spatial graph are embedded in some other space, most often the two or three-
-dimensional Euclidean space. This object brings a rich representation to
-spatial ecology and is particularly suited for systems of lakes and rivers, which can
-easily be represented by vertices and edges. There are two notions of
-distance in spatial networks. Euclidean distance represents the geographical
-distance between the vertices $(i, j)$, i.e.: $\sqrt{(x_i - x_j)^2 + (y_i - y_
-j)^2}$. Geodesic distance is the distance in the graph space, i.e.: the
-length of the shortest path \cite{dij59}. For example, two lakes could be
-very close on a map (short Euclidean distance) but the geodesic distance could be great if they are not directly
-linked by a river.
+A network is a discrete mathematical object made of two sets: a set of nodes
+(or vertices) and a set of edges connecting the nodes \cite{new10}. The
+term ``graph'' is often preferred in computer science and mathematics
+\cite{gro06}, with graph algorithms being an important and active area of
+research \cite{sed01}. A network is a combinatorial object: it is used to study
+how discrete entities are connected and how they combine together to create
+complex structures. They are used to study molecules, food webs, social
+networks, or even the relationship between variables in statistics
+\cite{wri21,new10}. We are especially interested in spatial networks, a special
+kind of network mixing the combinatorial properties of networks with a
+topological space \cite{kob94}. Thus, the vertices in a spatial graph are
+embedded in some other space, most often the two or three- dimensional Euclidean
+space. This object brings a rich representation to spatial ecology and is
+particularly suited for systems of lakes and rivers, which can easily be
+represented by vertices and edges. There are two notions of distance in spatial
+networks. Euclidean distance represents the geographical distance between the
+nodes $(i, j)$, i.e.: $\sqrt{(x_i - x_j)^2 + (y_i - y_ j)^2}$. Geodesic
+distance is the distance in the graph space, i.e.: the length of the shortest
+path \cite{dij59}. For example, two lakes could be very close on a map (short
+Euclidean distance) but the geodesic distance could be great if they are not
+directly linked by a river.
 
-The popularity of network theory stem for their ability to model complex
-structures while allowing us to extract useful metrics (Table 1). At a very high level, 
-a network can be described by a number of vertices (the order) and a number
-of edges (the size). Looking more closely, the relationship between vertices
-is influenced by paths, which are ordered series of nodes.
-Centrality is a central (no pun intended) concept in network theory, where it
-can be seen to as a measure of ``importance''. The simplest measure
-of the centrality of a node is its degree, which is the number of nodes directly connected to it. Of course, this is a very rough description of centrality. For
-example, two lakes can have the same degree, with one being connected to
-a small isolated cluster, while the other one is part of one of the biggest 
-network of lake. In this case, measures of centrality like eigen-centrality will
-weight the importance of the connection so a node connected to well-connected
-nodes will have higher centrality than a node connected to isolated nodes.
+The popularity of network theory stems for its ability to model complex
+structures while allowing us to extract useful metrics (Table 1). At a very high
+level, a network can be described by its number of nodes (the order) and 
+edges (the size). Looking more closely, the relationship between
+nodes is influenced by paths, which are ordered series of nodes. Centrality
+is a \emph{central} concept in network theory, where it can be seen
+to as a measure of ``importance''. The simplest measure of the centrality of a
+node is its degree, which is the number of nodes directly connected to it. Of
+course, this is a very rough description of centrality. For example, two lakes
+can have the same degree, with one being connected to a small isolated cluster,
+while the other one is part of one of the biggest network of lake. In this case,
+measures of centrality like eigen-centrality will weight the importance of the
+connection, so a node connected to well-connected nodes will have higher
+centrality than a node connected to isolated nodes.
 
 For simulations, spatial networks can easily be generated with the random
-geometric graph algorithm \cite{sed01}. In this algorithm, all vertices are
+geometric graph algorithm \cite{sed01}. In this algorithm, all nodes are
 assigned to a position in some two-dimensional space, most often the unit
-square. Then, all pairs of vertices within some threshold Euclidean distance $r$
-are connected with an edge. The resulting networks have the desirable
-property of locality: if a vertex $A$ is connected to two vertices $B$ and $C$.
-then $B$ and $C$ are more likely to be connected than two random vertices.
-Random geometric networks have been extensively studied \cite{app97a,app97
+square. Then, all pairs of nodes within some threshold Euclidean distance $r$
+are connected with an edge. The resulting networks have the desirable property
+of locality: if a node $A$ is connected to two vertices $B$ and $C$. then $B$
+and $C$ are more likely to be connected than two random vertices. Random
+geometric networks have been extensively studied \cite{app97a,app97
 b,app02a,app02b,pen03} and we provide a R function to generate them. The
-position of vertices is typically random, but we could also imagine alterations
-where they are either more aggregated or segregated than expected by chance alone.
+position of nodes is typically random, but we could also imagine alterations
+where they are either more aggregated or segregated than expected by chance
+alone.
 
 We also provide the code for a second structure that we call a random geometric
-tree.  It builds a tree from the shortest path tree \cite{dij59} of a vertex in
-a random geometric network. This random geometric tree does not exactly
-represent dendritic landscapes but is a convenient model to simulate lake
-connected by rivers to a series of smaller lakes.
+tree. The algorithm first builds a random geometric graph, then select a node
+from which to start the tree. It then calculaes the the shortest path tree
+\cite{dij59} from this node to all other ones. Edges that are not located along
+this tree are pruned. This random geometric tree does not exactly represent
+dendritic landscapes but is a convenient model to simulate lake connected by
+rivers to a series of smaller lakes.
 
@@
 
-% Table 1: main descriptors of spatial networks. Column 1: Name of the metric; Column 2: Definition
-% We might want to distinguish network level versus node-level metrics.
-Concepts
-Path A sequence of edges forming a sequence of vertices.
-Connection Two vertices are connected iff there is a path between them.
-Geodesic distance Length of the shortest path between two vertices.
-Network-level metrics
-Order Total number of vertices.
-Size Total number of edges.
-Connectivity A measure of robustness: the minimum number of elements to remove
-to isolate the vertices.
-Components The number of connected subsets.
-Vertex-level metrics
-Degree The number of edges of a vertex.
-Closeness centrality Average geodesic distance between a vertex and all other vertex.
-Eigenvector centrality A measure of centrality based on the concept that connection to
-highly connected vertices are more important.
-Betweenness centrality The number of shortest paths from all vertices to all others that
-pass through that vertex.
+\begin{table*}
+	\centering
+	\begin{tabular}{p{5cm}p{8cm}}
+		\hline
+		\textbf{Concept} & \textbf{Definition} \\
+		\hline
+		Path & A sequence of edges forming a sequence of nodes \\
+		Connection & Two nodes are connected if there is a path between them \\
+		Euclidean distance & Geographical distance between two nodes \\
+		Geodesic distance & Length of the shortest path between two nodes \\
+
+		\textbf{Network-level metrics} & \\
+		Order & Total number of nodes \\
+		Size & Total number of edges \\
+		Connectivity & A measure of robustness: the minimum number of elements to remove to isolate the nodes \\
+		Components &  The number of connected subsets\\
+
+		\textbf{Node-level metrics} & \\
+		Degree & The number of edges of a node \\
+		Closeness centrality & Average geodesic distance between a node and all other vertex \\
+		Eigenvector centrality & A measure of centrality based on the concept that connection to
+		highly connected nodes are more important \\
+		Betweenness centrality & The number of shortest paths from all nodes to all others that
+		pass through that note \\
+		\hline
+	\end{tabular}
+	\caption{Main descriptors of spatial networks used in this study.}
+\end{table*}
 
@@
 perspectives of metacommunity ecology: patch dynamics, neutral dynamics and
-species sorting. While the neutral model is interesting in itself, it is by
+species sorting. While the neutral model is interesting in itself, it is by its
 comparison with a model without any interactions (patch dynamics) and with niche
@@
 model \parencite(Levins1969) describes the stochastic colonizations and
-extinctions of population over a homogenous landscape. The basic unit is the
-population. The Levins model tracts the dynamics of occupancy (the fraction of
-the landscape that is occupied) with an ordinary differential equation and
-therefore assumes an infinite landscape. The simulation model we run is more
-realistic as it simulates a finite number $N$ of discrete patches. The rules
-described in the previous section were used to generate connectivity matrices
-along four scenarios (Fig. 1): global dispersal (connected graph), a lattice, a
-random geometric graph and a random tree graph. A patch $x$ shares $d_x$ links
-with neigbhouring patches (its degree). At each time step (the simulation model
-is discrete in time), the probability that a colonist coming from an occupied
-patch $y$ arrives at patch $x$ is $cd_y^{-1}$, where $c$ is the probability a
-colonization event takes place if all connected patches are occupied. The
-expected probability that a colonist arrives to patch $x$ from patch $y$ is then
-$C_{ixy}=cp_{iy}d_y^{-1}$, where $p_{i}y$ is the probability that patch $y$ is
-occupied by species $i$. The probability that an extinction occurs in a given
-patch is $e$. The Levins model is for a single species, but a basic
-metacommunity patch dynamics model could be run by aggregating $S$ independent
-metapopulation models \parencite(Hanski1997). There are no interactions in this
-simple model, which means there is no limit to local species richness and no
-carrying capacity. Competitive, mutualistic and predator-prey interactions have
-been added to this framework (e.g.
-\textcite{Tilman1994,Klausmeier1998,Holt1996}) but we will keep this model
+extinctions of a single species over a homogenous landscape. The basic unit is
+the population. The model tracts the dynamics of occupancy (the fraction of the
+landscape that is occupied) with an ordinary differential equation and therefore
+assumes an infinite landscape. The simulation model we run is more realistic as
+it simulates a finite number $N$ of discrete patches (or nodes in network
+terminology). The rules described in the previous section were used to generate
+connectivity matrices along four scenarios (Fig. 1): global dispersal (connected
+graph), a lattice, a random geometric graph and a random tree graph. A patch $x$
+shares $d_x$ links with neigbhouring patches (its degree). At each time step
+(the simulation model is discrete in time), the probability that a colonist
+coming from an occupied patch $y$ arrives at patch $x$ is $cd_y^{-1}$, where $c$
+is the probability a colonization event takes place if all connected patches are
+occupied. The expected probability that a colonist arrives to patch $x$ from
+patch $y$ is then $C_{ixy}=cp_{iy}d_y^{-1}$, where $p_{i}y$ is the probability
+that patch $y$ is occupied by species $i$. The probability that an extinction
+occurs in a given patch is $e$. The Levins model is for a single species, but a
+basic metacommunity patch dynamics model could be run by aggregating $S$
+independent metapopulation models \parencite(Hanski1997). There are no
+interactions in this simple model, which means there is no limit to local
+species richness and no carrying capacity. Competitive, mutualistic and
+predator-prey interactions have been added to this framework (e.g.
+\textcite{Tilman1994,Holt1996,Klausmeier1998,Gravel2011b}) but we will keep this model
 minimal for the sake of comparison with the neutral model.
@@
 Given that all species are the same, then we should expect the regional
-diversity to be $S$ if this condition is satisfied and $0$ if not. The situation
+diversity ($\gamma$) to be $S$ if this condition is satisfied and $0$ if not. The situation
 is however more complex in spatially explicit landscapes with complex
@@
 Neutral theory introduces strong competitive interactions by assuming there is a
-finite number of individuals that could occupy a  patch. There are different
-ways to simulate this \emph{zero-sum rule} \parencite{Bell2000,Hubbell2001}, but
-they all result in the same constraint that the increase in abundance of a
-species could only occur after an equivalent decrease by another species. One
-important change in the formulation of most neutral models relative the patch
-dynamics model presented above is therefore that it is individual-based, not
-population based. We therefore considered in our toy model of neutral dynamics
-that each local patch holds $J_x$ individuals. The model tracts the local
-abundance of all species $N_{ix}$ in each local patch. At each time step an
-individidual dies with probability $k$. Recruitment only occurs in vacant sites,
-similarly to a tree by tree replacement process in a closed canopy forest.
+finite number of individuals that could occupy a patch. There are different ways
+to simulate this \emph{zero-sum rule} \parencite{Bell2000,Hubbell2001}, but they
+all result in the same constraint that the increase in abundance of a species
+could only occur after an equivalent decrease by another species. One important
+change in the formulation of most neutral models relative the patch dynamics
+model presented above is therefore that it is individual-based, not population
+based. We therefore considered in our toy model of neutral dynamics that each
+local patch holds $J_x$ individuals. The model tracts the local abundance of all
+species $N_{ix}$ in each local patch. At each time step an individidual dies
+with probability $k$. Recruitment only occurs in vacant sites, similarly to a
+tree by tree replacement process in a closed canopy forest.
 
@@
 neutral models, making possible the coupling with the metacommunity and
-neighbouring patches. We adopt a simple formulation in our model based on
-\parencite{Gravel2006}. The approach is conceptually similar to placing a seed
-trap in a canopy gap and picking a seed at random among the ones falling
-in to determine the identity of the recruited species. The composition of the
-seed pool in that trap will be a mixture of local dispersal and immigrants from
-the metacommunity. For simplicity, we consider three spatial scales of dispersal
-but it would be easy to generalize the approach to a continuous seed dispersal
+neighbouring patches. We adopt a simple formulation based on
+\parencite{Gravel2006}. The approach is conceptually similar to placing a trap
+in a canopy gap and picking a seed at random among the ones falling in to
+determine the identity of the recruited species. The composition of the seed
+pool in that trap will be a mixture of local dispersal and immigrants from the
+metacommunity. For simplicity, we consider three spatial scales of dispersal but
+it would be easy to generalize the approach to a continuous seed dispersal
 kernel \parencite{Gravel2006}. The parameter $m$ is the probability that the
@@
 connected nodes because some nodes will spread their seeds across a higher
-number of nodes and thus contribute less to the seed pool. We consider simply
-$P_{ix} =\frac{\sum d_y^{-1}P_{iy}}{\sum dy^{-1}}$.  We assume for simplicity
+number of nodes and thus contribute less to the seed pool. We thus consider
+$P_{ix} =\frac{\sum P_{iy}d_y^{-1}}{\sum d_y^{-1}}$. We assume for simplicity
 (and without loss of generality, \parencite{Bell2000}) that the relative
 abundance in the metacommunity is uniform, i.-e. equal to $S^{-1}$. This
-immigration prevents the collapse of the metacommunity because otherwise all
-species except one will face extinction by ecological drift (speciation prevents
-this phenomenon to occur in \textcite{Hubbell2001}). The local recruitment
-probability is consequently $Pr_{ix} = MS^{-1} + mP_{ix} +
+immigration prevents the collapse of the metacommunity to a single species,
+since otherwise all species except one will face extinction by ecological drift
+(speciation prevents this phenomenon to occur in \textcite{Hubbell2001}). The
+local recruitment probability is consequently $Pr_{ix} = MS^{-1} + mP_{ix} +
 (1-m-M)N_{ix}J_x^{-1}$.
 
-The model is neutral because it assumes that the probabilities of a local
-recruitment, an immigration and a mortality event are all equal across species. Demographic
-stochasticity is the source of variations in abundance, but larger disturbances
-could be simulated as well, as long as they hit all species with the same
-probability, independently of their density. The fundamental feature of neutral
-dynamics is therefore the ecological drift, defined as population changes
-emerging from neutraly stable population dynamics. It can be measured as the
-variance between replicated time series of community dynamics
-\parencite{Gravel2011}. \textcite{Hubbell2001} provides a very comprehensive
+The model is neutral because it assumes that the probabilities of local
+recruitment, immigration and mortality events are all equal across species.
+Demographic stochasticity is the source of variations in abundance, but larger
+disturbances could be simulated as well, as long as they hit all species with
+the same probability, independently of their abundance. The fundamental feature of
+neutral dynamics is therefore the ecological drift, defined as population
+changes emerging from neutraly stable population dynamics. It can be measured as
+the variance between replicated time series of community dynamics
+\parencite{Gravel2011a}. \textcite{Hubbell2001} provides a very comprehensive
 analysis of the model, with specific attention to the effect of the different
-parameters on the drift (and consequently variance in abundance) and time to
+parameters on drift (and consequently variance in abundance) and time to
 extinction. Despite its simplicity, the neutral model is surprisingly rich in
@@
 analyzed the performance of neutral models to predict species abundance
-distributions, range-abundance relationship, spatial variation in abundance,
-species-area relationship, community turnover (beta-diversity) and
-co-occurrence. Recent trophic neutral models also predicted realistic ecological networks \parenciteCanard2012}. Other than the ecological equivalence assumption, one of the most
-criticized aspect of neutral models is the realism of the speciation process and
-the required speciation rates to sustain species richness
-\parencite{Ricklefs2003,Etienne2007}. Recent neutral models with more credible
-speciation models \parencite{Rosindell2009,Desjardins2012a} revealed the
-difficulty to maintain diversity in neutral models over macro-evolutionary time
-scales. These models however also generated interesting novel predictions on
-endemic species richness \parencite{Rosindell2011,Desjardins2012b}.
+distributions, the range-abundance relationship, spatial variation in abundance,
+the species-area relationship, community turnover (beta-diversity) and
+co-occurrence. Recent trophic neutral models were also found to predict realistic ecological
+network structures \parencite{Canard2012}. Other than the ecological equivalence
+assumption, one of the most criticized aspect of neutral models is the realism
+of the speciation process and the required speciation rates to sustain species
+richness \parencite{Ricklefs2003,Etienne2007}. Recent neutral models with more
+credible speciation models \parencite{Rosindell2009,Desjardins2012a} revealed
+the difficulty to maintain diversity in neutral models over macro-evolutionary
+time scales. These models nonetheless proposed interesting predictions
+on endemic species richness and island biogeography \parencite{Rosindell2011,Desjardins2012b}.
 
@@
 lottery among potential candidates as in the neutral model. The recruitment probability 
-is however biased by species specific response to local environmental conditions.
+is however biased by species specific responses to local environmental conditions.
 
@@
 local environmental conditions could be implemented at this stage with a biased
-survival probability. The $J_x$ individuals all experience a unique environmental condition 
-$E_{nx}$ called a microsite $n$. We considered a patch average $\overline{E_x}$, with a
-within-patch variance $\sigma_x$. The regional average is $\overline{E_R}$ and
-the regional variance $\sigma_R$ (for simplicity we considered normal
-distributions, but different distributions will lead to different regional
-similarity constraints \parencite{Mouquet2003,Tilman2004,Gravel2006}). We
-consider that a fraction $\lambda_{inx}$ of offsprings reaching the microsite
-where recruitment occurs will survive. The recruitment probability is therefore
-biased in favour of the species with highest survival because only some species will be able to cope with the local environmental conditions. For tractability we define the relative abundance in
-the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$. The
-calculation of the relative abundance in the seed rain is the same as the
-neutral model but the recruitment probability differs because only a fraction of
-offspring survive. It is formulated as $Pr_{ix} =
+survival probability. The $J_x$ individuals all experience a unique
+environmental condition $E_{nx}$ called a microsite $n$. We considered a patch
+average $\overline{E_x}$, with a within-patch variance $\sigma_x$. The regional
+average is $\overline{E_R}$ and the regional variance $\sigma_R$ (for simplicity
+we considered normal distributions of environmental conditions, but different
+distributions will lead to different regional similarity constraints
+\parencite{Mouquet2003,Tilman2004,Gravel2006}). We consider that a fraction
+$\lambda_{inx}$ of offsprings reaching the microsite where recruitment occurs
+will survive. The recruitment probability is therefore biased in favour of the
+species with highest survival because only some species will be able to cope
+with the local environmental conditions. We define the relative
+abundance in the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} +
+(1-m-M)N_{ix}J_x^{-1}$. The calculation of the relative abundance in the seed
+rain is the same as the neutral model but the recruitment probability differs
+because only a fraction of offspring survive. It is formulated as $Pr_{ix} =
 \frac{\lambda_{inx}Z_{ix}}{\sum \lambda_{jnx}Z_{jx}}$. The function describing
@@
 Analyses of similar models with a combination of dispersal and species-sorting
-shown that predictions are extremely variables and depend on the distribution of
-environmental conditions, niche optimums and breadth. For instance, a
-well-studied prediction of neutral models is the species abundance distribution.
-It was shown that niche models can predict similar distributions given appropriate parameters
-\parencite{Tilman2004,Gravel2006}. The main prediction is nonetheless that
-stable and predictable (meaning which species will coexist) if species are
-sufficiently dissimilar. Local species richness will first depend on the joint
-effects of local heterogeneity and niche breadth because coexistence requires a
-sufficient dissimilaritya mong species \parencite{Schwilk2005}. Local species richness could be
-increased by a mass effect when dispersal is consistently supplying individuals
-coming from more favorable locations (refuges). The limiting similarity required
-to maintain regional coexistence depends on the amount of dispersal because
-exchanges among communities homogenizes environmental conditions. This is one of
-the main result from the species sorting theory and a clever example of
-local-regional feedbacks: increasing dispersal promotes local coexistence, but
-on the other hand it diminishes regional coexistence. Only the best average
-competitors will remain at very high dispersal.  We therefore expect a
-hump-shaped relationship between dispersal and alpha diversity, with a peak at
+shown that predictions are extremely variables and depend on the frequency
+distributions of environmental conditions, niche optimums and breadth. For
+instance, a well-studied prediction of neutral models is the species abundance
+distribution. It was shown that niche models can predict similar distributions
+given appropriate parameters \parencite{Tilman2004,Gravel2006}. The main
+prediction is nonetheless that stable and predictable (meaning which species
+will coexist) if species are sufficiently dissimilar, which differs from neutral
+models. Local species richness will first depend on the joint effects of local
+heterogeneity and niche breadth because coexistence requires a sufficient
+dissimilaritya mong species \parencite{Schwilk2005}. Local species richness
+could be increased by a mass effect when dispersal is consistently supplying
+individuals coming from more favorable locations (refuges). The limiting
+similarity required to maintain regional coexistence depends on the amount of
+dispersal because exchanges among communities homogenizes environmental
+conditions. This is one of the main result from the species sorting theory and a
+clever example of local-regional feedbacks: increasing dispersal promotes local
+coexistence, but on the other hand it diminishes regional coexistence. Only the
+best average competitors will remain at very high dispersal. We therefore expect
+a hump-shaped relationship between dispersal and alpha ($\alpha$) diversity, with a peak at
 intermediate dispersal. On the other hand, we expect a monotonic decrease of
-beta and gamma diversity with dispersal \parencite{Mouquet2003}. This prediction
+beta ($\beta$) and $\gamma$ diversity with dispersal \parencite{Mouquet2003}. This prediction
 has been validated in some experiments \parencite{Venail2008, Logue2011}.
@@
 properties of the community under the scenarios represented by each
-meta-community model. Our analysis is not exhaustive, it is provided simply to
+metacommunity model. Our analysis is not exhaustive, it is provided simply to
 illustrate the interaction between metacommunity perspectives and landscape
-structures on alpha, beta and gamma diversity.
+structures on $\alpha$, $\beta$ and $\gamma$ diversity.
 
@@
 assumptions of metacommunity dynamics and network structure. We scaled the
-species richness by the maximal local species richness to facilitate comparison
+species richness by the maximal $\alpha$ diversity to facilitate comparison
 between models. The model parameterization is responsible for differences in
@@
 that both in the geographical and tree graph, the path dynamics model has a
-much more considerable variation in local species richness.  However, in all
-cases the alpha diversity increases with the node degree centrality, meaning
+much more considerable variation in $\alpha$ diversity.  However, in all
+cases the $\alpha$ diversity increases with the node degree centrality, meaning
 that nodes with more connections also host a more diverse community.
@@
 diversity as a function of the shape of the network, under the three dynamic
-models. We used Bray-Curtis measure of dissimilarity between patches. In
-Fig. 3, the distance is expressed as the euclidean (geographic) distance between
-two patches. Although this neglects how dispersal connects the different
-patches, there is already a clear signal of geographic distance on $\beta$
-diversity. In both the neutral and patch dynamics model, local communities
-become increasingly dissimilar when the distance between them increases. In
-other words, two communities which are close to each other will share a large
-proportion of their species pool, whereas two communities which are afar will share a
-small proportion.  In the species-sorting model, the relationship between
-distance and dissimilarity is similar, Nonetheless, it forms an enveloppe of
-points (with most points lying in the upper-left part of the graph). While two
-distant communities will be dissimilar, there is no telling how dissimilar two
-close communities will be. Note this relationship for SS varies significantly
-with the spatial distribution of microsites (not shown). At one extreme, if all
-patches hold the same average conditions, then we should expect no relationship
-between dissimilarity and distance. On the other hand, if the average conditions
-are highly variable among localities (as in here), then we should expect two
+models. We used Bray-Curtis measure of dissimilarity between patches. In Fig. 3,
+the distance is expressed as the Euclidean (geographic) distance between two
+patches. Although this neglects how dispersal connects the different patches,
+there is already a clear signal of geographic distance on $\beta$ diversity,
+indicating the importance of dispersal under the three scenarios. In both the
+neutral and patch dynamics model, local communities become increasingly
+dissimilar when the distance between them increases. In other words, two
+communities which are close to each other will share a large proportion of their
+species pool, whereas two communities which are afar will share a small
+proportion. The relationship between distance and dissimilarity is similar for
+species-sorting. Nonetheless, it forms an enveloppe of points (with most points
+lying in the upper-left part of the graph). While two distant communities will
+be dissimilar, there is no telling how dissimilar two close communities will be.
+Note this relationship for species-sorting varies significantly with the spatial
+distribution of microsites (not shown). At one extreme, if all patches hold the
+same average conditions, then we should expect no relationship between
+dissimilarity and distance. On the other hand, if the average conditions are
+highly variable among localities (as in here), then we should expect two
 communities close to be potentially dissimilar (if conditions are different) or
@@
 between spatial contingencies (here connectivity and distribution of
-environmental conditions).
+environmental conditions) and dispersal.
 
 To a vast extent, these relationships are preserved when looking at the
-topological distances (Fiig. 4), i.e. along how many edges should one travel to
+geodesic distances (Fiig. 4), i.e. along how many edges should one travel to
 connect two patchs. Interestingly enough, the distance-dissimilarity
@@
 thus a usefull tool to understand the joint effects of dispersal and community
-interactions. Our comparison of the distribution of alpha diversity was
+interactions. Our comparison of the distribution of $\alpha$ diversity was
 particularly meaningful in that respect. The simulation results show that
@@
 that we must move toward more realistic landscapes \parencite{Gilarranz2012}.
-For field ecologists, and particularly limnologists, our review emphasize that
-we need to go beyond geographic based analysis of beta-diversity (e.g.
-\parencite{Legendre2005}) to topological based analyses \parencite{Peterson2013}.
+For field ecologists, and particularly limnologists, our review emphasizes that
+we need to go beyond geographic based analysis of $\beta$ diversity (e.g.
+\parencite{Legendre2005}) to topological based analyses
+\parencite{Peterson2013}.
 
@@
 dispersal kernels. It makes a significant departure to simple island-mainland or
-global dispersal approaches used previously (e.g. \parencite{Tilman1994,
-Mouquet2002, Hubbell2001}). But dispersal is also spatially explicit in a
+global dispersal approaches used previously (e.g. \parencite{Tilman1994, Hubbell2001,
+Mouquet2002}). But dispersal is also spatially explicit in a
 lattice model and it does not make the landscape more realistic. We believe the
@@
 \parencite{Altermat2012}, the simulations show that the degree centrality has a
-significant impact on local species richness. Central nodes might also
-contribute more to maintain regional diversity, as they are essential for
+significant impact on $\alpha$ diversity. Central nodes might also
+contribute more to maintain $\gamma$ diversity, as they are essential for
 species to spread throughout the landscape. The nodes could be potentially
@@
 dynamics. We could even hypothesize it will vanish in the extreme case of niche
-differentiation (with low overlap for instance) and low mass effect. The neutral
+differentiation (with low overlap for instance) and low mass effect. In this particular case, the neutral
 versus niche comparison therefore illustrates that very strong unequal
@@
 reveal there could be spatial variation in coexistence mechanisms. If we take
-the species-sorting perspective for instance, we find that alpha diversity is
-higher in more central nodes. Since the environment is constant, it implies that
-diversity in these communities is maintained by a stronger mass effect. It
-results in spatial variation in the relative importance of species-sorting, the
-mass effect and to a certain extent the neutral drift. Because the degree
-centrality was the best variable explaining diversity, we should expect the
-degree distribution to strongly impact the relative importance of these
-coexistence mechanisms. For a given set of ecological processes and distribution
-of species traits, we might expect the coexistence mechanisms to differ from one
-landscape to another.
+the species-sorting perspective for instance, we find that $\alpha$ diversity is
+higher in more central nodes. Since the environment is on average the same from
+one patch to another, it implies that diversity in these communities is
+maintained by a stronger mass effect. It results in spatial variation in the
+relative importance of species-sorting, the mass effect and to a certain extent
+the neutral drift. Because the degree centrality was the best variable
+explaining diversity, we should expect the degree distribution to strongly
+impact the relative contribution of these coexistence mechanisms. For a given
+set of ecological processes and distribution of species traits, we might expect
+the coexistence mechanisms to differ from one landscape to another.
 
@@
 distribution. This would make a significant improvement over traditional null
-models in ecology (Gotelli1996) where there are no interactions and no dispersal
+models in ecology \parencite{Gotelli1996} in which there are no interactions and no dispersal
 limitations.
 
-The multivariate variance partitioning framework originally proposed by Borcard
-et al. (1992) and further developed by Borcard and Legendre (2002) has been
+The multivariate variance partitioning framework originally proposed by
+\textcite{Borcard1992} and further developed by \textcite{Borcard2002} has been
 widely used to quantify the relative importance of species sorting and dispersal
@@
 model species distribution as a function of environmental variables, taking into
-account the spatial autocorrelation of species distribution (Leduc1992,
-Borcard1992, Legendre1993). This methodology has been widely used over the last
-decade as a test of the neutral theory, its underlying assumption and a
-quantification of dispersal limitations (e.g. Svenning2004; Hardy2004;
-Gilbert2004; Cottenie2005). One problem of this approach is however that it
-makes a weak test of neutrality (McGill2003), based on the description of
-spatial community structure, rather than hypothesis testing. The different
-models we reviewed in this article could be better employed if used to generate
-expectations based on different hypotheses and then compare them. But most of
-all, parametrized spatially explicit neutral models could be more useful if
-used to generate expectations. For instance, neutral models could be used to
-predict the consequences of habitat destruction, fragmentation or a change in
-the connectivity matrix \parencite{Hubbell2008}. The spatially explicit
-description of the landscape is a major improvement toward that end, providing
-much flexibility in the scenarios that could be explored.
+account the spatial autocorrelation of species distribution
+\parencite{Leduc1992, Borcard1992, Legendre1993}. This methodology has been
+widely used over the last decade as a test of the neutral theory, its underlying
+assumption and a quantification of dispersal limitations (e.g.
+\textcite{Svenning2004; Hardy2004; Gilbert2004; Cottenie2005}. One problem of
+this approach is however that it makes a weak test of neutrality
+\parencite{McGill200}, based on the description of spatial community structure,
+rather than hypothesis testing. The different models we reviewed in this article
+could be better employed if used to generate null expectations of species
+distribution based on different hypotheses and then compare them. But most of
+all, parametrized spatially explicit neutral models could be more useful if used
+to predict biodiversity under different global change scenarios. For instance,
+neutral models could be used to predict the consequences of habitat destruction,
+fragmentation or a change in the connectivity matrix \parencite{Hubbell2008}.
+The spatially explicit description of the landscape is a major improvement
+toward that end, providing much flexibility in the scenarios that could be
+explored.
 
-%PDP
-Working with more complex representations on landscapes has several
-advantages. Real landscapes are not flat geometric objects, they are highly
-structured and diverse. This diverse structure has long been recognized as
-a key component of diversity.
-In the 19th century, Moritz Wagner noted that patterns of rivers \cite{coy04} could explain how beetles diversified.
-Yet, to this day, the relationship between spatial structures and biodiversity is not resolved, but networks provide powerful tools to analyze landscapes and generate
-testable predictions on the relationship between community assembly
-and spatial structures. For example, the neutral theory predict less diversity
-in isolated communities and constant speciation (regardless of isolation) \cite{eco08}. Adding the effect of gene flow changes the predictions, but
-the neutral theory do not predict adaptive radiation and might thus face
-problems for predicting many long-term ecological processes.
+Working with more complex representations on landscapes has several advantages.
+Real landscapes are not flat geometric objects, they are highly structured and
+diverse. This diverse structure has long been recognized as a key component of
+diversity. In the 19th century, Moritz Wagner noted that patterns of rivers
+\cite{coy04} could explain how beetles diversified. Yet, to this day, the
+relationship between spatial structures and biodiversity is not resolved, but
+networks provide powerful tools to analyze landscapes and generate testable
+predictions on the relationship between community assembly and spatial
+structures. For example, the neutral theory predict less diversity in isolated
+communities and constant speciation (regardless of isolation) \cite{eco08}.
+Adding the effect of gene flow changes the predictions, but the neutral theory
+do not predict adaptive radiation and might thus face problems for predicting
+many long-term ecological processes.
 
-%TP
 Finally, our analyses emphasize the need to expand on the canonial neutral
-theory. As pointed out by Wootoon2005Nature433, most of the unexplained
+theory. As pointed out by \textcite{Wootoon2005}, most of the unexplained
 deviation of empirical community from the prediction of accurately calibrated
-neutral models can be attributed to non-competitive interactions. Canard2012
+neutral models can be attributed to non-competitive interactions. \textcite{Canard2012}
 proposed that neutral processes can explain the network structure of trophic

### 2013-05-28 ms.tex
Figure legends and formatting

@@
 % TODO:
-% DG: First complete reading
-% DG: Figures
-% DG: Legends
+% DG: Finish figures
 % DG: Compile references
 % TP & PDP: fix bugs I generated!!!
-% DG: Abstract
 
@@
 
-2: Qu\'ebec Centre for Biodiversity Sciences, Stewart Biological Sciences Building, 1205 Dr.~Penfield Avenue, Montr\'eal (QC), H3A 1B1, Canada\\
+2: Qu\'ebec Centre for Biodiversity Sciences\\
 
-\textbf{Words in the abstract:}      \\
-\textbf{Words in the main text:}    \\
-\textbf{Words in the legends:}    \\
-\textbf{References:}             \\
-\textbf{Table:}                    \\
+\textbf{Keywords:} metacommunity; neutral theory; species sorting; metapopulation; spatial network; centrality \\
+
+\textbf{Words in the abstract:}
+
+\textbf{Words in the main text:}
+
+\textbf{Words in the legends:}
+
+\textbf{Figures:} 4
+
+\textbf{Tables:} 2
+
+\textbf{References:}
 
@@
 %========================================================%
-\section{Introduction}
+\section*{Abstract}
+The metacommunity perspective appears naturally as the appropriate conceptual
+framework to make ecology more predictive, integrating elements of theoretical
+ecology. The recent development of neutral theory appears as a step back in that
+direction because of the assumption of ecological equivalence and the absence of
+any effect of the environment on community organization. A remarkable strength
+of neutral theory is nonetheless to provide a general theory of diversity that
+accounts for a wide range of empirical observations. In this paper, we argue
+that neutral theory can be a useful tool to understand the impact of dispersal
+on community organization in landscapes of various complexities. Our main
+objective is to use neutral theory to stress the importance of complex landscape
+network structure on the distribution of diversity. We refer to the landscape
+organization as a "spatial contingency" that could potentially affect the
+coexistence mechanisms at play. We briefly review the main approaches to
+describe spatial networks and describe three simple toy models of metacommunity
+dynamics. We take this opportunity to review their assumptions and main
+predictions. We then conduct simple simulations of these models to reveal with
+simple examples the impact of spatial network structure on diversity
+distribution. The simulation results show that competitive interactions buffer
+the potential impact of landscape structure. The strongest centrality-species
+richness relationship was observed for the patch dynamics, a model without any
+interactions. On the other hand, strong and unequal competitive interactions
+minimized the effect of centrality. We conclude that the neutral model is thus a
+useful tool to understand the joint effects of dispersal and community
+interactions. Our analysis shows that ecologists must now integrate more
+realistic landscapes when analyzing community assembly from a metacommunity
+perspective.
+\newpage
+
+%========================================================%
+\section*{Introduction}
 
 Ecology needs to move toward a more predictive approach, integrating elements of
-theoretical ecology \parencite{Thuiller2013}. The metacommunity perspective \parencite{Leibold2004a}
-appears naturally as the appropriate conceptual framework to fill this
-challenge. The metacommunity concept builds on feedbacks between local scale
-processes, such as competitive interactions and local adaptation, and regional
-scale processes such as dispersal, gene flow and speciation. It is particularly relevant to limnology,
-where exchanges of organisms and nutrients affect community and ecosystem
-properties from the local (e.g. vertical mixing \parencite{Ryabov2011}) to the
-regional (e.g. connection of lakes \parencite{Leibold2004b, Gravel2010} scales.
-It emphasizes the importance of dispersal relative to pairwise interactions in
-the organization of ecological communities.
+theoretical ecology \parencite{Thuiller2013}. The metacommunity perspective
+\parencite{Leibold2004a} appears naturally as the appropriate conceptual
+framework to fill this challenge. The metacommunity concept builds on feedbacks
+between local scale processes, such as competitive interactions and local
+adaptation, and regional scale processes such as dispersal, gene flow and
+speciation. It is particularly relevant to limnology, where exchanges of
+organisms and nutrients affect community and ecosystem properties from the local
+(e.g. vertical mixing \parencite{Ryabov2011}) to the regional (e.g. connection
+of lakes \parencite{Leibold2004b, Gravel2010} scales. It emphasizes the
+importance of dispersal relative to pairwise interactions in the organization of
+ecological communities.
 
@@
 Neutral theory of biodiversity makes the provocative assumption that species are
-ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral ecological communities are
-driven only by demographic stochasticity and dispersa and thus, variation in the environment has no impact on
-demography. Neutral theory sparked
+ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral ecological
+communities are driven only by demographic stochasticity and dispersa and thus,
+variation in the environment has no impact on demography. Neutral theory sparked
 an historical debate still lasting after more than a decade
-\parencite{Chave2004, Etiennee2011, Rosindell2012,Clark2012}. It was stimulated
+\parencite{Chave2004, Etienne2011, Rosindell2012,Clark2012}. It was stimulated
 by the surprising ability of neutral models to fit some well studied empirical
@@
 landscape network structure on the distribution of diversity. We refer to the
-landscape organization as a "spatial contingency" \parencite{Peres-Neto2013}
+landscape organization as a \emph{spatial contingency} \parencite{Peres-Neto2013}
 that could potentially affect the coexistence mechanisms at play. We will
@@
 %========================================================%
-\section{Network representation of landscapes}
+\section*{Network representation of landscapes}
 
@@
 
-\begin{table*}
-	\centering
-	\begin{tabular}{p{5cm}p{8cm}}
-		\hline
-		\textbf{Concept} & \textbf{Definition} \\
-		\hline
-		Path & A sequence of edges forming a sequence of nodes \\
-		Connection & Two nodes are connected if there is a path between them \\
-		Euclidean distance & Geographical distance between two nodes \\
-		Geodesic distance & Length of the shortest path between two nodes \\
-
-		\textbf{Network-level metrics} & \\
-		Order & Total number of nodes \\
-		Size & Total number of edges \\
-		Connectivity & A measure of robustness: the minimum number of elements to remove to isolate the nodes \\
-		Components &  The number of connected subsets\\
-
-		\textbf{Node-level metrics} & \\
-		Degree & The number of edges of a node \\
-		Closeness centrality & Average geodesic distance between a node and all other vertex \\
-		Eigenvector centrality & A measure of centrality based on the concept that connection to
-		highly connected nodes are more important \\
-		Betweenness centrality & The number of shortest paths from all nodes to all others that
-		pass through that note \\
-		\hline
-	\end{tabular}
-	\caption{Main descriptors of spatial networks used in this study.}
-\end{table*}
-
 %========================================================%
@@
 Material.
-
-\begin{table*}
-	\centering
-	\begin{tabular}{llccc}
-	\hline
-	& Definition & Patch dynamics & Neutral & Species-sorting \\
-	\hline
-	\textbf{Variables} &     &    &    &   \\
-	$p$ & Occupancy  & X &    &   \\
-	$N$ & Local population size &    & X & X \\
-	$Z$ & Local rel. abund. &  & X & X \\
-	$P$ & Rel. abund. in the neighborhood &  & X & X \\
-	$s$ & Local species richness & X & X & X \\
-	$d$ & Node degree & X & X & X \\
-	$C$ & Prob. of a colonization event & X &  &  \\
-	$I$ & Prob. of a colonization event & X &  &  \\
-	$Pr$ & Recruitment prob. &  & X & X \\
-	$\lambda$ & Survival prob. &  &  & X \\
-
-	\textbf{Indices} & & & &\\
-	$x,y$ & Node location & X & X & X \\
-	$i,j$ & Species & X & X & X\\
-	$n$ & Microsite & & & X \\
-
-	\textbf{Parameters} & & & &\\
-	$S$ & Size of regional species pool & X & X & X \\
-	$c$ & Colonization prob. & X & & \\
-	$e$ & Extinction prob. & X & & \\
-	$J$ & Local carrying capacity & & X & X \\
-	$m$ & immigration prob. from neigh. & & X & X \\
-	$M$ & immigration prob. from metaco. & & X & X \\
-	$k$ & Death prob. & & X & X \\
-	$u$ & Niche optimum & & & X \\
-	$b$ & Niche breadth & & & X \\
-	$E$ & Microsite env. conditions & & & X \\
-	$\overline{E}$ & Local env. average & & & X \\
-	$\sigma$ & Local env. variance & & & X \\
-	$\overline{E_R}$ & Regioal env. average & & & X \\
-	$\sigma_R$ & Regional env. variance & & & X \\
-	\hline
-	\end{tabular}
-	\caption{List of variables, indices and parameters from the three models}
-\end{table*}
 
@@
 predator-prey interactions have been added to this framework (e.g.
-\textcite{Tilman1994,Holt1996,Klausmeier1998,Gravel2011b}) but we will keep this model
-minimal for the sake of comparison with the neutral model.
+\textcite{Tilman1994,Holt1996,Klausmeier1998,Gravel2011b}) but we will keep this
+model minimal for the sake of comparison with the neutral model.
 
@@
 Given that all species are the same, then we should expect the regional
-diversity ($\gamma$) to be $S$ if this condition is satisfied and $0$ if not. The situation
-is however more complex in spatially explicit landscapes with complex
-connectivity matrices \parencite{Hanski1998}. Spatially explicit dispersal
-usually reduces the occupancy and thereby the likelihood of persistence. The
-second prediction is that, given spatial variation in connectivity, there will
-be spatial variation in occurrence probability. Given the above formulation of a
-colonization event to occur, the probability that an empty location is colonized
-is $I_ix=1-\prod d_x(1-C_{ixy})$. This equation basically tells us that the
-colonization probability will increase asymptotically with the degree of a patch
-(because of the product). It is easy to show from metapopulation theory that the
-occurrence probability in a patch is then $p_ix=I_x(I_ix+e)^{-1}$. The feedback
-between local and regional dynamics arises because all $p_ix$ from the landscape
-are dependent from each other. Simulations are usually conducted to solve the
-model for a large landscape, but numerical solutions are theoretically possible.
-The aggregation across the $S$ species of the regional species pool is obtained
-by taking the summation of occurrence probabilities over all species, $s_x =
-\sum{p_i}$. Because in this model all species are equal, we expect the local
-species richness to be a linear function of the patch degree (number of edges). Multi-species
-analysis of metapopulation models also reveals interesting predictions on other
-aspects of community organization at various spatial scales such as the species-area
+diversity ($\gamma$) to be $S$ if this condition is satisfied and $0$ if not.
+The situation is however more complex in spatially explicit landscapes with
+complex connectivity matrices \parencite{Hanski1998}. Spatially explicit
+dispersal usually reduces the occupancy and thereby the likelihood of
+persistence. The second prediction is that, given spatial variation in
+connectivity, there will be spatial variation in occurrence probability. Given
+the above formulation of a colonization event to occur, the probability that an
+empty location is colonized is $I_ix=1-\prod d_x(1-C_{ixy})$. This equation
+basically tells us that the colonization probability will increase
+asymptotically with the degree of a patch (because of the product). It is easy
+to show from metapopulation theory that the occurrence probability in a patch is
+then $p_ix=I_x(I_ix+e)^{-1}$. The feedback between local and regional dynamics
+arises because all $p_ix$ from the landscape are dependent from each other.
+Simulations are usually conducted to solve the model for a large landscape, but
+numerical solutions are theoretically possible. The aggregation across the $S$
+species of the regional species pool is obtained by taking the summation of
+occurrence probabilities over all species, $s_x = \sum{p_i}$. Because in this
+model all species are equal, we expect the local species richness to be a linear
+function of the patch degree (number of edges). Multi-species analysis of
+metapopulation models also reveals interesting predictions on other aspects of
+community organization at various spatial scales such as the species-area
 relationship \parencite{Hanski1997}, and proved to be useful in conservation
@@
 disturbances could be simulated as well, as long as they hit all species with
-the same probability, independently of their abundance. The fundamental feature of
-neutral dynamics is therefore the ecological drift, defined as population
+the same probability, independently of their abundance. The fundamental feature
+of neutral dynamics is therefore the ecological drift, defined as population
 changes emerging from neutraly stable population dynamics. It can be measured as
@@
 the species-area relationship, community turnover (beta-diversity) and
-co-occurrence. Recent trophic neutral models were also found to predict realistic ecological
-network structures \parencite{Canard2012}. Other than the ecological equivalence
-assumption, one of the most criticized aspect of neutral models is the realism
-of the speciation process and the required speciation rates to sustain species
-richness \parencite{Ricklefs2003,Etienne2007}. Recent neutral models with more
-credible speciation models \parencite{Rosindell2009,Desjardins2012a} revealed
-the difficulty to maintain diversity in neutral models over macro-evolutionary
-time scales. These models nonetheless proposed interesting predictions
-on endemic species richness and island biogeography \parencite{Rosindell2011,Desjardins2012b}.
+co-occurrence. Recent trophic neutral models were also found to predict
+realistic ecological network structures \parencite{Canard2012}. Other than the
+ecological equivalence assumption, one of the most criticized aspect of neutral
+models is the realism of the speciation process and the required speciation
+rates to sustain species richness \parencite{Ricklefs2003,Etienne2007}. Recent
+neutral models with more credible speciation models
+\parencite{Rosindell2009,Desjardins2012a} revealed the difficulty to maintain
+diversity in neutral models over macro-evolutionary time scales. These models
+nonetheless proposed interesting predictions on endemic species richness and
+island biogeography \parencite{Rosindell2011,Desjardins2012b}.
 
@@
 best average competitors will remain at very high dispersal. We therefore expect
-a hump-shaped relationship between dispersal and alpha ($\alpha$) diversity, with a peak at
-intermediate dispersal. On the other hand, we expect a monotonic decrease of
-beta ($\beta$) and $\gamma$ diversity with dispersal \parencite{Mouquet2003}. This prediction
-has been validated in some experiments \parencite{Venail2008, Logue2011}.
+a hump-shaped relationship between dispersal and alpha ($\alpha$) diversity,
+with a peak at intermediate dispersal. On the other hand, we expect a monotonic
+decrease of beta ($\beta$) and $\gamma$ diversity with dispersal
+\parencite{Mouquet2003}. This prediction has been validated in some experiments
+\parencite{Venail2008, Logue2011}.
 
-\section{Results}
+\section*{Results}
 
@@
 %========================================================%
-\section{Discussion}
+\section*{Discussion}
 
@@
 models is the effect of unequal competitive interactions. The neutral model is
-thus a usefull tool to understand the joint effects of dispersal and community
+thus a useful tool to understand the joint effects of dispersal and community
 interactions. Our comparison of the distribution of $\alpha$ diversity was
@@
 dispersal kernels. It makes a significant departure to simple island-mainland or
-global dispersal approaches used previously (e.g. \parencite{Tilman1994, Hubbell2001,
-Mouquet2002}). But dispersal is also spatially explicit in a
+global dispersal approaches used previously (e.g. \parencite{Tilman1994,
+Hubbell2001, Mouquet2002}). But dispersal is also spatially explicit in a
 lattice model and it does not make the landscape more realistic. We believe the
@@
 \parencite{Altermat2012}, the simulations show that the degree centrality has a
-significant impact on $\alpha$ diversity. Central nodes might also
-contribute more to maintain $\gamma$ diversity, as they are essential for
-species to spread throughout the landscape. The nodes could be potentially
-quantified as keystone for the metacommunity \parencite{Mouquet2013}.
-Interestingly, but not surprisingly, this effect is weaker with species sorting
-dynamics. We could even hypothesize it will vanish in the extreme case of niche
-differentiation (with low overlap for instance) and low mass effect. In this particular case, the neutral
-versus niche comparison therefore illustrates that very strong unequal
+significant impact on $\alpha$ diversity. Central nodes might also contribute
+more to maintain $\gamma$ diversity, as they are essential for species to spread
+throughout the landscape. The nodes could be potentially quantified as keystone
+for the metacommunity \parencite{Mouquet2013}. Interestingly, but not
+surprisingly, this effect is weaker with species sorting dynamics. We could even
+hypothesize it will vanish in the extreme case of niche differentiation (with
+low overlap for instance) and low mass effect. In this particular case, the
+neutral versus niche comparison therefore illustrates that very strong unequal
 competitive interactions could overwhelm the impact of dispersal.
@@
 assumption and a quantification of dispersal limitations (e.g.
-\textcite{Svenning2004; Hardy2004; Gilbert2004; Cottenie2005}. One problem of
+\textcite{Svenning2004, Hardy2004, Gilbert2004, Cottenie2005}. One problem of
 this approach is however that it makes a weak test of neutrality
@@
 %========================================================%
-\section{Acknowledgements}
+\section*{Acknowledgements}
 DG received financial support from NSERC and Canada Research Chair program. TP
-is funded by a MELS-FQRNT post-doctoral fellowship and PDP by a NSERC
+is supported by a MELS-FQRNT post-doctoral fellowship and PDP by a NSERC
 fellowship.
@@
 \newpage
+
 %========================================================%
@@
 \textbf{Illustration of the four simulated landscapes}. The color code
-represents the local species richness simulated with a neutral model, ranked
-from the poorest (red) to the richest (blue). Parameters are: $P = 100$, $N =
-25$, $m = 0.2$, $M = 0.01$, $k = 0.1$, $J_x = 100$.
+represents the $\alpha$ diversity simulated with a neutral model, ranked from
+the poorest (red) to the richest (blue). Parameters: $N = 25$, $r = 0.3 $, $S =
+100$, $m = 0.2$, $M = 0.01$, $k = 0.1$, $J_x = 100$. Simulations run 1000 time
+steps.
 
@@
 \subsection*{Figure 2}
-\textbf{Relationship between local species richness and metrics of node
-centrality}. The upper two panels are simulation results conducted with the
-random geometric graph illustrated at Fig. 1 and the lower two panels are runs
-with the random tree graph.
+\textbf{Relationship between $\alpha$ diversity and node centrality}. The upper
+two panels are simulation results conducted with the random geometric graph
+illustrated at Fig. 1 and the lower two panels are runs with the random tree
+graph. Parameters:$S = 100$, $c = 0.4$, $e = 0.1$, $J_x = 100$, $m = 0.2$, $M =
+0.01$, $k = 0.1$, $ u \epsilon [0,100]$, $b = 15$, $ E_x \epsilon [0,100]$,
+$\overline{E_R} = 50$, $\sigma{E_R} = 5$. Simulations run 1000 time steps
 
@@
 \subsection*{Figure 3}
-\textbf{Bray curtis dissimilarity as a function of geographic distance}.
+\textbf{Bray curtis dissimilarity as a function of Euclidean distance}. Parameters as in Fig. 2.
 
@@
 \subsection*{Figure 4}
-\textbf{Bray curtis dissimilarity as a function of topological distance}.
+\textbf{Bray curtis dissimilarity as a function of geodesic distance}. Parameters as in Fig. 2.
 
@@
 %========================================================%
-
-\section*{Figures}
-
 %------------------------
@@
 
+%========================================================%
+
+\begin{table*}[c]
+	\centering
+	\begin{tabular}{p{5cm}p{8cm}}
+		\hline
+		\textbf{Concept} & \textbf{Definition} \\
+		\hline
+		Path & A sequence of edges forming a sequence of nodes \\
+		Connection & Two nodes are connected if there is a path between them \\
+		Euclidean distance & Geographical distance between two nodes \\
+		Geodesic distance & Length of the shortest path between two nodes \\
+
+		\textbf{Network-level metrics} & \\
+		Order & Total number of nodes \\
+		Size & Total number of edges \\
+		Connectivity & A measure of robustness: the minimum number of elements to remove to isolate the nodes \\
+		Components &  The number of connected subsets\\
+
+		\textbf{Node-level metrics} & \\
+		Degree & The number of edges of a node \\
+		Closeness centrality & Average geodesic distance between a node and all other vertex \\
+		Eigenvector centrality & A measure of centrality based on the concept that connection to
+		highly connected nodes are more important \\
+		Betweenness centrality & The number of shortest paths from all nodes to all others that
+		pass through that note \\
+		\hline
+	\end{tabular}
+	\caption{Main descriptors of spatial networks used in this study.}
+\end{table*}
+\newpage
+
+%------------------------
+
+\begin{table*}[c]
+	\centering
+	\begin{tabular}{llccc}
+	\hline
+	& Definition & Patch dynamics & Neutral & Species-sorting \\
+	\hline
+	\textbf{Variables} &     &    &    &   \\
+	$p$ & Occupancy  & X &    &   \\
+	$N$ & Local population size &    & X & X \\
+	$Z$ & Local rel. abund. &  & X & X \\
+	$P$ & Rel. abund. in the neighborhood &  & X & X \\
+	$s$ & Local species richness & X & X & X \\
+	$d$ & Node degree & X & X & X \\
+	$C$ & Prob. of a colonization event & X &  &  \\
+	$I$ & Prob. of a colonization event & X &  &  \\
+	$Pr$ & Recruitment prob. &  & X & X \\
+	$\lambda$ & Survival prob. &  &  & X \\
+
+	\textbf{Indices} & & & &\\
+	$x,y$ & Node location & X & X & X \\
+	$i,j$ & Species & X & X & X\\
+	$n$ & Microsite & & & X \\
+
+	\textbf{Parameters} & & & &\\
+	$S$ & Size of regional species pool & X & X & X \\
+	$c$ & Colonization prob. & X & & \\
+	$e$ & Extinction prob. & X & & \\
+	$J$ & Local carrying capacity & & X & X \\
+	$m$ & immigration prob. from neigh. & & X & X \\
+	$M$ & immigration prob. from metaco. & & X & X \\
+	$k$ & Death prob. & & X & X \\
+	$u$ & Niche optimum & & & X \\
+	$b$ & Niche breadth & & & X \\
+	$E$ & Microsite env. conditions & & & X \\
+	$\overline{E}$ & Local env. average & & & X \\
+	$\sigma$ & Local env. variance & & & X \\
+	$\overline{E_R}$ & Regioal env. average & & & X \\
+	$\sigma_R$ & Regional env. variance & & & X \\
+	\hline
+	\end{tabular}
+	\caption{List of variables, indices and parameters from the three models}
+\end{table*}
+\newpage
+
 \end{document}

### 2013-05-28 ms.tex
References

@@
 \usepackage[style=bes]{biblatex}
-\bibliography{./library}
+\bibliography{library}
 
@@
 (e.g. vertical mixing \parencite{Ryabov2011}) to the regional (e.g. connection
-of lakes \parencite{Leibold2004b, Gravel2010} scales. It emphasizes the
+of lakes \parencite{Leibold2004b, Gravel2010a} scales. It emphasizes the
 importance of dispersal relative to pairwise interactions in the organization of
@@
 some aspects of community dynamics. It links to an old philosophical debate
-between realism and instrumentalism \parencite{Wennekes2011}. Because every
+between realism and instrumentalism \parencite{Wennekes2012}. Because every
 ecological model is a simplification of reality, scientists have to subjectively
@@
 understanding or for prediction (another old philosophical debate, see
-\textcite{Schmueli2010}). Obviously neutral theory could only be instrumental.
+\textcite{Shmueli2010}). Obviously neutral theory could only be instrumental.
 The question then is if such a 'general, large-scale, but vague' theory
-\parencite{Wennekes2011} is a satisfying approximation.
+\parencite{Wennekes2012} is a satisfying approximation.
 
@@
 landscape network structure on the distribution of diversity. We refer to the
-landscape organization as a \emph{spatial contingency} \parencite{Peres-Neto2013}
+landscape organization as a \emph{spatial contingency} \parencite{Peres-Neto2012}
 that could potentially affect the coexistence mechanisms at play. We will
@@
 on the variance of dispersal. A second generation of neutral models (e.g.
-\textcite{Economo2008, Economo2011,Desjardins2012a,Desjardins2012b}, and even
-experiments \parencite{Altermat2012}, recently introduced more realistic
+\textcite{Economo2008, Economo2011,Desjardins2012a,Desjardins2012b}), and even
+experiments \parencite{Carrara2012}, recently introduced more realistic
 landscapes and found surprising contributions of spatial contingencies. We start
@@
 A network is a discrete mathematical object made of two sets: a set of nodes
-(or vertices) and a set of edges connecting the nodes \cite{new10}. The
+(or vertices) and a set of edges connecting the nodes \parencite{new10}. The
 term ``graph'' is often preferred in computer science and mathematics
-\cite{gro06}, with graph algorithms being an important and active area of
-research \cite{sed01}. A network is a combinatorial object: it is used to study
+\parencite{gro06}, with graph algorithms being an important and active area of
+research \parencite{sed01}. A network is a combinatorial object: it is used to study
 how discrete entities are connected and how they combine together to create
@@
 networks, or even the relationship between variables in statistics
-\cite{wri21,new10}. We are especially interested in spatial networks, a special
+\parencite{wri21,new10}. We are especially interested in spatial networks, a special
 kind of network mixing the combinatorial properties of networks with a
-topological space \cite{kob94}. Thus, the vertices in a spatial graph are
+topological space \parencite{kob94}. Thus, the vertices in a spatial graph are
 embedded in some other space, most often the two or three- dimensional Euclidean
@@
 distance is the distance in the graph space, i.e.: the length of the shortest
-path \cite{dij59}. For example, two lakes could be very close on a map (short
+path \parencite{dij59}. For example, two lakes could be very close on a map (short
 Euclidean distance) but the geodesic distance could be great if they are not
@@
 For simulations, spatial networks can easily be generated with the random
-geometric graph algorithm \cite{sed01}. In this algorithm, all nodes are
+geometric graph algorithm \parencite{sed01}. In this algorithm, all nodes are
 assigned to a position in some two-dimensional space, most often the unit
@@
 and $C$ are more likely to be connected than two random vertices. Random
-geometric networks have been extensively studied \cite{app97a,app97
-b,app02a,app02b,pen03} and we provide a R function to generate them. The
+geometric networks have been extensively studied \parencite{app97a,app97b,app02a,app02b,pen03} 
+and we provide a R function to generate them. The
 position of nodes is typically random, but we could also imagine alterations
@@
 from which to start the tree. It then calculaes the the shortest path tree
-\cite{dij59} from this node to all other ones. Edges that are not located along
+\parencite{dij59} from this node to all other ones. Edges that are not located along
 this tree are pruned. This random geometric tree does not exactly represent
@@
 biology, where patterns of connections can be used to study and influence the
-flow of organisms \cite{min07,fal07,min08,gar08,urb09,dal10}. In the neutral
-theory, networks were pioneered by Economo and Keitt \cite{eco08, eco10}. They
+flow of organisms \parencite{min07,fal07,min08,gar08,urb09,dal10}. In the neutral
+theory, networks were pioneered by Economo and Keitt \parencite{Economo2008, eco10}. They
 used networks to study how different spatial structures influenced diversity.
 They were also used to study how the spatial structure influenced nonsympatric
-speciation \cite{des12,des12b}.
+speciation \parencite{Desjardins2012a,Desjardins2012b}.
 
@@
 predator-prey interactions have been added to this framework (e.g.
-\textcite{Tilman1994,Holt1996,Klausmeier1998,Gravel2011b}) but we will keep this
+\textcite{Tilman1994a,Holt1996,Klausmeier1998,Gravel2011b}) but we will keep this
 model minimal for the sake of comparison with the neutral model.
@@
 ecology with predictions of extinctions following habitat destruction
-\parencite{Nee1994,Rybicki2013}.
+\parencite{Tilman1994b,Rybicki2013}.
 
@@
 neutral models with more credible speciation models
-\parencite{Rosindell2009,Desjardins2012a} revealed the difficulty to maintain
+\parencite{Rosindell2010,Desjardins2012a} revealed the difficulty to maintain
 diversity in neutral models over macro-evolutionary time scales. These models
@@
 dispersal kernels. It makes a significant departure to simple island-mainland or
-global dispersal approaches used previously (e.g. \parencite{Tilman1994,
+global dispersal approaches used previously (e.g. \parencite{Tilman1994a,
 Hubbell2001, Mouquet2002}). But dispersal is also spatially explicit in a
@@
 heterogeneity of dispersal. In agreement with previous theoretical
-\parencite{Economo2011, Desjardins2012} and experimemental studies
-\parencite{Altermat2012}, the simulations show that the degree centrality has a
+\parencite{Economo2011, Desjardins2012b} and experimemental studies
+\parencite{Carrara2012}, the simulations show that the degree centrality has a
 significant impact on $\alpha$ diversity. Central nodes might also contribute
@@
 this approach is however that it makes a weak test of neutrality
-\parencite{McGill200}, based on the description of spatial community structure,
+\parencite{McGill2003}, based on the description of spatial community structure,
 rather than hypothesis testing. The different models we reviewed in this article
@@
 structures. For example, the neutral theory predict less diversity in isolated
-communities and constant speciation (regardless of isolation) \cite{eco08}.
+communities and constant speciation (regardless of isolation) \parencite{Economo2008}.
 Adding the effect of gene flow changes the predictions, but the neutral theory
@@
 Finally, our analyses emphasize the need to expand on the canonial neutral
-theory. As pointed out by \textcite{Wootoon2005}, most of the unexplained
+theory. As pointed out by \textcite{Wootton2005}, most of the unexplained
 deviation of empirical community from the prediction of accurately calibrated
@@
 \newpage
-
 %========================================================%

### 2013-05-30 ms.tex
Updates on figures 4-5
Small edit on the ms

@@
 \bibliography{library}
-
-% TODO:
-% DG: Finish figures
-% DG: Compile references
-% TP & PDP: fix bugs I generated!!!
 
@@
 
-The popularity of network theory stems for its ability to model complex
+The popularity of network theory stems from its ability to model complex
 structures while allowing us to extract useful metrics (Table 1). At a very high
@@
 from which to start the tree. It then calculaes the the shortest path tree
-\parencite{dij59} from this node to all other ones. Edges that are not located along
-this tree are pruned. This random geometric tree does not exactly represent
-dendritic landscapes but is a convenient model to simulate lake connected by
-rivers to a series of smaller lakes.
+\parencite{dij59} from this node to all other ones. This random geometric tree
+does not exactly represent dendritic landscapes but is a convenient model to
+simulate a lake connected by rivers to a series of smaller lakes.
 
@@
 \begin{figure}[ht!]
-	\centering\includegraphics[width=0.65\textwidth]{BetaGeoDist.pdf}
+	\centering\includegraphics[width=0.65\textwidth]{BetaGeoDist.png}
 \end{figure}
@@
 \begin{figure}[ht!]
-	\centering\includegraphics[width=0.65\textwidth]{BetaTopoDist.pdf}
+	\centering\includegraphics[width=0.65\textwidth]{BetaTopoDist.png}
 \end{figure}

### 2013-05-30 ms.tex

@@
 
-\textbf{Title:}  Using neutral theory to reveal the contribution of dispersal to community assembly in complex landscapes\\
+\textbf{Title:} Using neutral theory to reveal the contribution of meta-community processes to assembly in complex landscapes\\
 
@@
 \section*{Abstract}
-The metacommunity perspective appears naturally as the appropriate conceptual
-framework to make ecology more predictive, integrating elements of theoretical
-ecology. The recent development of neutral theory appears as a step back in that
+The metacommunity perspective appears as an appropriate conceptual
+framework to make ecology more predictive. The recent development of neutral theory appears as a step back in that
 direction because of the assumption of ecological equivalence and the absence of
@@
 accounts for a wide range of empirical observations. In this paper, we argue
-that neutral theory can be a useful tool to understand the impact of dispersal
-on community organization in landscapes of various complexities. Our main
+that neutral theory can be useful to understand the impact of dispersal
+on community assembly in landscapes of various complexities. Our main
 objective is to use neutral theory to stress the importance of complex landscape
@@
 decide the level of details they put in, leaving out some elements they consider
-unimportant. The realism perspective requires that all assumptions of the theory
-must be true, while the utility of the theory is more important to
+unimportant. The realist perspective requires all assumptions of the theory
+to be true, while the utility of the theory is more important to
 instrumentalism. The utilitarian value of a theory could either be for
@@
 To a vast extent, these relationships are preserved when looking at the
-geodesic distances (Fiig. 4), i.e. along how many edges should one travel to
+geodesic distances (Fig. 4), i.e. along how many edges should one travel to
 connect two patchs. Interestingly enough, the distance-dissimilarity
@@
 more to maintain $\gamma$ diversity, as they are essential for species to spread
-throughout the landscape. The nodes could be potentially quantified as keystone
+throughout the landscape. Such nodes could be potentially quantified as keystone
 for the metacommunity \parencite{Mouquet2013}. Interestingly, but not

### 2013-05-31 codeR/examples.R
final revisions before submission

@@ -41,12 +41,12 @@ source("lottery_model.R")
 # Default parameters
 m = 0.2
 M = 0.01
-d = 0.1
+k = 0.1
 J = 100
 sdN = Inf
 sdE = 5
 spatial_graph = geograph_fn(n,r)
-results = lottery_model(m,M,d,S,J,sdE,sdN,spatial_graph,nsteps)
+results = lottery_model(m,M,k,S,J,sdE,sdN,spatial_graph,nsteps)
 
 # Figure 1: time series
 x11(height = 5.5, width = 6)
@@ -68,12 +68,12 @@ source("lottery_model.R")
 # Default parameters
 m = 0.2
 M = 0.01
-d = 0.1
+k = 0.1
 J = 100
 sdN = 15
 sdE = 5
 spatial_graph = geograph_fn(n,r)
-results = lottery_model(m,M,d,S,J,sdE,sdN,spatial_graph,nsteps)
+results = lottery_model(m,M,k,S,J,sdE,sdN,spatial_graph,nsteps)
 
 # Figure 1: time series
 x11(height = 5.5, width = 6)

### 2013-05-31 codeR/lottery_model.R

@@ -16,7 +16,7 @@ lottery_model = function(m,M,d,S,J,sdE,sdN,spatial_graph,nsteps) {
 	# Args: 
 	# 	m: immigration probability from the neighbourhood
 	#	M: immigration probability from outside the metacommunity 
-	#	d: local death rate 	
+	#	k: local death rate 	
 	# 	S: number of species
 	#	J: local community size
 	# 	sdE: within patch standard deviation of the environment
@@ -43,7 +43,6 @@ lottery_model = function(m,M,d,S,J,sdE,sdN,spatial_graph,nsteps) {
 	Env = matrix(nr=n, nc = J)
 	for(i in 1:n) Env[i,] = rnorm(J,mean = runif(1,0,100), sd = sdE) 	# For a random distribution of environments
 #	for(i in 1:n) Env[i,] = rnorm(J,mean = i/n*100, sd = 0)				# For a uniform distribution of environments		
-		
 	####################
 	# Niche optimums
 	u = runif(S, 0, 100) 	# For a random distribution of optimums
@@ -75,7 +74,7 @@ lottery_model = function(m,M,d,S,J,sdE,sdN,spatial_graph,nsteps) {
 		for(i in 1:n) {	
 			# Kill individuals at random
 			rand = runif(J,0,1)
-			localC[[i]][rand < d,] = 0
+			localC[[i]][rand < k,] = 0
 		
 			####################
 			# Calculate recruitment probability		

### 2013-05-31 codeR/simulations.R

@@ -1,5 +1,5 @@
 ########################################################################
-# Source for the figures 1-4 in the paper
+# Source code for the figures 1-4 in the paper
 # Gravel, Poisot and Desjardins. 2013. Using neutral theory to reveal the contribution of dispersal to community assembly in complex landscapes. J. Limnology. 
 #
 # By: Dominique Gravel (dominique_gravel@uqar.ca) 
@@ -110,17 +110,17 @@ betaPatchGeotree = as.matrix(vegdist(patchGeotree[[1]],method = "bray",diag=T,up
 # Specific parameters
 m = 0.2
 M = 0.01
-d = 0.1
+k = 0.1
 J = 100
 sdN = Inf
 sdE = 5
 nsteps = 1000
 
 # Simulations
-neutralConnected = lottery_model(m,M,d,S,J,sdE,sdN,connected,nsteps)
-neutralLattice = lottery_model(m,M,d,S,J,sdE,sdN,lattice,nsteps)
-neutralGeograph = lottery_model(m,M,d,S,J,sdE,sdN,geograph,nsteps)
-neutralGeotree = lottery_model(m,M,d,S,J,sdE,sdN,geotree,nsteps)
+neutralConnected = lottery_model(m,M,k,S,J,sdE,sdN,connected,nsteps)
+neutralLattice = lottery_model(m,M,k,S,J,sdE,sdN,lattice,nsteps)
+neutralGeograph = lottery_model(m,M,k,S,J,sdE,sdN,geograph,nsteps)
+neutralGeotree = lottery_model(m,M,k,S,J,sdE,sdN,geotree,nsteps)
 
 # Alpha diversity
 alphaNeutralConnected = alpha_fn(neutralConnected[[1]])
@@ -139,17 +139,17 @@ betaNeutralGeotree = as.matrix(vegdist(neutralGeotree[[1]],method = "bray",diag=
 # Specific parameters
 m = 0.2
 M = 0.01
-d = 0.1
+k = 0.1
 J = 100
 sdN = 15
 sdE = 5
 nsteps = 1000
 
 # Simulations
-ssConnected = lottery_model(m,M,d,S,J,sdE,sdN,connected,nsteps)
-ssLattice = lottery_model(m,M,d,S,J,sdE,sdN,lattice,nsteps)
-ssGeograph = lottery_model(m,M,d,S,J,sdE,sdN,geograph,nsteps)
[…]

### 2013-09-24 revisions/ms.tex
Minor_revisions

--- 2013-05-30 ms.tex
@@
 
-\textbf{Words in the abstract:}
+\textbf{Words in the abstract:}   276
 
-\textbf{Words in the main text:}
+\textbf{Words in the main text:}  5992
 
-\textbf{Words in the legends:}
+\textbf{Words in the legends:}   103
 
@@
 
-\textbf{References:}
+\textbf{References:} 82
 
@@
 \section*{Abstract} 
-The metacommunity perspective appears as an appropriate conceptual
-framework to make ecology more predictive. The recent development of neutral theory appears as a step back in that
-direction because of the assumption of ecological equivalence and the absence of
-any effect of the environment on community organization. A remarkable strength
-of neutral theory is nonetheless to provide a general theory of diversity that
-accounts for a wide range of empirical observations. In this paper, we argue
-that neutral theory can be useful to understand the impact of dispersal
-on community assembly in landscapes of various complexities. Our main
-objective is to use neutral theory to stress the importance of complex landscape
-network structure on the distribution of diversity. We refer to the landscape
-organization as a "spatial contingency" that could potentially affect the
-coexistence mechanisms at play. We briefly review the main approaches to
-describe spatial networks and describe three simple toy models of metacommunity
-dynamics. We take this opportunity to review their assumptions and main
-predictions. We then conduct simple simulations of these models to reveal with
-simple examples the impact of spatial network structure on diversity
-distribution. The simulation results show that competitive interactions buffer
-the potential impact of landscape structure. The strongest centrality-species
-richness relationship was observed for the patch dynamics, a model without any
+The metacommunity perspective appears as an appropriate conceptual framework to
+make ecology more predictive. It is particularly relevant to limnology, where
+exchanges of organisms and nutrients affect community and ecosystem properties
+from the local to the regional scales. The recent development of neutral theory
+appears as a step back in that direction because of the assumption of ecological
+equivalence and the absence of any effect of the environment on community
+organization. A remarkable strength of neutral theory is nonetheless to provide
+a general theory of diversity that accounts for a wide range of empirical
+observations. In this paper, we argue that neutral theory can be useful to
+understand the impact of dispersal on community assembly in landscapes of
+various complexities. Our analysis focus on spatially explicit landscapes
+conceptualized as networks of local communities (e.g. lakes) connected to each
+other by dispersal channels (e.g. rivers). The main objective of the paper is to
+use neutral theory to stress the importance of landscape structure on the
+distribution of diversity. We refer to the landscape organization as a "spatial
+contingency" that could potentially affect the coexistence mechanisms at play.
+We briefly review the main approaches to describe spatial networks and describe
+three simple toy models of metacommunity dynamics. We take this opportunity to
+review their assumptions and main predictions. We then conduct simulations of
+these models to reveal with simple examples the impact of spatial network
+structure on diversity distribution. The simulation results show that
+competitive interactions buffer the potential impact of landscape structure. The
+strongest relationship between node position in the landscape and species
+richness was observed for the patch dynamics model without any
 interactions. On the other hand, strong and unequal competitive interactions
-minimized the effect of centrality. We conclude that the neutral model is thus a
-useful tool to understand the joint effects of dispersal and community
-interactions. Our analysis shows that ecologists must now integrate more
+minimized the effect of node position. We conclude that the neutral model is a
+useful tool to understand the joint effects of dispersal and ecological
+interactions. Our analysis shows that limnologists must now integrate more
 realistic landscapes when analyzing community assembly from a metacommunity
@@
 organisms and nutrients affect community and ecosystem properties from the local
-(e.g. vertical mixing \parencite{Ryabov2011}) to the regional (e.g. connection
-of lakes \parencite{Leibold2004b, Gravel2010a} scales. It emphasizes the
-importance of dispersal relative to pairwise interactions in the organization of
-ecological communities.
+(e.g. vertical mixing \parencite{Ryabov2011}) to the regional (e.g. connections
+of lakes \parencite{Leibold2004b, Muneepeerakul2008, Gravel2010a} scales. The
+metacommunity perspective emphasizes the importance of dispersal relative to
+pairwise interactions in the organization of ecological communities.
 
 At first sight, the development of neutral theory appears as a step back.
-Neutral theory of biodiversity makes the provocative assumption that species are
-ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral ecological
-communities are driven only by demographic stochasticity and dispersa and thus,
+The neutral theory of biodiversity makes the provocative assumption that species are
+ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral community
+dynamics are driven only by demographic stochasticity and dispersal and thus,
 variation in the environment has no impact on demography. Neutral theory sparked
-an historical debate still lasting after more than a decade
+a historical debate still lasting after more than a decade
 \parencite{Chave2004, Etienne2011, Rosindell2012,Clark2012}. It was stimulated
@@
 \parencite{Bell2001}. Even if new studies rejecting neutral theory are
-consistently published (e.g. \textcite{Ricklefs2012}), a consensus is forming
-that neutral theory is a well-developed null hypothesis for niche theory and
-could even be used as an adequate approximation of ecological dynamics in some
-situations. \textcite{Bell2001} nicely envisioned two perspectives to neutral
-theory that are still standing today. Under the weak perspective, neutral theory
-provides a set of realistic predictions of community organization despite false
-assumptions. Even if being fundamentally wrong, neutral theory would still
-useful when used as a null hypothesis \parencite{Gotelli2006}. It is considered
-as an improvement over traditional null hypotheses based on randomization
-\parencite{Gotelli2000} because it readily integrates dispersal. On the other
-had, the strong version posits that neutral theory is a satisfying approximation
-to community dynamics and an appropriate theory to explain the distribution of
-biodiversity. It implies that the right mechanisms have been identified and that
-the consistently observed differences among species do not impact community
-organization.
+consistently published (e.g. \textcite{Ricklefs2012}, but see the comments
+associated with this study, e.g. \textcite{Etienne2012b}), a consensus is
+forming that neutral theory is a well-developed null hypothesis for niche theory
+and could even be used as an adequate approximation of ecological dynamics in
+some situations. \textcite{Bell2001} nicely envisioned two perspectives to
+neutral theory that are still standing today. Under the weak perspective,
+neutral theory provides a set of realistic predictions of community organization
+despite false assumptions. Even if being fundamentally wrong, neutral theory
+would still useful when used as a null hypothesis \parencite{Gotelli2006}. It is
+considered as an improvement over traditional null hypotheses based on
+randomization \parencite{Gotelli2000} because it readily integrates dispersal.
+On the other had, the strong version posits that neutral theory is a satisfying
+approximation to community dynamics and an appropriate theory to explain the
+distribution of biodiversity. It implies that the right mechanisms have been
+identified and that the consistently observed differences among species do not
+strongly impact community organization.
 
@@
 ecological model is a simplification of reality, scientists have to subjectively
-decide the level of details they put in, leaving out some elements they consider
+decide the level of detail they put in, leaving out some elements they consider
 unimportant. The realist perspective requires all assumptions of the theory
@@
 species and promote ecological drift \parencite{Gravel2011a}. Much has been said
-the existence of demographic stochasticity \parencite{Clark2012}, and we
-therefore will let this discussion for other papers. A second answer is that
+about the existence of demographic stochasticity \parencite{Clark2012}, and we
+will therefore leave this discussion for other papers. A second explanation is that
 dispersal and historical contingencies might have a much more profound impact on
@@
 dynamics. The debate over the equivalence assumption and demographic
-stochasticity has perhaps overlook the recognition of how much dispersal
+stochasticity might have overlooked the recognition of how much dispersal
 influence community assembly.
@@
 landscape network structure on the distribution of diversity. We refer to the
-landscape organization as a \emph{spatial contingency} \parencite{Peres-Neto2012}
-that could potentially affect the coexistence mechanisms at play. We will
-therefore move from a perspective where dispersal is either global or spatially
-explicit (e.g. over a lattice), and spatial constant, to a perspective focusing
+landscape organization as a \emph{spatial contingency}
+\parencite{Peres-Neto2012} that could potentially affect the coexistence
+mechanisms at play. We will therefore move from a perspective where dispersal is
+either global or constant over space (e.g. a lattice), to a perspective focusing
 on the variance of dispersal. A second generation of neutral models (e.g.
-\textcite{Economo2008, Economo2011,Desjardins2012a,Desjardins2012b}), and even
-experiments \parencite{Carrara2012}, recently introduced more realistic
-landscapes and found surprising contributions of spatial contingencies. We start
-with a short review of the main approaches to describe spatial networks. Then we
-describe three simple toy models of metacommunity dynamics, using this
-opportunity to review their assumptions and main predictions. We provide as
-Supplementary Material the R scripts for the toy models and all simulations
-conducted for this paper. We then conduct simple simulations of these models to
-reveal with simple examples the impact of spatial network structure on diversity
-distribution. We conclude with a discussion on the operationally of the
-framework.
+\textcite{Economo2008, Economo2011,Desjardins2012a,Desjardins2012b}), field
+surveys \parencite{Muneepeerakul2008} and even experiments
+\parencite{Carrara2012}, recently introduced more realistic landscapes and found
+surprising contributions of spatial contingencies. We start with a short review
+of the main approaches to describe spatial networks. Then we describe three
+simple toy models of metacommunity dynamics, using this opportunity to review
+their assumptions and main predictions. We provide as Supplementary Material the
+R scripts for the toy models and all simulations conducted for this paper. We
+then conduct simple simulations of these models to reveal with simple examples
+the impact of spatial network structure on diversity distribution. We conclude
+with a discussion on the operationality of the metacommunity framework.
 
@@
 kind of network mixing the combinatorial properties of networks with a
-topological space \parencite{kob94}. Thus, the vertices in a spatial graph are
+topological space \parencite{kob94}. Thus, the nodes in a spatial graph are
 embedded in some other space, most often the two or three- dimensional Euclidean
@@
 particularly suited for systems of lakes and rivers, which can easily be
-represented by vertices and edges. There are two notions of distance in spatial
+represented by nodes and edges. There are two notions of distance in spatial
 networks. Euclidean distance represents the geographical distance between the
@@
 can have the same degree, with one being connected to a small isolated cluster,
-while the other one is part of one of the biggest network of lake. In this case,
+while the other one is part of one of the biggest network of lakes. In this case,
 measures of centrality like eigen-centrality will weight the importance of the
@@
 
-For simulations, spatial networks can easily be generated with the random
-geometric graph algorithm \parencite{sed01}. In this algorithm, all nodes are
-assigned to a position in some two-dimensional space, most often the unit
-square. Then, all pairs of nodes within some threshold Euclidean distance $r$
-are connected with an edge. The resulting networks have the desirable property
-of locality: if a node $A$ is connected to two vertices $B$ and $C$. then $B$
-and $C$ are more likely to be connected than two random vertices. Random
-geometric networks have been extensively studied \parencite{app97a,app97b,app02a,app02b,pen03} 
-and we provide a R function to generate them. The
+In this paper we study four types of networks (Fig. 1). First, for the sake of
+comparisons, we study a complete network where dispersal is global. Then we
+consider a lattice, which is the most regular type of spatially explicit
+network. Then after we simulated two types of random spatial networks. We
+considered random geometric graphs \parencite{sed01}. In this algorithm, all
+nodes are assigned to a position in some two-dimensional space, most often the
+unit square. Then, all pairs of nodes within some threshold Euclidean distance
+$r$ are connected with an edge. The resulting networks have the desirable
+property of locality: if a node $A$ is connected to two vertices $B$ and $C$.
+then $B$ and $C$ are more likely to be connected than two random vertices.
+Random geometric networks have been extensively studied \parencite{app97a,
+app97b, app02a, app02b, pen03} and we provide a R function to generate them. The
 position of nodes is typically random, but we could also imagine alterations
 where they are either more aggregated or segregated than expected by chance
-alone.
-
-We also provide the code for a second structure that we call a random geometric
-tree. The algorithm first builds a random geometric graph, then select a node
-from which to start the tree. It then calculaes the the shortest path tree
-\parencite{dij59} from this node to all other ones. This random geometric tree
-does not exactly represent dendritic landscapes but is a convenient model to
-simulate a lake connected by rivers to a series of smaller lakes.
+alone. Finally, we also provided the code for a second structure that we call a
+random geometric tree. The algorithm first builds a random geometric graph, then
+selects a node from which to start the tree. It then calculates the the shortest
+path tree \parencite{dij59} from this node to all other ones and remove edges
+not located along this tree. This random geometric tree does not exactly
+represent dendritic landscapes but is a convenient model to simulate a lake
+connected by rivers to a series of smaller lakes.
 
 Spatial graphs are increasingly popular in spatial ecology and conservation
-biology, where patterns of connections can be used to study and influence the
-flow of organisms \parencite{min07,fal07,min08,gar08,urb09,dal10}. In the neutral
-theory, networks were pioneered by Economo and Keitt \parencite{Economo2008, eco10}. They
-used networks to study how different spatial structures influenced diversity.
-They were also used to study how the spatial structure influenced nonsympatric
-speciation \parencite{Desjardins2012a,Desjardins2012b}.
+biology, where the structure of connections can be used to study and influence
+the flow of organisms \parencite{min07,fal07,min08,gar08,urb09,dal10}. In the
+neutral theory, networks were pioneered by Economo and Keitt
+\parencite{Economo2008, eco10}. They used networks to study how different
+spatial structures influenced diversity. They were also used to study how the
+spatial structure influenced nonsympatric speciation
+\parencite{Desjardins2012a,Desjardins2012b}.
 
@@
 differentiation (species sorting) that we will be able to fully understand the
-interaction between these processes and landscape structure. Despite neutral,
+interaction between these processes and landscape structure. Despite being neutral,
 competitive interactions in neutral models are very strong because of the
@@
 The simplest metacommunity model is a $S$ species extension of traditional
-metapopulation models \parencite(Hanski1999). The standard Levins metapopulation
-model \parencite(Levins1969) describes the stochastic colonizations and
+metapopulation models \parencite{Hanski1999}. The standard Levins metapopulation
+model \parencite{Levins1969} describes the stochastic colonizations and
 extinctions of a single species over a homogenous landscape. The basic unit is
@@
 basic metacommunity patch dynamics model could be run by aggregating $S$
-independent metapopulation models \parencite(Hanski1997). There are no
+independent metapopulation models \parencite{Hanski1997}. There are no
 interactions in this simple model, which means there is no limit to local
@@
 the above formulation of a colonization event to occur, the probability that an
-empty location is colonized is $I_ix=1-\prod d_x(1-C_{ixy})$. This equation
+empty location is colonized by an offspring coming from a least one of the $r$ neighbouring patches is $I_{ix}=1-\prod_{y=1}^r d_x(1-C_{ixy})$. This equation
 basically tells us that the colonization probability will increase
@@
 to show from metapopulation theory that the occurrence probability in a patch is
-then $p_ix=I_x(I_ix+e)^{-1}$. The feedback between local and regional dynamics
-arises because all $p_ix$ from the landscape are dependent from each other.
+then $p_{ix}=I_{ix}(I_{ix}+e)^{-1}$. The feedback between local and regional dynamics
+arises because all $p_{ix}$ from the landscape are dependent from each other.
 Simulations are usually conducted to solve the model for a large landscape, but
@@
 based. We therefore considered in our toy model of neutral dynamics that each
-local patch holds $J_x$ individuals. The model tracts the local abundance of all
+local patch holds $J_x$ individuals. The model tracks the local abundance of all
 species $N_{ix}$ in each local patch. At each time step an individidual dies
@@
 The formulation of the recruitment probability is the central piece of all
-neutral models, making possible the coupling with the metacommunity and
-neighbouring patches. We adopt a simple formulation based on
+neutral models, making the coupling with the metacommunity and
+neighbouring patches possible. We adopt a simple formulation based on
 \parencite{Gravel2006}. The approach is conceptually similar to placing a trap
@@
 $1-m-M$ is the probability it comes from local dispersal. The fraction
-$N_{ix}J_x^{-1}$ is the local relative abundance and $P_ix$ is the relative
-abundance of species $i$ in the seed pool coming from neigbouring patches $x$.
+$N_{ix}J_x^{-1}$ is the local relative abundance and $P_{ix}$ is the relative
+abundance of species $i$ in the seed pool coming from patches connected to patch $x$.
 The relative abundance in the neighborhood is weighted by the degree of the
@@
 $P_{ix} =\frac{\sum P_{iy}d_y^{-1}}{\sum d_y^{-1}}$. We assume for simplicity
-(and without loss of generality, \parencite{Bell2000}) that the relative
-abundance in the metacommunity is uniform, i.-e. equal to $S^{-1}$. This
+that the relative abundance in the metacommunity is uniform, i.e. equal to
+$S^{-1}$, but other distributions could be used. The metacommunity is usually
+modelled with a log-series (e.g. \textcite{Hubbell2001}), but it has been shown
+that for a range of realistic dispersal rates a uniform distribution does not
+have distinguishable impacts relative to a log-series \parencite{Bell2000}. This
 immigration prevents the collapse of the metacommunity to a single species,
@@
 (speciation prevents this phenomenon to occur in \textcite{Hubbell2001}). The
-local recruitment probability is consequently $Pr_{ix} = MS^{-1} + mP_{ix} +
-(1-m-M)N_{ix}J_x^{-1}$.
+local recruitment probability $Pr_{ix}$ is consequently $Pr_{ix} = MS^{-1} +
+mP_{ix} + (1-m-M)N_{ix}J_x^{-1}$. Self-replacement occurs when an individual of
+species $i$ is replaced by an offspring of the same species coming either from
+the remaining local individuals or from the metacommunity.
 
@@
 disturbances could be simulated as well, as long as they hit all species with
-the same probability, independently of their abundance. The fundamental feature
+the same probability, independently of their identity. The fundamental feature
 of neutral dynamics is therefore the ecological drift, defined as population
@@
 %------------------------
-\subsection*{Species-sorting and mass effect}
+\subsection*{Species sorting and mass effect}
 
-The species-sorting and the mass effect perspectives build on the notion of
+The species sorting and the mass effect perspectives build on the notion of
 species-specific responses to a spatially varying environment
@@
 The lottery dynamics described above for the neutral model assume there is a
-very large number of offsprings that are candidate for recruitment but only one
-will survive and develop to the adult stage. The effect of a differentiation to
+very large number of offspring that are candidate for recruitment but only one
+will survive and develop to the adult stage. The effect of a specific response to
 local environmental conditions could be implemented at this stage with a biased
@@
 species with highest survival because only some species will be able to cope
-with the local environmental conditions. We define the relative
+with the microsite environmental conditions. We define the relative
 abundance in the seed rain as $Z_{ix} = MS^{-1} + mP_{ix} +
@@
 
-Analyses of similar models with a combination of dispersal and species-sorting
-shown that predictions are extremely variables and depend on the frequency
-distributions of environmental conditions, niche optimums and breadth. For
+Analyses of similar models with a combination of dispersal and species sorting
+shows that predictions are extremely variables and depend on the frequency
+distributions of environmental conditions, niche optimums and niche breadth. For
 instance, a well-studied prediction of neutral models is the species abundance
@@
 given appropriate parameters \parencite{Tilman2004,Gravel2006}. The main
-prediction is nonetheless that stable and predictable (meaning which species
-will coexist) if species are sufficiently dissimilar, which differs from neutral
-models. Local species richness will first depend on the joint effects of local
-heterogeneity and niche breadth because coexistence requires a sufficient
-dissimilaritya mong species \parencite{Schwilk2005}. Local species richness
-could be increased by a mass effect when dispersal is consistently supplying
-individuals coming from more favorable locations (refuges). The limiting
-similarity required to maintain regional coexistence depends on the amount of
-dispersal because exchanges among communities homogenizes environmental
-conditions. This is one of the main result from the species sorting theory and a
-clever example of local-regional feedbacks: increasing dispersal promotes local
-coexistence, but on the other hand it diminishes regional coexistence. Only the
-best average competitors will remain at very high dispersal. We therefore expect
+prediction is nonetheless that stable and predictable coexistence is possible if
+species are sufficiently dissimilar, which differs from neutral models. Local
+species richness will first depend on the joint effects of local heterogeneity
+and niche breadth because coexistence requires a sufficient dissimilarity among
+species \parencite{Schwilk2005}. Local species richness could be increased by a
+mass effect when dispersal is consistently supplying individuals coming from
+more favorable locations (refuges). The limiting similarity required to maintain
+regional coexistence depends on the amount of dispersal because exchanges among
+communities homogenizes environmental conditions. This is one of the main results
+from the species sorting theory and a clever example of local-regional
+feedbacks: increasing dispersal promotes local coexistence, but on the other
+hand it diminishes regional coexistence. Only the best average competitors will
+remain at very high dispersal. We therefore expect
 a hump-shaped relationship between dispersal and alpha ($\alpha$) diversity,
 with a peak at intermediate dispersal. On the other hand, we expect a monotonic
-decrease of beta ($\beta$) and $\gamma$ diversity with dispersal
+decrease of beta ($\beta$) and $\gamma$ diversity with increasing dispersal
 \parencite{Mouquet2003}. This prediction has been validated in some experiments
@@
 
-\section*{Results}
+%========================================================%
+\section*{Illustrative examples}
 
 In this section we provide simple simulation results to illustrate the impact of
-saptial contingencies on species distribution and coexistence. We consider four
+spatial contingencies on species distribution and coexistence. We consider four
 different landscapes, illustrated at Fig. 1. with the outcome of simulations
-using the neutral model. All of these networks have the same number of nodes
-(e.g. spatial sampling sites), but both different number of edges (e.g.
-dispersal routes between sampling sites) and patterns of connectivity between
-nodes. We ask how these differences in connectivity will shape the emerging
-properties of the community under the scenarios represented by each
-metacommunity model. Our analysis is not exhaustive, it is provided simply to
-illustrate the interaction between metacommunity perspectives and landscape
-structures on $\alpha$, $\beta$ and $\gamma$ diversity.
+using the neutral model. All of these networks have the same number of nodes,
+but differ both in number of edges (e.g. dispersal routes between sampling
+sites) and connectivity between nodes. We ask how these differences
+in topological structures will shape the emerging properties of the community under the
+scenarios represented by each metacommunity model. Our analysis is not
+exhaustive, it is provided simply to illustrate the interaction between
+metacommunity perspectives and landscape structures on $\alpha$, $\beta$ and
+$\gamma$ diversity.
 
@@
 relationship between centrality and richness ought to be looked at. It appears
-that both in the geographical and tree graph, the path dynamics model has a
+that both in the random geographical and tree networks, the path dynamics model has a
 much more considerable variation in $\alpha$ diversity. However, in all
@@
 that nodes with more connections also host a more diverse community.
-Eigen-centrality gave a far less clear-cut result, which can probably be
+Eigen-centrality gave a less clear-cut result, which can probably be
 attributed to the fact that our networks are relatively small in size.
@@
 diversity as a function of the shape of the network, under the three dynamic
-models. We used Bray-Curtis measure of dissimilarity between patches. In Fig. 3,
-the distance is expressed as the Euclidean (geographic) distance between two
-patches. Although this neglects how dispersal connects the different patches,
-there is already a clear signal of geographic distance on $\beta$ diversity,
-indicating the importance of dispersal under the three scenarios. In both the
-neutral and patch dynamics model, local communities become increasingly
+models. The connected graph was not considered because there is no variation in
+path length. We used Bray-Curtis measure of dissimilarity between patches. In
+Fig. 3, the distance is expressed as the Euclidean (geographic) distance between
+two patches. Although this neglects how dispersal connects the different
+patches, there is already a clear signal of geographic distance on $\beta$
+diversity, indicating the importance of dispersal under the three scenarios. In
+both the neutral and patch dynamics models, local communities become increasingly
 dissimilar when the distance between them increases. In other words, two
@@
 proportion. The relationship between distance and dissimilarity is similar for
-species-sorting. Nonetheless, it forms an enveloppe of points (with most points
+species-sorting. Nonetheless, it forms an envelope of points (with most points
 lying in the upper-left part of the graph). While two distant communities will
@@
 distribution of microsites (not shown). At one extreme, if all patches hold the
-same average conditions, then we should expect no relationship between
+same average and variance in microsite conditions, then we should expect no relationship between
 dissimilarity and distance. On the other hand, if the average conditions are
@@
 
-Our objective in this paper was to review the main assumptions of three
+The main objective of this paper was to review the main assumptions of three
 metacommunity models and illustrate how the implementation of more realistic
@@
 predict the impact of dispersal on community organization. The review of the
-different models shows that the fundamental difference between a neutral model
+different models shows that the fundamental difference between the neutral model
 and the patch dynamics model is the effect of competitive interactions on
-distribution, while the difference between the neutral and the species sorting
+distribution, while the difference between neutral and species sorting
 models is the effect of unequal competitive interactions. The neutral model is
@@
 \parencite{Legendre2005}) to topological based analyses
-\parencite{Peterson2013}.
+\parencite{dal10, Peterson2013}.
 
-The network approach to the study of spatially explicit landscapes is a major
+The network approach to the study of spatially explicit landscapes was a major
 advancement in metacommunity ecology. It is a first step to make the concept
 operational because it accounts for more realistic landscape structures and
-dispersal kernels. It makes a significant departure to simple island-mainland or
+dispersal kernels. It is a significant departure to island-mainland or
 global dispersal approaches used previously (e.g. \parencite{Tilman1994a,
@@
 lattice model and it does not make the landscape more realistic. We believe the
-fundamental contribution of this approach is the consideration of spatial
+fundamental contribution of this approach is accounting for spatial
 heterogeneity of dispersal. In agreement with previous theoretical
@@
 \parencite{Carrara2012}, the simulations show that the degree centrality has a
-significant impact on $\alpha$ diversity. Central nodes might also contribute
-more to maintain $\gamma$ diversity, as they are essential for species to spread
+significant impact on $\alpha$ diversity. Central nodes might also be important
+to maintain $\gamma$ diversity, as they promote spreading
 throughout the landscape. Such nodes could be potentially quantified as keystone
@@
 surprisingly, this effect is weaker with species sorting dynamics. We could even
-hypothesize it will vanish in the extreme case of niche differentiation (with
-low overlap for instance) and low mass effect. In this particular case, the
+hypothesize it will vanish with very strong niche differentiation (which would occur with
+low niche overlap for instance) and low mass effect. In this particular case, the
 neutral versus niche comparison therefore illustrates that very strong unequal
@@
 The network approach and the comparison between metacommunity perspectives
-reveal there could be spatial variation in coexistence mechanisms. If we take
-the species-sorting perspective for instance, we find that $\alpha$ diversity is
-higher in more central nodes. Since the environment is on average the same from
-one patch to another, it implies that diversity in these communities is
-maintained by a stronger mass effect. It results in spatial variation in the
+reveals there could be spatial variation in coexistence mechanisms. If we take
+the species-sorting perspective for instance, we find that $\alpha$ diversity
+could be higher in more central nodes under some dispersal rates. Since the
+environment is on average the same from one patch to another, and thus should
+hold a similar number of species, it implies that diversity in these communities
+is maintained by a stronger mass effect. It results in spatial variation in the
 relative importance of species-sorting, the mass effect and to a certain extent
@@
 envision to parametrize them and simulate real landscapes. The recruitment
-probabilities defined above could be used as statistical models (likelihood
+probabilities defined above could all be used as statistical models (likelihood
 functions) to fit to empirical data. Prior information could be used to define
@@
 widely used over the last decade as a test of the neutral theory, its underlying
-assumption and a quantification of dispersal limitations (e.g.
-\textcite{Svenning2004, Hardy2004, Gilbert2004, Cottenie2005}. One problem of
-this approach is however that it makes a weak test of neutrality
-\parencite{McGill2003}, based on the description of spatial community structure,
-rather than hypothesis testing. The different models we reviewed in this article
-could be better employed if used to generate null expectations of species
-distribution based on different hypotheses and then compare them. But most of
-all, parametrized spatially explicit neutral models could be more useful if used
-to predict biodiversity under different global change scenarios. For instance,
-neutral models could be used to predict the consequences of habitat destruction,
+assumption and to quantify dispersal limitations (e.g. \textcite{Svenning2004,
+Hardy2004, Gilbert2004, Cottenie2005}. This approach is however only a weak test
+of neutrality \parencite{McGill2003}, based on the description of spatial
+community structure, rather than hypothesis testing. The different models we
+reviewed in this article could be better employed if used to generate null
+expectations of species distribution based on different hypotheses and then
+compare them. But most of all, parametrized spatially explicit neutral models
+could be useful if used to predict biodiversity under different global change
+scenarios. For instance in freshwater systems, neutral models combined with a
+dendritic network analysis were shown to predict well the distribution of
+diversity in central United-States \parencite{Muneepeerakul2008}. Fitted neutral
+models could be used to predict the consequences of habitat destruction,
 fragmentation or a change in the connectivity matrix \parencite{Hubbell2008}.
@@
 Real landscapes are not flat geometric objects, they are highly structured and
-diverse. This diverse structure has long been recognized as a key component of
-diversity. In the 19th century, Moritz Wagner noted that patterns of rivers
-\cite{coy04} could explain how beetles diversified. Yet, to this day, the
+diversified. This structure has long been recognized as a key component of
+biological diversity. In the 19th century, Moritz Wagner noted that patterns of
+rivers \cite{coy04} could explain how beetles diversified. Yet, to this day, the
 relationship between spatial structures and biodiversity is not resolved, but
 networks provide powerful tools to analyze landscapes and generate testable
-predictions on the relationship between community assembly and spatial
-structures. For example, the neutral theory predict less diversity in isolated
-communities and constant speciation (regardless of isolation) \parencite{Economo2008}.
-Adding the effect of gene flow changes the predictions, but the neutral theory
-do not predict adaptive radiation and might thus face problems for predicting
-many long-term ecological processes.
+predictions on community assembly. For example, the neutral theory predicts less
+diversity in isolated communities and constant speciation (regardless of
+isolation) \parencite{Economo2008}. Adding the effect of gene flow changes the
+predictions \parencite{Desjardins2012b}, but the neutral theory does not predict
+adaptive radiation and might thus face problems for predicting many long-term
+evolutionary processes.
 
-Finally, our analyses emphasize the need to expand on the canonial neutral
+Finally, our analyses emphasize the need to expand on the canonical neutral
 theory. As pointed out by \textcite{Wootton2005}, most of the unexplained
-deviation of empirical community from the prediction of accurately calibrated
+deviation of empirical communities from the predictions of accurately calibrated
 neutral models can be attributed to non-competitive interactions. \textcite{Canard2012}
@@
 interactions with a good accuracy. Incorporating reasonable complexity in the
-mechanisms addressed by neutral models is not a theoreticians' exercise: it will
+mechanisms addressed by neutral models is not a theoreticial exercise: it will
 re-enforce the usefulness of the neutral theory as an operational concept,
@@
 dissimilarity at the between-site and regional scales. These predictions are the
-benchmark against which empirical relev\'es of species richness and community
+benchmark against which empirical surveys of species richness and community
 structure ought to be compared, and coming up with realistic parameters to
@@
 steps.
+Note that over replicated runs the global dispersal and the lattice scenarios would have a very regular distribution of diversity (spatially uniform, with declining diversity at the edges for the lattice), while the structure should be maintained for the two random networks.
 
@@
 0.01$, $k = 0.1$, $ u \epsilon [0,100]$, $b = 15$, $ E_x \epsilon [0,100]$,
-$\overline{E_R} = 50$, $\sigma{E_R} = 5$. Simulations run 1000 time steps
+$\overline{E_R} = 50$, $\sigma{E_R} = 5$. Simulations run 1000 time steps.
 
@@
 \subsection*{Figure 3}
-\textbf{Bray curtis dissimilarity as a function of Euclidean distance}. Parameters as in Fig. 2.
+\textbf{Bray curtis dissimilarity as a function of Euclidean distance}.
+The dissimilarity among all pairs of local communities is illustrated as a function of the distance for three types of spatial networks and three metacommunity perspectives.
+Parameters as in Fig.
+2.
 
@@
 \subsection*{Figure 4}
-\textbf{Bray curtis dissimilarity as a function of geodesic distance}. Parameters as in Fig. 2.
+\textbf{Bray curtis dissimilarity as a function of geodesic distance}.
+The dissimilarity among all pairs of local communities is illustrated as a function of the distance for three types of spatial networks and three metacommunity perspectives.
+Parameters as in Fig.
+2.
 
@@
 %------------------------
+
 \subsection*{Figure 2}
@@
 \begin{figure}[ht!]
-	\centering\includegraphics[width=0.75\textwidth]{Centrality.pdf}
+	\centering\includegraphics[width=0.85\textwidth]{Centrality.pdf}
 \end{figure}
@@
 \begin{figure}[ht!]
-	\centering\includegraphics[width=0.65\textwidth]{BetaGeoDist.png}
+	\centering\includegraphics[width=0.85\textwidth]{BetaGeoDist.png}
 \end{figure}
@@
 \begin{figure}[ht!]
-	\centering\includegraphics[width=0.65\textwidth]{BetaTopoDist.png}
+	\centering\includegraphics[width=0.85\textwidth]{BetaTopoDist.png}
 \end{figure}
@@
 		Geodesic distance & Length of the shortest path between two nodes \\
-
+		\\
 		\textbf{Network-level metrics} & \\
@@
 		Components & The number of connected subsets\\
-
+		\\
 		\textbf{Node-level metrics} & \\
 		Degree & The number of edges of a node \\
-		Closeness centrality & Average geodesic distance between a node and all other vertex \\
+		Closeness centrality & Average geodesic distance between a node and all other vertices \\
 		Eigenvector centrality & A measure of centrality based on the concept that connection to
@@
 		Betweenness centrality & The number of shortest paths from all nodes to all others that
-		pass through that note \\
+		pass through that node \\
 		\hline
 	\end{tabular}
-	\caption{Main descriptors of spatial networks used in this study.}
+	\caption{Main descriptors of spatial networks used in this study.
+In the case of spatial networks describing dispersal, nodes refer to local communities and edges to a connexion by dispersal between them.}
 \end{table*}
@@
 
-\begin{table*}[c]
+\begin{table*}[l]
 	\centering
@@
 	$d$ & Node degree & X & X & X \\
-	$C$ & Prob. of a colonization event & X & & \\
-	$I$ & Prob. of a colonization event & X & & \\
+	$C$ & Col. prob. from patch $y$ to patch $x$ & X & & \\
+	$I$ & Col. prob. at patch $x$ given neighbours & X & & \\
 	$Pr$ & Recruitment prob. & & X & X \\
@@
 	\textbf{Parameters} & & & &\\
+	$r$ & Number of patches & X & X & X \\
 	$S$ & Size of regional species pool & X & X & X \\
@@
 	$\sigma$ & Local env. variance & & & X \\
-	$\overline{E_R}$ & Regioal env. average & & & X \\
+	$\overline{E_R}$ & Regional env. average & & & X \\
 	$\sigma_R$ & Regional env. variance & & & X \\

### 2013-09-24 submission1/ms.tex
Minor_revisions

--- 2013-05-30 ms.tex
@@
 \section*{Abstract} 
-The metacommunity perspective appears as an appropriate conceptual
-framework to make ecology more predictive. The recent development of neutral theory appears as a step back in that
-direction because of the assumption of ecological equivalence and the absence of
-any effect of the environment on community organization. A remarkable strength
-of neutral theory is nonetheless to provide a general theory of diversity that
-accounts for a wide range of empirical observations. In this paper, we argue
-that neutral theory can be useful to understand the impact of dispersal
-on community assembly in landscapes of various complexities. Our main
-objective is to use neutral theory to stress the importance of complex landscape
-network structure on the distribution of diversity. We refer to the landscape
-organization as a "spatial contingency" that could potentially affect the
-coexistence mechanisms at play. We briefly review the main approaches to
-describe spatial networks and describe three simple toy models of metacommunity
-dynamics. We take this opportunity to review their assumptions and main
-predictions. We then conduct simple simulations of these models to reveal with
-simple examples the impact of spatial network structure on diversity
-distribution. The simulation results show that competitive interactions buffer
-the potential impact of landscape structure. The strongest centrality-species
-richness relationship was observed for the patch dynamics, a model without any
-interactions. On the other hand, strong and unequal competitive interactions
-minimized the effect of centrality. We conclude that the neutral model is thus a
-useful tool to understand the joint effects of dispersal and community
-interactions. Our analysis shows that ecologists must now integrate more
-realistic landscapes when analyzing community assembly from a metacommunity
-perspective.
+The metacommunity perspective appears as an appropriate
+conceptual framework to make ecology more predictive. The recent development of
+neutral theory appears as a step back in that direction because of the
+assumption of ecological equivalence and the absence of any effect of the
+environment on community organization. A remarkable strength of neutral theory
+is nonetheless to provide a general theory of diversity that accounts for a wide
+range of empirical observations. In this paper, we argue that neutral theory can
+be useful to understand the impact of dispersal on community assembly in
+landscapes of various complexities. Our main objective is to use neutral theory
+to stress the importance of complex landscape structure on the
+distribution of diversity. We refer to the landscape organization as a "spatial
+contingency" that could potentially affect the coexistence mechanisms at play.
+We briefly review the main approaches to describe spatial networks and describe
+three simple toy models of metacommunity dynamics. We take this opportunity to
+review their assumptions and main predictions. We then conduct 
+simulations of these models to reveal with simple examples the impact of spatial
+network structure on diversity distribution. The simulation results show that
+competitive interactions buffer the potential impact of landscape structure. The
+strongest centrality-species richness relationship was observed for the patch
+dynamics, a model without any interactions. On the other hand, strong and
+unequal competitive interactions minimized the effect of centrality. We conclude
+that the neutral model is a useful tool to understand the joint effects of
+dispersal and community interactions. Our analysis shows that ecologists must
+now integrate more realistic landscapes when analyzing community assembly from a
+metacommunity perspective. 
 \newpage
@@
 Neutral theory of biodiversity makes the provocative assumption that species are
-ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral ecological
-communities are driven only by demographic stochasticity and dispersa and thus,
+ecologically equivalent \parencite{Bell2000,Hubbell2001}. Neutral community
+dynamics are driven only by demographic stochasticity and dispersal and thus,
 variation in the environment has no impact on demography. Neutral theory sparked
@@
 species and promote ecological drift \parencite{Gravel2011a}. Much has been said
-the existence of demographic stochasticity \parencite{Clark2012}, and we
-therefore will let this discussion for other papers. A second answer is that
+about the existence of demographic stochasticity \parencite{Clark2012}, and we
+will therefore let this discussion for other papers. A second explanation is that
 dispersal and historical contingencies might have a much more profound impact on
@@
 landscape network structure on the distribution of diversity. We refer to the
-landscape organization as a \emph{spatial contingency} \parencite{Peres-Neto2012}
-that could potentially affect the coexistence mechanisms at play. We will
-therefore move from a perspective where dispersal is either global or spatially
-explicit (e.g. over a lattice), and spatial constant, to a perspective focusing
+landscape organization as a \emph{spatial contingency}
+\parencite{Peres-Neto2012} that could potentially affect the coexistence
+mechanisms at play. We will therefore move from a perspective where dispersal is
+either global or constant over space (e.g. a lattice), to a perspective focusing
 on the variance of dispersal. A second generation of neutral models (e.g.
@@
 and $C$ are more likely to be connected than two random vertices. Random
-geometric networks have been extensively studied \parencite{app97a,app97b,app02a,app02b,pen03} 
-and we provide a R function to generate them. The
+geometric networks have been extensively studied \parencite{app97a, app97b,
+app02a, app02b, pen03} and we provide a R function to generate them. The
 position of nodes is typically random, but we could also imagine alterations
@@
 tree. The algorithm first builds a random geometric graph, then select a node
-from which to start the tree. It then calculaes the the shortest path tree
-\parencite{dij59} from this node to all other ones. This random geometric tree
-does not exactly represent dendritic landscapes but is a convenient model to
-simulate a lake connected by rivers to a series of smaller lakes.
+from which to start the tree. It then calculates the the shortest path tree
+\parencite{dij59} from this node to all other ones and remove edges not located
+along this tree. This random geometric tree does not exactly represent dendritic
+landscapes but is a convenient model to simulate a lake connected by rivers to a
+series of smaller lakes.
 
 Spatial graphs are increasingly popular in spatial ecology and conservation
-biology, where patterns of connections can be used to study and influence the
-flow of organisms \parencite{min07,fal07,min08,gar08,urb09,dal10}. In the neutral
-theory, networks were pioneered by Economo and Keitt \parencite{Economo2008, eco10}. They
-used networks to study how different spatial structures influenced diversity.
-They were also used to study how the spatial structure influenced nonsympatric
-speciation \parencite{Desjardins2012a,Desjardins2012b}.
+biology, where the structure of connections can be used to study and influence
+the flow of organisms \parencite{min07,fal07,min08,gar08,urb09,dal10}. In the
+neutral theory, networks were pioneered by Economo and Keitt
+\parencite{Economo2008, eco10}. They used networks to study how different
+spatial structures influenced diversity. They were also used to study how the
+spatial structure influenced nonsympatric speciation
+\parencite{Desjardins2012a,Desjardins2012b}.
 
@@
 the above formulation of a colonization event to occur, the probability that an
-empty location is colonized is $I_ix=1-\prod d_x(1-C_{ixy})$. This equation
+empty location is colonized is $I_{ix}=1-\prod d_x(1-C_{ixy})$. This equation
 basically tells us that the colonization probability will increase
@@
 to show from metapopulation theory that the occurrence probability in a patch is
-then $p_ix=I_x(I_ix+e)^{-1}$. The feedback between local and regional dynamics
-arises because all $p_ix$ from the landscape are dependent from each other.
+then $p_{ix}=I_x(I_{ix}+e)^{-1}$. The feedback between local and regional dynamics
+arises because all $p_{ix}$ from the landscape are dependent from each other.
 Simulations are usually conducted to solve the model for a large landscape, but
@@
 $1-m-M$ is the probability it comes from local dispersal. The fraction
-$N_{ix}J_x^{-1}$ is the local relative abundance and $P_ix$ is the relative
+$N_{ix}J_x^{-1}$ is the local relative abundance and $P_{ix}$ is the relative
 abundance of species $i$ in the seed pool coming from neigbouring patches $x$.
@@
 
-Analyses of similar models with a combination of dispersal and species-sorting
+Analyses of similar models with a combination of dispersal and species sorting
 shown that predictions are extremely variables and depend on the frequency
@@
 given appropriate parameters \parencite{Tilman2004,Gravel2006}. The main
-prediction is nonetheless that stable and predictable (meaning which species
-will coexist) if species are sufficiently dissimilar, which differs from neutral
-models. Local species richness will first depend on the joint effects of local
-heterogeneity and niche breadth because coexistence requires a sufficient
-dissimilaritya mong species \parencite{Schwilk2005}. Local species richness
-could be increased by a mass effect when dispersal is consistently supplying
-individuals coming from more favorable locations (refuges). The limiting
-similarity required to maintain regional coexistence depends on the amount of
-dispersal because exchanges among communities homogenizes environmental
-conditions. This is one of the main result from the species sorting theory and a
-clever example of local-regional feedbacks: increasing dispersal promotes local
-coexistence, but on the other hand it diminishes regional coexistence. Only the
-best average competitors will remain at very high dispersal. We therefore expect
+prediction is nonetheless that stable and predictable coexistence is possible if
+species are sufficiently dissimilar, which differs from neutral models. Local
+species richness will first depend on the joint effects of local heterogeneity
+and niche breadth because coexistence requires a sufficient dissimilarity among
+species \parencite{Schwilk2005}. Local species richness could be increased by a
+mass effect when dispersal is consistently supplying individuals coming from
+more favorable locations (refuges). The limiting similarity required to maintain
+regional coexistence depends on the amount of dispersal because exchanges among
+communities homogenizes environmental conditions. This is one of the main result
+from the species sorting theory and a clever example of local-regional
+feedbacks: increasing dispersal promotes local coexistence, but on the other
+hand it diminishes regional coexistence. Only the best average competitors will
+remain at very high dispersal. We therefore expect
 a hump-shaped relationship between dispersal and alpha ($\alpha$) diversity,
@@
 
-\section*{Results}
+%========================================================%
+\section*{Illustrative examples}
 
@@
 different landscapes, illustrated at Fig. 1. with the outcome of simulations
-using the neutral model. All of these networks have the same number of nodes
-(e.g. spatial sampling sites), but both different number of edges (e.g.
-dispersal routes between sampling sites) and patterns of connectivity between
-nodes. We ask how these differences in connectivity will shape the emerging
-properties of the community under the scenarios represented by each
-metacommunity model. Our analysis is not exhaustive, it is provided simply to
-illustrate the interaction between metacommunity perspectives and landscape
-structures on $\alpha$, $\beta$ and $\gamma$ diversity.
+using the neutral model. All of these networks have the same number of nodes,
+but both different number of edges (e.g. dispersal routes between sampling
+sites) and connectivity between nodes. We ask how these differences
+in topological structures will shape the emerging properties of the community under the
+scenarios represented by each metacommunity model. Our analysis is not
+exhaustive, it is provided simply to illustrate the interaction between
+metacommunity perspectives and landscape structures on $\alpha$, $\beta$ and
+$\gamma$ diversity.
 
@@
 that nodes with more connections also host a more diverse community.
-Eigen-centrality gave a far less clear-cut result, which can probably be
+Eigen-centrality gave a ess clear-cut result, which can probably be
 attributed to the fact that our networks are relatively small in size.
@@
 predict the impact of dispersal on community organization. The review of the
-different models shows that the fundamental difference between a neutral model
-and the patch dynamics model is the effect of competitive interactions on
-distribution, while the difference between the neutral and the species sorting
+different models shows that the fundamental difference between neutral model
+and patch dynamics model is the effect of competitive interactions on
+distribution, while the difference between neutral and species sorting
 models is the effect of unequal competitive interactions. The neutral model is
@@
 \parencite{Legendre2005}) to topological based analyses
-\parencite{Peterson2013}.
+\parencite{dal10, Peterson2013}.
 
@@
 operational because it accounts for more realistic landscape structures and
-dispersal kernels. It makes a significant departure to simple island-mainland or
+dispersal kernels. It is a significant departure to island-mainland or
 global dispersal approaches used previously (e.g. \parencite{Tilman1994a,
@@
 reveal there could be spatial variation in coexistence mechanisms. If we take
-the species-sorting perspective for instance, we find that $\alpha$ diversity is
-higher in more central nodes. Since the environment is on average the same from
-one patch to another, it implies that diversity in these communities is
-maintained by a stronger mass effect. It results in spatial variation in the
+the species-sorting perspective for instance, we find that $\alpha$ diversity
+could be higher in more central nodes under some dispersal rates. Since the
+environment is on average the same from one patch to another, and thus should
+hold a similar number of species, it implies that diversity in these communities
+is maintained by a stronger mass effect. It results in spatial variation in the
 relative importance of species-sorting, the mass effect and to a certain extent
@@
 widely used over the last decade as a test of the neutral theory, its underlying
-assumption and a quantification of dispersal limitations (e.g.
-\textcite{Svenning2004, Hardy2004, Gilbert2004, Cottenie2005}. One problem of
-this approach is however that it makes a weak test of neutrality
-\parencite{McGill2003}, based on the description of spatial community structure,
-rather than hypothesis testing. The different models we reviewed in this article
-could be better employed if used to generate null expectations of species
-distribution based on different hypotheses and then compare them. But most of
-all, parametrized spatially explicit neutral models could be more useful if used
-to predict biodiversity under different global change scenarios. For instance,
-neutral models could be used to predict the consequences of habitat destruction,
-fragmentation or a change in the connectivity matrix \parencite{Hubbell2008}.
-The spatially explicit description of the landscape is a major improvement
-toward that end, providing much flexibility in the scenarios that could be
-explored.
+assumption and to quantify dispersal limitations (e.g. \textcite{Svenning2004,
+Hardy2004, Gilbert2004, Cottenie2005}. This approach is however only a weak test
+of neutrality \parencite{McGill2003}, based on the description of spatial
+community structure, rather than hypothesis testing. The different models we
+reviewed in this article could be better employed if used to generate null
+expectations of species distribution based on different hypotheses and then
+compare them. But most of all, parametrized spatially explicit neutral models
+could be useful if used to predict biodiversity under different global
+change scenarios. For instance, neutral models could be used to predict the
+consequences of habitat destruction, fragmentation or a change in the
+connectivity matrix \parencite{Hubbell2008}. The spatially explicit description
+of the landscape is a major improvement toward that end, providing much
+flexibility in the scenarios that could be explored.
 
@@
 Real landscapes are not flat geometric objects, they are highly structured and
-diverse. This diverse structure has long been recognized as a key component of
-diversity. In the 19th century, Moritz Wagner noted that patterns of rivers
-\cite{coy04} could explain how beetles diversified. Yet, to this day, the
+diversified. This structure has long been recognized as a key component of
+biological diversity. In the 19th century, Moritz Wagner noted that patterns of
+rivers \cite{coy04} could explain how beetles diversified. Yet, to this day, the
 relationship between spatial structures and biodiversity is not resolved, but
 networks provide powerful tools to analyze landscapes and generate testable
-predictions on the relationship between community assembly and spatial
-structures. For example, the neutral theory predict less diversity in isolated
-communities and constant speciation (regardless of isolation) \parencite{Economo2008}.
-Adding the effect of gene flow changes the predictions, but the neutral theory
-do not predict adaptive radiation and might thus face problems for predicting
-many long-term ecological processes.
+predictions on community assembly. For example, the neutral theory predict less
+diversity in isolated communities and constant speciation (regardless of
+isolation) \parencite{Economo2008}. Adding the effect of gene flow changes the
+predictions \parencite{Desjardins2012b}, but the neutral theory does not predict
+adaptive radiation and might thus face problems for predicting many long-term
+evolutionary processes.
 
-Finally, our analyses emphasize the need to expand on the canonial neutral
+Finally, our analyses emphasize the need to expand on the canonical neutral
 theory. As pointed out by \textcite{Wootton2005}, most of the unexplained
-deviation of empirical community from the prediction of accurately calibrated
+deviation of empirical communities from the predictions of accurately calibrated
 neutral models can be attributed to non-competitive interactions. \textcite{Canard2012}

### 2013-09-24 revisions/supplementary_code/examples.R

@@ -1,0 +1,90 @@
+########################################################################
+# Examples of dynamics on spatial graphs with different metacommunity perspectives
+#
+# By: Dominique Gravel (dominique_gravel@uqar.ca) 
+# Last edited September 2013
+# For latest version:
+# https://github.com/DominiqueGravel/ms_neutral_theory
+#
+########################################################################
+###############################
+# General parameters
+S = 100
+n = 25
+r = 0.3
+nsteps = 100
+
+########################################
+########################################
+# Patch dynamics example
+source("spatial_graphs.R")
+source("patch_model.R")
+c = 0.2
+e = 0.1
+spatial_graph = geograph(n, r)
+results = patch_model(c, e, S, spatial_graph, nsteps)
+
+# Figure 1: time series
+x11(height = 5.5, width = 6)
+plot(c(1:nsteps),apply(results[[2]][,2:(S+1)],1,sum),xlab = "Time", ylab = "Average local species richness",cex.lab = 1.5, cex.axis = 1.25, type = "l", ylim = c(0,S))
+
+# Figure 2: local species richness
+s = apply(results[[1]],1,sum)
+vec.col = numeric(length(s))
+RK = rank(s)
+for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
+plot_spatial(spatial_graph, vec.col)
+
+########################################
+########################################
+# Neutral example
+source("spatial_graphs.R")
+source("lottery_model.R")
+# Default parameters
+m = 0.2
+M = 0.01
+k = 0.1
+J = 100
+sdN = Inf
+sdE = 5
+spatial_graph = geograph_fn(n,r)
+results = lottery_model(m,M,k,S,J,sdE,sdN,spatial_graph,nsteps)
+
+# Figure 1: time series
+x11(height = 5.5, width = 6)
+plot(c(1:nsteps),apply(results[[2]][,2:(S+1)],1,sum),xlab = "Time", ylab = "Average local species richness",cex.lab = 1.5, cex.axis = 1.25, type = "l", ylim = c(0,S))
+
+# Figure 2: local species richness
+s = apply(results[[1]],1,sum)
+vec.col = numeric(length(s))
+RK = rank(s)
+for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
[…]

### 2013-09-24 revisions/supplementary_code/lottery_model.R

@@ -1,0 +1,123 @@
+#######################################################################
+# Individual based lottery model with spatially explicit
+# dispersal and environmental heterogeneity
+# 
+# The model collapse to a neutral model when the niche breadth tends to infinity
+#
+# Store the results in a time series of occupancy, a presence-absence matrix
+# and a local abundance matrix
+# 
+# By: Dominique Gravel (dominique_gravel@uqar.ca) 
+# Last edited September 2013
+# For latest version:
+# https://github.com/DominiqueGravel/ms_neutral_theory
+########################################################################
+
+# Main function
+lottery_model = function(m,M,d,S,J,sdE,sdN,spatial_graph,nsteps) {
+	# Args: 
+	# 	m: immigration probability from the neighbourhood
+	#	M: immigration probability from outside the metacommunity 
+	#	k: local death rate 	
+	# 	S: number of species
+	#	J: local community size
+	# 	sdE: within patch standard deviation of the environment
+	# 	sdN: niche breadth
+	#	spatial_graph: a spatial graph object
+	#	nsteps: number of time steps to run the simulation
+	#
+	# Returns:
+	#	A list with a time series of occupancy for each species, a site presence-absence matrix	
+	#	and a site-abundance matrix from the last time step
+	#	
+	########################################
+	# Prepare the simulation
+	
+	####################
+	# Adjency matrix
+	adjMat = spatial_graph[[2]] 	# The original matrix
+	n = nrow(adjMat)				# Number of nodes
+	degrees = apply(adjMat,2,sum) 	# Number of degrees for each node
+	w = adjMat*matrix(degrees^-1,nr = n,nc=n, byrow=T) # Weighted adjancy matrix
+
+	####################
+	# Local environmental conditions
+	Env = matrix(nr=n, nc = J)
+	for(i in 1:n) Env[i,] = rnorm(J,mean = runif(1,0,100), sd = sdE) 	# For a random distribution of environments
+#	for(i in 1:n) Env[i,] = rnorm(J,mean = i/n*100, sd = 0)				# For a uniform distribution of environments		
+	####################
+	# Niche optimums
+	u = runif(S, 0, 100) 	# For a random distribution of optimums
[…]

### 2013-09-24 revisions/supplementary_code/patch_model.R

@@ -1,0 +1,84 @@
+########################################################################
+# Multi-species discrite-time Levins metapopulation model
+# Spatially explicit local dispersal
+# Spatially uniform environment
+# 
+# Store the results in a time series of occupancy
+# and a presence absence matrix
+#
+# By: Dominique Gravel (dominique_gravel@uqar.ca) 
+# Last edited September 2013
+# For latest version:
+# https://github.com/DominiqueGravel/ms_neutral_theory
+########################################################################
+
+patch_model = function(c, e, S, spatial_graph, nsteps) {
+	# Runs a muti-species Levins model on a spatial graph
+	#
+	# Args: 
+	# 	c: colonization probability 
+	#	e: extinction probability
+	# 	S: number of species
+	#	spatial_graph: a spatial graph object
+	#	nsteps: number of time steps to run the simulation
+	#
+	# Returns:
+	#	A list with a time series of occupancy for each species and a presence-absence
+	# 	matrix from the last time step
+	#
+	########################################
+	# Prepare the simulation
+	
+	####################
+	# Adjency matrix
+	adjMat = spatial_graph[[2]] # The original matrix
+	n = nrow(adjMat)	# Number of nodes
+	degrees = apply(adjMat,2,sum) # Number of degrees for each node
+	ColProb =  c*adjMat*matrix(degrees^-1,nr = n,nc=n, byrow=T) # Colonization probability weighted by the number of out degrees
+
+	####################
+	# Initialization of the metacommunity in the presence-absence matrix pres
+	# Global distribution at the beginning of the run
+	pres = matrix(1, nr = n, nc = S)
+
+	####################
+	# Matrix in which we record the occupancy over time  
+	Series = matrix(nr=nsteps,nc=S+1)
+	
+	########################################
+	# Loop over all time steps
+	for(time in 1:nsteps) {
+		
+		####################
+		# Test if there is extinction
+		ExtMat = matrix(0,nr=n,nc=S)
+		randExt = matrix(runif(n*S,0,1),nr=n,nc=S)
+		ExtMat[pres == 1 & randExt < e] = -1
+
+		####################		
+		# Test if there is colonization
+		ColMat = matrix(0,nr=n,nc=S)
+		randCol = matrix(runif(n*S,0,1),nr=n,nc=S)
+
+		# Calculate how many populations are connected to the focal patch
[…]

### 2013-09-24 revisions/supplementary_code/simulations (enregistré automatiquement).R

@@ -1,0 +1,320 @@
+########################################################################
+# Source code for the figures 1-4 in the paper
+# Gravel, Poisot and Desjardins. 2013. Using neutral theory to reveal the contribution of dispersal to community assembly in spatially explicit landscapes. J. Limnology. 
+#
+# By: Dominique Gravel (dominique_gravel@uqar.ca) 
+# Last edited September 2013
+# For latest version:
+# https://github.com/DominiqueGravel/ms_neutral_theory
+########################################################################
+source("spatial_graphs.R")
+source("patch_model.R")
+source("lottery_model.R")
+library(vegan)
+
+###############################
+# Useful functions
+alpha_fn = function(pres) {
+	s = apply(pres,1,sum)
+	vec.col = numeric(length(s))
+	RK = rank(s)
+	for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
+	return(vec.col)	
+}
+
+plot_alpha = function(spatial_graph, vec.col) {
+  	# Plots a spatial graph 
+  	#
+  	# Args:
+  	#   spatia_graph: the output of one of the spatial graphs
+  	#
+	par(mar=c(1,1,1,1))
+	XY = spatial_graph[[1]]
+	adjMat = spatial_graph[[2]]
+
+	plot(XY[,1],XY[,2],xlab = "", ylab = "",cex = 1.5,labels = F)
+	adjVec = stack(as.data.frame(adjMat))[,1]
+	XX = expand.grid(XY[,1],XY[,1])
+	YY = expand.grid(XY[,2],XY[,2])
+	XX = subset(XX,adjVec==1)
+	YY = subset(YY,adjVec==1)
+	arrows(x0 = XX[,1],x1=XX[,2],y0 = YY[,1], y1 = YY[,2], length = 0,lwd = 0.2, col = "darkgrey")
+	points(XY[,1],XY[,2],pch=21,bg=vec.col,cex = 1.5)
+}
+
+###############################
+# General parameters
+S = 100
+n = 25
+r = 0.3
+
+###############################
+# Draw the four landscapes
[…]

### 2013-09-24 revisions/supplementary_code/simulations.R

@@ -1,0 +1,285 @@
+########################################################################
+# Source code for the figures 1-4 in the paper
+# Gravel, Poisot and Desjardins. 2013. Using neutral theory to reveal the contribution of dispersal to community assembly in spatially explicit landscapes. J. Limnology. 
+#
+# By: Dominique Gravel (dominique_gravel@uqar.ca) 
+# Last edited September 2013
+# For latest version:
+# https://github.com/DominiqueGravel/ms_neutral_theory
+########################################################################
+source("spatial_graphs.R")
+source("patch_model.R")
+source("lottery_model.R")
+library(vegan)
+
+###############################
+# Useful functions
+alpha_fn = function(pres) {
+	s = apply(pres,1,sum)
+	vec.col = numeric(length(s))
+	RK = rank(s)
+	for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
+	return(vec.col)	
+}
+
+plot_alpha = function(spatial_graph, vec.col) {
+  	# Plots a spatial graph 
+  	#
+  	# Args:
+  	#   spatia_graph: the output of one of the spatial graphs
+  	#
+	par(mar=c(1,1,1,1))
+	XY = spatial_graph[[1]]
+	adjMat = spatial_graph[[2]]
+
+	plot(XY[,1],XY[,2],xlab = "", ylab = "",cex = 1.5,labels = F)
+	adjVec = stack(as.data.frame(adjMat))[,1]
+	XX = expand.grid(XY[,1],XY[,1])
+	YY = expand.grid(XY[,2],XY[,2])
+	XX = subset(XX,adjVec==1)
+	YY = subset(YY,adjVec==1)
+	arrows(x0 = XX[,1],x1=XX[,2],y0 = YY[,1], y1 = YY[,2], length = 0,lwd = 0.2, col = "darkgrey")
+	points(XY[,1],XY[,2],pch=21,bg=vec.col,cex = 1.5)
+}
+
+###############################
+# General parameters
+S = 100
+n = 25
+r = 0.3
+
+###############################
+# Draw the four landscapes
+connected = connected_fn(n)
[…]

### 2013-09-24 revisions/supplementary_code/spatial_graphs.R

@@ -1,0 +1,279 @@
+########################################################################
+# Functions to generate networks ("graphs") for spatial ecology.
+#
+# Stores the graph in a adjacency matrix
+#
+# Functions to generate graphs:
+# - geograph_fn: returns a random geometric graph.
+# - geotree_fn: returns a random geometric tree (of doom!).
+# - lattice_fn: returns a lattice type of spatial graph.
+# - connected_fn: returns a fully connected graph.
+#
+# Useful functions:
+# - plot_spatial: custom function to plot spatial graphs.
+# - deg_cen: computes degree centrality.
+# - cls_cen: computes closeness centrality.
+# - eig_cen: computers eigen-centrality.
+#
+# Associated functions:
+# - connected: tests if the graph is fully connected.
+# - test_con: recursive function used by 'connected'.
+# - spm: Generates the shortest path distance matrix.
+# - spd: Generates the shortest path distances starting with a given source.
+# - spt: Generates the shortest path tree as an adjacency matrix.
+#
+# Examples:
+# > plot_spatial(geograph_fn(n = 25, r = 0.3))
+# > plot_spatial(geotree_fn(n = 25, r = 0.3))
+# > plot_spatial(connected_fn(n = 25))
+# > plot_spatial(lattice_fn(n = 25))
+#
+# By: Dominique Gravel (dominique_gravel@uqar.ca)
+#     Philippe Desjardins-Proulx (philippe.d.proulx@gmail.com)
+#
+# Last edited September 2013
+# For latest version:
+# https://github.com/DominiqueGravel/ms_neutral_theory
+########################################################################
+
+library(igraph)
+
+geograph_fn = function(n = 64, r = 0.32) {
+  # Generates a random geometric graph.
+  #
+  # Args:
+  #   n: Number of vertices.
+  #   r: Threshold distance to connect vertices.
+  #
+  # Returns:
+  #   A list with the xy coordinates and the adjacency matrix.
+  # Note: the algorithm tests if the graph is connected
+  repeat {
+    xy = cbind(runif(n), runif(n))
+    distMat = as.matrix(dist(xy, method = 'euclidean', upper = T, diag = T))
+    adjMat = matrix(0, nr = n, nc = n)
[…]

### 2013-09-24 submission1/supplementary_code/simulations.R

@@ -1,0 +1,279 @@
+########################################################################
+# Source code for the figures 1-4 in the paper
+# Gravel, Poisot and Desjardins. 2013. Using neutral theory to reveal the contribution of dispersal to community assembly in spatially explicit landscapes. J. Limnology. 
+#
+# By: Dominique Gravel (dominique_gravel@uqar.ca) 
+# Last edited September 2013
+# For latest version:
+# https://github.com/DominiqueGravel/ms_neutral_theory
+########################################################################
+source("spatial_graphs.R")
+source("patch_model.R")
+source("lottery_model.R")
+library(vegan)
+
+###############################
+# Useful functions
+alpha_fn = function(pres) {
+	s = apply(pres,1,sum)
+	vec.col = numeric(length(s))
+	RK = rank(s)
+	for(i in 1:n) vec.col[i] = rainbow(n,start = 0, end = 0.7)[RK[i]]
+	return(vec.col)	
+}
+
+plot_alpha = function(spatial_graph, vec.col) {
+  	# Plots a spatial graph 
+  	#
+  	# Args:
+  	#   spatia_graph: the output of one of the spatial graphs
+  	#
+	par(mar=c(1,1,1,1))
+	XY = spatial_graph[[1]]
+	adjMat = spatial_graph[[2]]
+
+	plot(XY[,1],XY[,2],xlab = "", ylab = "",cex = 1.5,labels = F)
+	adjVec = stack(as.data.frame(adjMat))[,1]
+	XX = expand.grid(XY[,1],XY[,1])
+	YY = expand.grid(XY[,2],XY[,2])
+	XX = subset(XX,adjVec==1)
+	YY = subset(YY,adjVec==1)
+	arrows(x0 = XX[,1],x1=XX[,2],y0 = YY[,1], y1 = YY[,2], length = 0,lwd = 0.2, col = "darkgrey")
+	points(XY[,1],XY[,2],pch=21,bg=vec.col,cex = 1.5)
+}
+
+###############################
+# General parameters
+S = 100
+n = 25
+r = 0.3
+
+###############################
+# Draw the four landscapes
+connected = connected_fn(n)
[…]

### 2013-09-25 revisions/ms.tex
Tim revisions

@@
 number of nodes and thus contribute less to the seed pool. We thus consider
-$P_{ix} =\frac{\sum P_{iy}d_y^{-1}}{\sum d_y^{-1}}$. We assume for simplicity
+$P_{ix} =\sum P_{iy}d_y^{-1}/\sum d_y^{-1}$. We assume for simplicity
 that the relative abundance in the metacommunity is uniform, i.e. equal to
@@
 because only a fraction of offspring survive. It is formulated as $Pr_{ix} =
-\frac{\lambda_{inx}Z_{ix}}{\sum \lambda_{jnx}Z_{jx}}$. The function describing
+\lambda_{inx}Z_{ix}/\sum \lambda_{jnx}Z_{jx}$. The function describing
 the relationship between a microsite condition and survival could take various
@@
 with a peak at intermediate dispersal. On the other hand, we expect a monotonic
-decrease of beta ($\beta$) and $\gamma$ diversity with increasing dispersal
+decrease of $\beta$ and $\gamma$ diversity with increasing dispersal
 \parencite{Mouquet2003}. This prediction has been validated in some experiments
@@
 spatial contingencies on species distribution and coexistence. We consider four
-different landscapes, illustrated at Fig. 1. with the outcome of simulations
+different landscapes, illustrated at Fig. 1 with the outcome of simulations
 using the neutral model. All of these networks have the same number of nodes,
@@
 dispersal kernels. It is a significant departure to island-mainland or
-global dispersal approaches used previously (e.g. \parencite{Tilman1994a,
+global dispersal approaches used previously (e.g. \textcite{Tilman1994a,
 Hubbell2001, Mouquet2002}). But dispersal is also spatially explicit in a
@@
 assumption and to quantify dispersal limitations (e.g. \textcite{Svenning2004,
-Hardy2004, Gilbert2004, Cottenie2005}. This approach is however only a weak test
+Hardy2004, Gilbert2004, Cottenie2005}). This approach is however only a weak test
 of neutrality \parencite{McGill2003}, based on the description of spatial
@@
 biological diversity. In the 19th century, Moritz Wagner noted that patterns of
-rivers \cite{coy04} could explain how beetles diversified. Yet, to this day, the
+rivers \parencite{coy04} could explain how beetles diversified. Yet, to this day, the
 relationship between spatial structures and biodiversity is not resolved, but
@@
 diversity in isolated communities and constant speciation (regardless of
-isolation) \parencite{Economo2008}. Adding the effect of gene flow changes the
+isolation -- \parencite{Economo2008}. Adding the effect of gene flow changes the
 predictions \parencite{Desjardins2012b}, but the neutral theory does not predict
@@
 0.01$, $k = 0.1$, $ u \epsilon [0,100]$, $b = 15$, $ E_x \epsilon [0,100]$,
-$\overline{E_R} = 50$, $\sigma{E_R} = 5$. Simulations run 1000 time steps.
+$\overline{E_R} = 50$, $\sigma{E_R} = 5$. Simulations were run 1000 time steps.
 
